### 2020-07-02 analysis/preliminaryAnalysis.Rmd
Initial commit

---
title: "Post-replication citation patterns in psychological science: Preliminary analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = T)
```

```{r load_packages}
library(knitr) # for literate programming
library(papaja) # for article template
library(tidyverse) # for data munging
library(tidylog) # inline analysis feedback
library(here) # for finding files
library(ggpubr) # for arranging plots
library(kableExtra) # for tables 
```

```{r perform_preprocessing}
# loads raw data, performs preprocessing, saves preprocessed files
source(here('analysis','preprocessing.R'))
```

```{r load_data}
# loads the processed data files output by the preprocessing performed above
load(here('data','processed','d_citations.R'))
load(here('data','processed','d_reference.R'))
load(here('data','processed','d_contentAnalysis.R'))
```

```{r load_functions}
# load custom functions
source(here('analysis','functions.R'))
```

```{r add_standardized_citation_counts}
# identify the five different case studies we are looking at
caseNames <- c('baumeister', 'sripada', 'strack', 'carter', 'caruso')
replicationYears <- c(2016, 2016, 2016, 2014, 2014)

# set up some colour palettes
library(RColorBrewer)
p1 <- c('lightgrey','darkgrey', brewer.pal(3, "Pastel2")[1], brewer.pal(3, "Pastel2")[2], brewer.pal(4, "Pastel1")[4])
p2 <- c('lightgrey', brewer.pal(3, "Pastel2")[3], brewer.pal(4, "Pastel1")[1])

# summarizes data at the year level for each case and adds standardized citation counts for target citations and reference class citations
d_summary <- data.frame() # create empty list to hold data frames for each case
for(i in seq(1,5)){ # loop through cases
  d <- standardizeCitations(
    citationData = d_citations %>% filter(case == caseNames[i]), 
    referenceData = d_reference %>% filter(case == caseNames[i]),
    contentData = d_contentAnalysis %>% filter(case == caseNames[i]),
    replicationYear = replicationYears[i])
  d_summary <- bind_rows(d_summary, d) # append to dataframe
}
```

Hi all,

We've now completed all planned data collection and coding for this project. This report contains a preliminary analysis of the data conducted by Tom & Sophia. The study protocol(s) can be found here:

* Original protocol: https://osf.io/eh5qd/
* Amended protocol (we added more up-to-date citation data): https://osf.io/pdvb5/

# Citation count analysis

## Raw citations

Firstly, we'll plot raw citation curves for the five cases to get a flavour of what we are working with. The x-axis starts in the year the target study was published. The black arrow indicates the publication date of the contradictory replication study. The black line represents the raw annual citation count to the target article. Note that the citation data was acquired on April 26th, 2019, which partly contributes to the fall in citations in 2019 in all cases.

Throughout this report, we've made larger plots for the Baumeister and Strack cases because their citation counts are substantially larger than the other cases (Sripada, Carter, Caruso).

```{r fig.height = 6, fig.width = 8}
ggarrange(citationCurve('strack', standardized = F, plotReference = F),
          citationCurve('baumeister', standardized = F, plotReference = F),
          nrow = 2, ncol = 1, common.legend = T) %>% 
  annotate_figure(
    left = text_grob("Raw citation count", rot = 90),
    bottom = text_grob("Publication year"))
```

```{r fig.height = 3, fig.width = 8}
ggarrange(citationCurve('sripada', standardized = F, plotReference = F),
          citationCurve('carter', standardized = F, plotReference = F),
          citationCurve('caruso', standardized = F, plotReference = F),
          nrow = 1, ncol = 3, common.legend = T) %>% 
  annotate_figure(
    left = text_grob("Raw citation count", rot = 90),
    bottom = text_grob("Publication year"))
```

## Standardized citations with reference class

Now we'll plot standardized citation curves and include citations to the set of reference articles (the dashed grey line). In the graphs below, the annual citation counts for both the target article and reference class are standardized against the calendar year with the highest citation count for the target article (i.e., citations received in that year are set at the standardized value of 100). As defined in the protocol, the reference class in each case is all of the articles published in the same year and in the same journal as the target article. For example, for Baumeister et al. (1998), the reference class is all articles published in 1998 in the Journal of Personality and Social Psychology.

```{r fig.height = 6, fig.width = 8}
ggarrange(citationCurve('strack', standardized = T, plotReference = T),
          citationCurve('baumeister', standardized = T, plotReference = T),
          nrow = 2, ncol = 1, common.legend = T) %>% 
  annotate_figure(
    left = text_grob("Standardized citation count", rot = 90),
    bottom = text_grob("Publication year"))
```

```{r fig.height = 3, fig.width = 8}
ggarrange(citationCurve('sripada', standardized = T, plotReference = T),
          citationCurve('carter', standardized = T, plotReference = T),
          citationCurve('caruso', standardized = T, plotReference = T),
          nrow = 1, ncol = 3, common.legend = T) %>% 
  annotate_figure(
    left = text_grob("Standardized citation count", rot = 90),
    bottom = text_grob("Publication year"))
```

## Issues

Before moving on, there are a few things to note or questions I want to raise:

(1) As mentioned above, the fall in citations in 2019 is partly due to the citation data being acquired part way through the year (April 26th, 2019). Now we are in 2020, it should be straightforward to update these curves with all of the 2019 data - however this may create complications for overlaying the qualitative data (next section) on the citation curves because (1) previous years will likely have some minor changes due to database updates (hopefully they'll be so few changes we can update our qualitative assessment without too much trouble); (2) 2019 will be left with a large gap, unless we code additional articles.

(2) Does it make sense to standardize the citation counts to 100 when the raw counts are below 100 for all cases other than Baumeister?

(3) Can we really learn much from the Sripada, Caruso, and Carter cases given the low citation count?

# Qualitative assessment of citation content

We conducted a qualitative assessment of citing articles within a pre-defined "assessment period". The assessment period included the year prior to and all years after publication of the replication. We assessed all citing articles in the assessment period except for in the Baumeister case - due to a very large number of citing articles we only examined a random sample of 40% of the citing articles from the assessment period, specifically 2015 (n = 76) and 2017 to April 2019 (n = 177). In all cases, the assessment period does not include the year in which the replication was published. This is because articles may already be in the publication pipeline for which the authors have no knowledge of the results of the contradictory replication.

Across all cases, `r d_contentAnalysis %>% count(excluded) %>% filter(excluded == T) %>% pull(n)` citing articles had to be excluded because they could not be accessed (n = `r d_contentAnalysis %>% count(exclusionReason) %>% filter(exclusionReason == "No access") %>% pull(n)`), were non-English language (n = `r d_contentAnalysis %>% count(exclusionReason) %>% filter(exclusionReason == "Non-English language") %>% pull(n)`), or because the target article was cited in the references but not in the actual text (n = `r d_contentAnalysis %>% count(exclusionReason) %>% filter(exclusionReason == "Cited in references but not in text") %>% pull(n)`). Excluded articles are represented on the graphs below in light grey.

## Claim correction vs. claim perpetuation

One goal of the study was to see if we would observe a post-replication pattern more consistent with "claim correction" or "claim perpetuation".

A claim correction pattern might involve a decrease in the number of favourable citations reflecting declining field-level endorsement of the original claim. This may be accompanied by an increase in the number of unfavourable citations (an active correction effect) or no change/decline in the number of unfavourable citations (a passive correction effect).

A claim perpetuation pattern might involve an increase or maintain of the number of favourable citations reflecting increasing or maintained field-level endorsement of the original claim. This may be accompanied by an increase in the number of unfavourable citations (a challenged perpetuation effect) or no change/decline in the number of unfavourable citations (an unchallenged perpetuation effect).

To assess this, we classified citations to the target article as favourable, equivocal, unfavourable, or unclassifiable (a general or methodological reference not related to belief in the claim under scruntiny). Each article was coded by two individuals with disagreements resolved through discussion/arbitration. The data shown below represents the agreed upon classifications.

```{r fig.height = 6, fig.width = 8}
ggarrange(citationCurve('strack', areaPlot = 'classification'),
          citationCurve('baumeister', areaPlot = 'classification'),
          nrow = 2, ncol = 1, common.legend = T) %>% 
  annotate_figure(
    left = text_grob("Standardized citation citation count", rot = 90),
    bottom = text_grob("Publication year"))
```

```{r fig.height = 3, fig.width = 8}
ggarrange(citationCurve('sripada', areaPlot = 'classification'),
          citationCurve('carter', areaPlot = 'classification'),
          citationCurve('caruso', areaPlot = 'classification'),
          nrow = 1, ncol = 3, common.legend = T) %>% 
  annotate_figure(
    left = text_grob("Standardized citation citation count", rot = 90),
    bottom = text_grob("Publication year"))
```

Recall that we only examined 40% of the articles in the Baumeister case which is why there's a lot of white space.

It may be helpful to zoom in on the relevant part of the timeline too to take a closer look (we'll just do this for Strack and Baumeister):

```{r fig.height = 3, fig.width = 8}
ggarrange(citationCurve('strack', areaPlot = 'classification', zoom = T),
          citationCurve('baumeister', areaPlot = 'classification', zoom = T),
          nrow = 1, ncol = 2, common.legend = T) %>% 
  annotate_figure(
    left = text_grob("Standardized citation count", rot = 90),
    bottom = text_grob("Publication year"))
```

There's a (to me) surprising preponderance of favourable citations in both cases - consistent with a "claim perpetuation" pattern. This is arguably 'unchallenged' because there are very few unfavourable citations. Here are the data in tabular format:

```{r}
d_summary %>%
  filter(case == 'baumeister',
         pubYear %in% c(2015,2017,2018,2019)) %>%
  select(year = pubYear, targetCitations = citesTarget, favourable, equivocal, unfavourable, unclassifiable, excluded = excluded_yes) %>%
  kable(caption = 'BAUMEISTER') %>%
  kable_styling(full_width = F)
```

```{r}
d_summary %>%
  filter(case == 'strack',
         pubYear %in% c(2015,2017,2018,2019)) %>%
  select(year = pubYear, targetCitations = citesTarget, favourable, equivocal, unfavourable, unclassifiable, excluded = excluded_yes) %>%
  kable(caption = 'STRACK') %>%
  kable_styling(full_width = F)
```

```{r}
d_summary %>%
  filter(case == 'sripada',
         pubYear %in% c(2015,2017,2018,2019)) %>%
  select(year = pubYear, targetCitations = citesTarget, favourable, equivocal, unfavourable, unclassifiable, excluded = excluded_yes) %>%
  kable(caption = 'SRIPADA') %>%
  kable_styling(full_width = F)
```

```{r}
d_summary %>%
  filter(case == 'carter',
         pubYear %in% c(2013,2015,2016,2017,2018,2019)) %>%
  select(year = pubYear, targetCitations = citesTarget, favourable, equivocal, unfavourable, unclassifiable, excluded = excluded_yes) %>%
  kable(caption = 'CARTER') %>%
  kable_styling(full_width = F)
```

```{r}
d_summary %>%
  filter(case == 'caruso',
         pubYear %in% c(2013,2015,2016,2017,2018,2019)) %>%
  select(year = pubYear, targetCitations = citesTarget, favourable, equivocal, unfavourable, unclassifiable, excluded = excluded_yes) %>%
  kable(caption = 'CARUSO') %>%
  kable_styling(full_width = F)
```

## Citation balance vs. citation bias

We examined whether articles that cite the target study also cite the contradictory replication (a 'balanced citation effect') or neglect to do so (a 'citation bias effect').

Here we are only looking at articles published in the post-replication period because only they had the opportunity to cite the replication. Recall that this period does not include the year in which the replication was published.

```{r fig.height = 6, fig.width = 8}
ggarrange(citationCurve('strack', areaPlot = 'citesReplication'),
          citationCurve('baumeister', areaPlot = 'citesReplication'),
          nrow = 2, ncol = 1, common.legend = T) %>% 
  annotate_figure(
    left = text_grob("Standardized citation count", rot = 90),
    bottom = text_grob("Publication year"))
```

```{r fig.height = 3, fig.width = 8}
ggarrange(citationCurve('sripada', areaPlot = 'citesReplication'),
          citationCurve('carter', areaPlot = 'citesReplication'),
          citationCurve('caruso', areaPlot = 'citesReplication'),
          nrow = 1, ncol = 3, common.legend = T) %>% 
  annotate_figure(
    left = text_grob("Standardized citation count", rot = 90),
    bottom = text_grob("Publication year"))
```

It may be helpful to zoom in on the relevant part of the timeline too to take a closer look (we'll just do this for Strack and Baumeister):

```{r fig.height = 3, fig.width = 8}
ggarrange(citationCurve('strack', areaPlot = 'citesReplication', zoom = T),
          citationCurve('baumeister', areaPlot = 'citesReplication', zoom = T),
          nrow = 1, ncol = 2, common.legend = T) %>% 
  annotate_figure(
    left = text_grob("Standardized citation count", rot = 90),
    bottom = text_grob("Publication year"))
```

Here's the data in tabular format:

```{r}
d_summary %>%
  filter(case == 'baumeister',
         pubYear %in% c(2017,2018,2019)) %>%
  mutate(replicationCitations_percent = round((citesRep_yes/citesTarget)*100,0)) %>%
  select(year = pubYear, targetCitations = citesTarget, excluded = excluded_yes, replicationCitations = citesRep_yes, replicationCitations_percent) %>%
  kable(caption = 'BAUMEISTER') %>%
  kable_styling(full_width = F)
```

```{r}
d_summary %>%
  filter(case == 'strack',
         pubYear %in% c(2017,2018,2019)) %>%
  mutate(replicationCitations_percent = round((citesRep_yes/citesTarget)*100,0)) %>%
  select(year = pubYear, targetCitations = citesTarget, excluded = excluded_yes, replicationCitations = citesRep_yes, replicationCitations_percent) %>%
  kable(caption = 'STRACK') %>%
  kable_styling(full_width = F)
```

```{r}
d_summary %>%
  filter(case == 'sripada',
         pubYear %in% c(2017,2018,2019)) %>%
  mutate(replicationCitations_percent = round((citesRep_yes/citesTarget)*100,0)) %>%
  select(year = pubYear, targetCitations = citesTarget, excluded = excluded_yes, replicationCitations = citesRep_yes, replicationCitations_percent) %>%
  kable(caption = 'SRIPADA') %>%
  kable_styling(full_width = F)
```

```{r}
d_summary %>%
  filter(case == 'carter',
         pubYear %in% c(2015,2016,2017,2018,2019)) %>%
  mutate(replicationCitations_percent = round((citesRep_yes/citesTarget)*100,0)) %>%
  select(year = pubYear, targetCitations = citesTarget, excluded = excluded_yes, replicationCitations = citesRep_yes, replicationCitations_percent) %>%
  kable(caption = 'CARTER') %>%
  kable_styling(full_width = F)
```

```{r}
d_summary %>%
  filter(case == 'caruso',
         pubYear %in% c(2015,2016,2017,2018,2019)) %>%
  mutate(replicationCitations_percent = round((citesRep_yes/citesTarget)*100,0)) %>%
  select(year = pubYear, targetCitations = citesTarget, excluded = excluded_yes, replicationCitations = citesRep_yes, replicationCitations_percent) %>%
  kable(caption = 'CARUSO') %>%
  kable_styling(full_width = F)
```
And here's the data summarised at the case level with percentages:

```{r}
rbind(
  d_summary %>%
  filter(case %in% c('baumeister','strack','sripada'),
         pubYear %in% c(2017,2018,2019)),
  d_summary %>%
    filter(case %in% c('carter', 'caruso'),
           pubYear %in% c(2015,2016,2017,2018,2019))
) %>%
  group_by(case) %>%
  summarise(targetCitations = sum(citesTarget,na.rm=T), 
            replicationCitations = sum(citesRep_yes,na.rm=T)) %>%
  mutate(`replicationCitations (%)` = round((replicationCitations/targetCitations)*100,0)) %>%
  kable(caption = 'Citations to replication study by articles citing target study in post-replication years') %>%
  kable_styling(full_width = F)
```

There seems to be evidence of citation bias (citing the target article but not the replication) in all cases, though the effect size smaller in the Sripada case (probably because this study is probably only well known because of its connection to the Baumeister study - recall that it was used as a methodological proxy for that more influential study in the replication attempt).

## Principled defense vs. hollow defense

Proponents of the original claim may seek to make the case that the contradictory replication actually does not undermine their position. In this case, one might expect counter-arguments that defend the original claim to be presented in articles that cite the replication study and favourably cite the original study.

Here's the data in tabular format:

```{r}
rbind(
  d_summary %>%
  filter(case %in% c('baumeister','strack','sripada'),
         pubYear %in% c(2017,2018,2019)),
  d_summary %>%
    filter(case %in% c('carter', 'caruso'),
           pubYear %in% c(2015,2016,2017,2018,2019))
) %>%
  group_by(case) %>%
  summarise(counterargs_yes = sum(counterargs_yes,na.rm=T),
            counterargs_no = sum(counterargs_no,na.rm=T)) %>%
  mutate(
    `Replication cited` = counterargs_yes+counterargs_no,
    counterargs_percent = round(
    (counterargs_yes/(`Replication cited`))*100,
    0)) %>%
  select(case,
         `Replication cited`,
         `Counter arguments used` = counterargs_yes,
         -counterargs_no,
         `Counter arguments used (%)` = counterargs_percent) %>%
  kable(caption = 'Use of counter arguments when target study cited favourably and replication study cited') %>%
  kable_styling(full_width = F)
```

The numbers are quite small here because we are looking only at articles that cited the target study and also cited the replication study favourably (note we actually coded for counter arguments whenever the replication study was cited, but this doesn't increase the counts a great deal as is less straightforward to interpret). It looks like there are examples of 'principled defense' (citing the target study favourably and providing counter arguments against the replication) and 'hollow defense' (citing the target study favourably but providing no counter arguments against the replication), with a tendency towards the latter.

### 2020-07-02 analysis/functions.R

@@ -1,0 +1,359 @@
+# functions
+
+# function to standardize citations
+# standardizes the citation counts for the target article and reference class against the year in which the maximum number of citations to the target article were received
+
+standardizeCitations <- function(citationData, referenceData, contentData, replicationYear){
+  
+  # summarize the content data on whether articles cite the replication study
+  repData <- contentData %>% 
+    filter(excluded == F, timePeriod == 'post') %>% 
+    count(pubYear, citesReplication) %>% 
+    pivot_wider(id_cols = 'pubYear', 
+                names_from = citesReplication, 
+                values_from = n, 
+                values_fill = list(n=0)) %>% # use zero for 'missing' values
+    rename("citesRep_no" = `FALSE`, "citesRep_yes" = `TRUE`)
+  
+  # summarize the content data on citation classifications
+  classificationData <- contentData %>% 
+    filter(excluded == F) %>% 
+    count(pubYear, citationClassificationAgreed) %>% 
+    pivot_wider(id_cols = 'pubYear', 
+                names_from = citationClassificationAgreed, 
+                values_from = n, 
+                values_fill = list(n=0)) # use zero for 'missing' values
+  
+  # summarize yearly exclusions for content data
+  exclusionData <- contentData %>%
+    count(pubYear, excluded) %>%
+    pivot_wider(id_cols = 'pubYear',
+                names_from = excluded,
+                values_from = n, 
+                values_fill = list(n=0)) %>%
+    rename("excluded_no" = `FALSE`, "excluded_yes" = `TRUE`)
+  
+  # summarize counter arguments for content data
+  counterargData <- contentData %>% 
+    mutate(counterArguments = factor(counterArguments)) %>% # make a factor so we can retain levels even if count is zero
+    filter(excluded == F, 
+           timePeriod == 'post', 
+           citesReplication == T,
+           citationClassificationAgreed == 'favourable') %>% 
+    count(pubYear, counterArguments, .drop = F) %>% 
+    pivot_wider(id_cols = 'pubYear', 
+                names_from = counterArguments, 
+                values_from = n, 
+                values_fill = list(n=0)) %>% # use zero for 'missing' values
+    rename("counterargs_no" = `FALSE`, "counterargs_yes" = `TRUE`)
+    
+  # merge the summarized content data
[…]

### 2020-07-02 analysis/preprocessing.R

@@ -1,0 +1,79 @@
+# Preprocessing
+
+# identify the five different case studies we are looking at
+caseNames <- c('baumeister', 'sripada', 'strack', 'carter', 'caruso')
+
+## Citation data
+
+### Load files
+d_citations <- data.frame() # create empty list to hold data frames for each case
+for(i in caseNames){ # loop through case names
+  filePath <- here('data','raw', paste0('citations_',i,'.csv')) # create file path
+  d <- read_csv(filePath, col_types = cols(.default = "c")) # load the data
+  d <- d %>% mutate(case = i)
+  d_citations <- bind_rows(d_citations, d) # append to dataframe
+}
+
+### Apply munging
+d_citations <- d_citations %>%
+  mutate(pubYear = as.numeric(PY))
+
+### Save file
+save(d_citations, file = here('data','processed','/d_citations.R'))
+
+## Reference data
+
+### Load files
+d_reference <- data.frame() # create empty list to hold data frames for each case
+for(i in caseNames){ # loop through case names
+  filePath <- here('data','raw', paste0('reference_',i,'.csv')) # create file path
+  d <- read_csv(filePath) # load the data
+  d <- d %>% mutate(case = i)
+  d_reference <- bind_rows(d_reference, d) # append to dataframe
+}
+
+### Apply munging
+
+### Save file
+save(d_reference, file = here('data','processed','/d_reference.R'))
+
+## Content analysis data
+
+### Load files
+d_contentAnalysis <- data.frame() # create empty list to hold data frames for each case
+for(i in caseNames){ # loop through case names
+  filePath <- here('data','raw', paste0('contentAnalysis_',i,'.csv')) # create file path
+  d <- read_csv(filePath, col_types = cols(.default = "c")) # load the data
+  d <- d %>% mutate(case = i)
+  d_contentAnalysis <- bind_rows(d_contentAnalysis, d) # append to dataframe
+}
+
+### Apply munging
+d_contentAnalysis <- d_contentAnalysis %>% 
+  mutate_at(vars( # change the columns below
+    case,
+    firstCoder,
+    secondCoder,
+    exclusionReason,
+    articleType,
+    citationClassificationOriginal,
+    citationClassificationAgreed
+  ),factor) %>% # to the factor type 
+mutate(pubYear = as.numeric(pubYear))
+
[…]

### 2020-07-13 analysis/preliminaryAnalysis.Rmd
syncing with latest version

@@
 ---
-title: "Post-replication citation patterns in psychological science: Preliminary analysis"
+title: "PRCP Preliminary Analysis"
 output: html_document
@@
 
+# Post-replication citation patterns in psychological science: Preliminary analysis
+
 Hi all,
@@
 
-Now we'll plot standardized citation curves and include citations to the set of reference articles (the dashed grey line).
+Now we'll plot standardized citation curves and include citations to the set of reference articles.
 In the graphs below, the annual citation counts for both the target article and reference class are standardized against the calendar year with the highest citation count for the target article (i.e., citations received in that year are set at the standardized value of 100).
@@
 
-Across all cases, `r d_contentAnalysis %>% count(excluded) %>% filter(excluded == T) %>% pull(n)` citing articles had to be excluded because they could not be accessed (n = `r d_contentAnalysis %>% count(exclusionReason) %>% filter(exclusionReason == "No access") %>% pull(n)`), were non-English language (n = `r d_contentAnalysis %>% count(exclusionReason) %>% filter(exclusionReason == "Non-English language") %>% pull(n)`), or because the target article was cited in the references but not in the actual text (n = `r d_contentAnalysis %>% count(exclusionReason) %>% filter(exclusionReason == "Cited in references but not in text") %>% pull(n)`).
+Across all cases, `r d_contentAnalysis %>% count(excluded) %>% filter(excluded == T) %>% pull(n)` citing articles had to be excluded because they could not be accessed (n = `r d_contentAnalysis %>% count(exclusionReason) %>% filter(exclusionReason == "No access") %>% pull(n)`), non-English language (n = `r d_contentAnalysis %>% count(exclusionReason) %>% filter(exclusionReason == "Non-English language") %>% pull(n)`), or because the target article was cited in the references but not in the actual text (n = `r d_contentAnalysis %>% count(exclusionReason) %>% filter(exclusionReason == "Cited in references but not in text") %>% pull(n)`).
 Excluded articles are represented on the graphs below in light grey.
 
-## Claim correction vs. claim perpetuation
+## Claim correction vs. claim perpetuation.
 
@@
             replicationCitations = sum(citesRep_yes,na.rm=T)) %>%
-  mutate(`replicationCitations (%)` = round((replicationCitations/targetCitations)*100,0)) %>%
+  mutate(replicationCitations_percent = round((replicationCitations/targetCitations)*100,0)) %>%
   kable(caption = 'Citations to replication study by articles citing target study in post-replication years') %>%
@@
 
-## Principled defense vs. hollow defense
+## Principled defense vs. hollow defense.
 
@@
 In this case, one might expect counter-arguments that defend the original claim to be presented in articles that cite the replication study and favourably cite the original study.
-
-Here's the data in tabular format:
 
@@
             counterargs_no = sum(counterargs_no,na.rm=T)) %>%
-  mutate(
-    `Replication cited` = counterargs_yes+counterargs_no,
-    counterargs_percent = round(
-    (counterargs_yes/(`Replication cited`))*100,
+  mutate(counterargs_percent = round(
+    (counterargs_yes/(counterargs_yes+counterargs_no))*100,
     0)) %>%
-  select(case,
-         `Replication cited`,
-         `Counter arguments used` = counterargs_yes,
-         -counterargs_no,
-         `Counter arguments used (%)` = counterargs_percent) %>%
-  kable(caption = 'Use of counter arguments when target study cited favourably and replication study cited') %>%
+  kable(caption = 'Use of counter arguments when replication study cited') %>%
   kable_styling(full_width = F)
 ```
-
-The numbers are quite small here because we are looking only at articles that cited the target study and also cited the replication study favourably (note we actually coded for counter arguments whenever the replication study was cited, but this doesn't increase the counts a great deal as is less straightforward to interpret).
-It looks like there are examples of 'principled defense' (citing the target study favourably and providing counter arguments against the replication) and 'hollow defense' (citing the target study favourably but providing no counter arguments against the replication), with a tendency towards the latter.

### 2020-07-20 analysis/preliminaryAnalysis.Rmd
text changes
minor text edits

--- 2020-07-02 analysis/preliminaryAnalysis.Rmd
@@
 The black line represents the raw annual citation count to the target article.
-Note that the citation data was acquired on April 26th, 2019, which partly contributes to the fall in citations in 2019 in all cases.
+Note that the citation data was acquired from the WOS database on April 26th, 2019, which partly contributes to the fall in citations in 2019 in all cases.
 
@@
 
-There seems to be evidence of citation bias (citing the target article but not the replication) in all cases, though the effect size smaller in the Sripada case (probably because this study is probably only well known because of its connection to the Baumeister study - recall that it was used as a methodological proxy for that more influential study in the replication attempt).
+There seems to be evidence of citation bias (citing the target article but not the replication) in all cases, though the effect is smaller in the Sripada case (probably because this study is probably only well known because of its connection to the Baumeister study - recall that it was used as a methodological proxy for that more influential study in the replication attempt).
 

### 2021-02-02 analysis/preprocessing.R
adding coauthor data and relevant processing

@@ -86,4 +86,20 @@ d_contentAnalysis <- d_contentAnalysis %>%
 
 ### Save file
 save(d_contentAnalysis, file = here('data','processed','/d_contentAnalysis.rds'))
+
+## Coauthor data
+
+### Load files
+d_coauthors <- data.frame() # create empty list to hold data frames for each case
+for(i in caseNames){ # loop through case names
+  filePath <- here('data','primary', paste0(i,'Coauthors.csv')) # create file path
+  d <- read_csv(filePath, col_types = cols(.default = "c")) # load the data
+  d <- d %>% mutate(case = i)
+  d_coauthors <- bind_rows(d_coauthors, d) # append to dataframe
+}
+
+### Apply munging
+
+### Save file
+save(d_coauthors, file = here('data','processed','/d_coauthors.rds'))
    

### 2021-02-03 analysis/functions.R
citation curve alpha reduced

@@ -442,7 +442,7 @@ citationCurve <- function(
   if (areaPlot == "classification") {
     basePlot <- basePlot +
       geom_area(
-        alpha = .5,
+        alpha = .4,
         aes(fill = var),
         data = d %>%
           filter(
@@ -458,7 +458,7 @@ citationCurve <- function(
           mutate(var = fct_drop(var))
       ) +
       geom_area(
-        alpha = .5,
+        alpha = .4,
         aes(fill = var),
         data = d %>%
           filter(
@@ -488,7 +488,7 @@ citationCurve <- function(
   if (areaPlot == "citesReplication") {
     basePlot <- basePlot +
       geom_area(
-        alpha =.5,
+        alpha =.4,
         aes(fill = var),
         data = d %>%
           filter(

### 2021-02-04 manuscript/appendix.Rmd
initial commit of manuscript

# Appendix A: Protocol Changes {#appA}

# Appendix B title {#appB}

Tabular data for citation curves and citation valence

```{r results = "asis"}
options(knitr.kable.NA = "-") # set symbol to use for NA cells

d_summary %>%
  filter(case == 'baumeister' & pubYear %in% c(2015,2017,2018,2019) |
         case == 'strack' & pubYear %in% c(2015,2017,2018,2019) |
         case == 'sripada' & pubYear %in% c(2015,2017,2018,2019) |
         case == 'carter' & pubYear %in% c(2013,2015,2016,2017,2018,2019) |
         case == 'caruso' & pubYear %in% c(2013,2015,2016,2017,2018,2019)) %>%
  mutate(favourable = paste0(favourable,' (',round(favourable_prop*100,0),'%)'),
         unfavourable = paste0(unfavourable,' (',round(unfavourable_prop*100,0),'%)'),
         equivocal = paste0(equivocal,' (',round(equivocal_prop*100,0),'%)'),
         unclassifiable = paste0(unclassifiable,' (',round(unclassifiable_prop*100,0),'%)'),
         excluded = paste0(excluded_yes,' (',round(excluded_yes_prop*100,0),'%)'),
         citesRep = paste0(citesRep_yes,' (',round(citesRep_yes_prop*100,0),'%)'),
         citesRep = ifelse(case %in% c('baumeister','sripada','strack') & pubYear == 2015, NA, citesRep), # set citesRep to NA in pre-replication year
         citesRep = ifelse(case %in% c('carter','caruso') & pubYear == 2013, NA, citesRep)) %>% # set citesRep to NA in pre-replication year
  select(year = pubYear, `citing\narticles` = citesTarget, excluded, unclassifiable, favourable, equivocal, unfavourable, citesRep) %>%
  kable(caption = 'Counts (percentages) for citations to the original study, exclusions, citation valence, and concomitant citations to replication study in the time periods subjected to qualitative analysis. Percentages for the Baumeister case are extrapolations based on 40\\% random samples drawn from the pre-replication and post-replication time periods.', booktabs=T, longtable=T) %>%
  pack_rows("Baumeister", 1, 4) %>%
  pack_rows("Sripada", 5, 8) %>%
  pack_rows("Strack", 9, 12) %>%
  pack_rows("Carter", 13, 18) %>%
  pack_rows("Caruso", 19, 24)
```

```{r}
# DATA SUMMARISED FOR POST-REPLICATION PERIOD
d_summary %>%
  filter(year == 'post-replication') %>%
  mutate(favourable = paste0(favourable,' (',round(favourable_prop,2),')'),
         unfavourable = paste0(unfavourable,' (',round(unfavourable_prop,2),')'),
         equivocal = paste0(equivocal,' (',round(equivocal_prop,2),')'),
         unclassifiable = paste0(unclassifiable,' (',round(unclassifiable_prop,2),')'),
         excluded = paste0(excluded_yes,' (',round(excluded_yes_prop,2),')'),
         citesRep = paste0(citesRep_yes,' (',round(citesRep_yes_prop,2),')')) %>%
  select(case, year = pubYear, `citing\narticles` = citesTarget, excluded, unclassifiable, favourable, equivocal, unfavourable, citesRep)
```

### 2021-02-04 manuscript/manuscript.Rmd
initial commit of manuscript

---
title             : "Post-replication citation patterns in psychology: Four case studies"
shorttitle        : "Post-replication citation patterns"

author: 
  - name          : "Tom E. Hardwicke"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Nieuwe Achtergracht 129B, Department of Psychology, University of Amsterdam, 1018 WS Amsterdam, The Netherlands"
    email         : "tom.hardwicke@uva.nl"
  - name          : "Dénes Szűcs"
    affiliation   : "3"
  - name          : "Robert T. Thibault"
    affiliation   : "4,5"
  - name          : "Sophia Crüwell"
    affiliation   : "2"
  - name          : "Olmo R. van den Akker"
    affiliation   : "6"
  - name          : "Michèle B. Nuijten"
    affiliation   : "6"
  - name          : "John P. A. Ioannidis"
    affiliation   : "2,7,8"

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, University of Amsterdam"
  - id            : "2"
    institution   : "Meta-Research Innovation Center Berlin (METRIC-B), QUEST Center for Transforming Biomedical Research, Charité – Universitätsmedizin Berlin"
  - id            : "3"
    institution   : "Department of Psychology, University of Cambridge, UK"
  - id            : "4"
    institution   : "School of Psychological Science, University of Bristol"
  - id            : "5"
    institution   : "MRC Integrative Epidemiology Unit at the University of Bristol"
  - id            : "6"
    institution   : "Department of Methodology and Statistics, Tilburg School of Social and Behavioral Sciences, Tilburg University"
  - id            : "7"
    institution   : "Meta-Research Innovation Center at Stanford (METRICS), Stanford University"
  - id            : "8"
    institution   : "Departments of Medicine, of Health Research and Policy, of Biomedical Data Science, and of Statistics, Stanford University"

author_note: >

abstract: >
   Abstract here.

keywords          : "Replication, self-correction, citations, citation bias, meta-research"

bibliography      : ["references.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes

lang              : "english"
class             : "man"

output            : 
  papaja::apa6_pdf:
    includes:
      after_body: "appendix.tex"
---

```{r load_packages, warning=F}
# load packages
library(knitr) # for literate programming
library(papaja) # for article template
library(tidyverse) # for data munging
library(tidylog) # inline tidyverse feedback
library(here) # for finding files
library(ggpubr) # for arranging plots
library(kableExtra) # for table formatting
```

```{r perform_preprocessing}
# loads raw data, performs preprocessing, saves preprocessed files
source(here('analysis','preprocessing.R'))
```

```{r load_data}
# loads the processed data files output by the preprocessing performed above
load(here('data','processed','d_citations.rds'))
load(here('data','processed','d_reference.rds'))
load(here('data','processed','d_contentAnalysis.rds'))
load(here('data','processed','d_coauthors.rds'))
```

```{r load_functions}
# load custom functions
source(here('analysis','functions.R'))
```

```{r add_standardized_citation_counts}
# identify the five different case studies we are looking at
caseNames <- c('baumeister', 'sripada', 'strack', 'carter', 'caruso')
replicationYears <- c(2016, 2016, 2016, 2014, 2014)

# set up some colour palettes
teal <- "#00798c"
blue <- 'dodgerblue4'
red <-"#CC6677"
yellow <- "#DDCC77"
lightgrey <- "grey75"
darkgrey <- "grey35"
beige <- "#D55E00"

p1 <- c(lightgrey,darkgrey,red,yellow,teal)
p2 <- c(lightgrey,beige,blue)

# summarizes data at the year level for each case and adds standardized citation counts for target citations and reference class citations
d_summary <- data.frame() # create empty list to hold data frames for each case
for(i in seq(1,5)){ # loop through cases
  d <- standardizeCitations(
    thisCase = caseNames[i],
    citationData = d_citations %>% filter(case == caseNames[i]), 
    referenceData = d_reference %>% filter(case == caseNames[i]),
    contentData = d_contentAnalysis %>% filter(case == caseNames[i]),
    replicationYear = replicationYears[i])
  d_summary <- bind_rows(d_summary, d) # append to dataframe
}
```

# Results

```{r}
excluded_n <- d_contentAnalysis %>% filter(excluded == T) %>% nrow() 
excluded_reason <-  d_contentAnalysis %>% filter(excluded == T) %>% count(exclusionReason)
noAccess_n <- excluded_reason %>% filter(exclusionReason == 'No access') %>% pull(n)
nonEnglish_n <- excluded_reason %>% filter(exclusionReason == 'Non-English language') %>% pull(n)
refCiteOnly_n <- excluded_reason %>% filter(exclusionReason == 'Cited in references but not in text') %>% pull(n)
noCite_n <- excluded_reason %>% filter(exclusionReason == 'Does not cite target or replication') %>% pull(n)
```

We excluded `r excluded_n` citing articles from the qualitative analysis because we could not access the full text (n = `r noAccess_n`), they were non-English language (n = `r nonEnglish_n`), they included a citation to the original study in the reference section, but not in the main text (n = `r refCiteOnly_n`), or after manual inspection ßthey did not actually appear to cite the original study at all (n = `r noCite_n`).

## Article characteristics

```{r}
artTyp <- d_contentAnalysis %>% 
  filter(excluded == F) %>% 
  count(articleType) %>%
  mutate(percent = round(n/sum(n),2)*100)

noData <- artTyp %>% filter(articleType == 'No empirical data')
metaAnalysis <- artTyp %>% filter(articleType == 'Data synthesis - meta-analysis')
caseStudy <- artTyp %>% filter(articleType == 'Empirical data - case study')
comment <- artTyp %>% filter(articleType == 'Empirical data - commentary including analysis')
fieldStudy <- artTyp %>% filter(articleType == 'Empirical data - field study')
labStudy <- artTyp %>% filter(articleType == 'Empirical data - laboratory study')
multipleStudy <- artTyp %>% filter(articleType == 'Empirical data - multiple study types are reported')
surveyStudy <- artTyp %>% filter(articleType == 'Empirical data - survey')
```

The research design of citing articles was classified as 'laboratory study' (n = `r labStudy$n`, `r labStudy$percent`%), 'no empirical data' (n = `r noData$n`, `r noData$percent`%), 'survey' (n = `r surveyStudy$n`, `r surveyStudy$percent`%), 'field study' (n = `r fieldStudy$n`, `r fieldStudy$percent`%), 'multiple study types' (n = `r multipleStudy$n`, `r multipleStudy$percent`%), 'meta-analysis' (n = `r metaAnalysis$n`, `r metaAnalysis$percent`%), 'commentary including data analysis' (n = `r comment$n`, `r comment$percent`%), or 'case study' (n = `r caseStudy$n`, `r caseStudy$percent`%).

## Annual citation counts and citation valence

```{r}
# identify how many times valence classifications were changed following discussion between primary and secondary coder
classChange <- d_contentAnalysis %>% filter(excluded == F) %>% mutate(classificationChange = citationClassificationOriginal != citationClassificationAgreed) %>% count(classificationChange) %>% mutate(percent = round(n/sum(n)*100))
```

Figure \@ref(fig:citeCurves) shows standardized annual citation counts for each original study, the respective reference class (citations to all articles published in the same year and same journal as the original study), and classifications of citation valence (favourable, equivocal, unfavourable). The data can also be viewed in tabular format in (LINK TO APPENDIX). Valence classifications by the primary coder were modified after discussion with the secondary coder in `r classChange %>% filter(classificationChange == T) %>% pull(n)` (`r classChange %>% filter(classificationChange == T) %>% pull(percent)`%) cases.

```{r citeCurves, fig.cap='Standardized annual citation counts (solid line) for the five original studies with citation valence (favourable, equivocal, unfavourable, unclassifiable) illustrated by coloured areas in pre-replication and post-replication assessment periods. Dashed line depicts citations to the reference class (all articles published in the same journal and same year as the target article). Annual citation counts are standardized against the year in which the replication was published (citation counts in the replication year, indicated by a black arrow, are set at the standardized value of 100). Citation valence classifications for the Baumeister case are extrapolated to all articles in the assessment period based on a 40% random sample.', fig.width=12, fig.height=12.5, fig.path='figs/'}
ggarrange(
  ggarrange(citationCurve('strack', thisTitle = 'Strack et al. (1988)', standardized = T, plotReference = T, areaPlot = "classification"),
          citationCurve('baumeister', thisTitle = 'Baumeister et al. (1998)', standardized = T, plotReference = T, areaPlot = "classification"),
          nrow = 2, ncol = 1, common.legend = T), 
  ggarrange(citationCurve('sripada', thisTitle = 'Sripada et al. (2014)', standardized = T, plotReference = T, areaPlot = "classification"),
          citationCurve('carter', thisTitle = 'Carter et al. (2011)', standardized = T, plotReference = T, areaPlot = "classification"),
          citationCurve('caruso', thisTitle = 'Caruso et al. (2013)', standardized = T, plotReference = T, areaPlot = "classification"),
          nrow = 1, ncol = 3, common.legend = F),
  nrow = 2, ncol = 1, heights = c(2,1)) %>% 
  annotate_figure(
    left = text_grob("Standardized citation count", rot = 90, size = 20),
    bottom = text_grob("Publication year", size = 20))
```

## Citation balance and citation bias

Figure \@ref(fig:repPlot) shows the proportion of citing articles that also cited or did not cite the replication study after it was published (excluding the publication year itself). In the Strack and Baumeister cases, a considerable majority of articles citing the original study did not cite the replication study, indicating substantial *citation bias*. In the Baumeister case the proportion of articles citing the replication study actually fell from 20% to 16% between 2017-2019. In the Strack case, the proportion increased from 13% to 39%. In the Carter and Caruso cases, the proportion never exceeded 40%, also consistent with substantial *citation bias*. In the Sripada case, it was much more common for the replication study to be cited (>88%) reflecting a *balanced citation* effect.

```{r repPlot, fig.cap='Standardized annual citation counts (solid line) for the five original studies with citation balance/bias (i.e., whether the replication is cited) illustrated by coloured areas in the post-replication assessment period. Dashed line depicts citations to the reference class (all articles published in the same journal and same year as the target article). Annual citation counts are standardized against the year in which the replication was published (citation counts in the replication year, indicated by a black arrow, are set at the standardized value of 100). Replication citation proportions for the Baumeister case are extrapolated to all articles in the assessment period based on a 40% random sample.', fig.width=12, fig.height=12.5, fig.path='figs/'}
ggarrange(
  ggarrange(citationCurve('strack', thisTitle = 'Strack et al. (1988)', areaPlot = 'citesReplication', zoom = T),
          citationCurve('baumeister', thisTitle = 'Baumeister et al. (1998)', areaPlot = 'citesReplication', zoom = T),
          nrow = 2, ncol = 1, common.legend = T), 
  ggarrange(citationCurve('sripada', thisTitle = 'Sripada et al. (2014)', areaPlot = 'citesReplication', zoom = T),
          citationCurve('carter', thisTitle = 'Carter et al. (2011)', areaPlot = 'citesReplication', zoom = T),
          citationCurve('caruso', thisTitle = 'Caruso et al. (2013)', areaPlot = 'citesReplication', zoom = T),
          nrow = 1, ncol = 3, common.legend = F),
  nrow = 2, ncol = 1, heights = c(2,1)) %>% 
  annotate_figure(
    left = text_grob("Standardized citation count", rot = 90, size = 20),
    bottom = text_grob("Publication year", size = 20))
```

## Explicit defence and absent defence

```{r}
ca <- d_contentAnalysis %>% 
  filter(citesReplication == T) %>% 
  group_by(case, counterArguments, .drop = F) %>%
  summarise(n = n()) %>%
  mutate(all = paste0(n,' (',round(n/sum(n),2),')')) %>%
  select(-n)

ca_everycase <- d_contentAnalysis %>% 
  filter(citesReplication == T) %>% 
  group_by(counterArguments, .drop = F) %>%
  summarise(n = n()) %>%
  mutate(case = factor('all cases'),
         all = paste0(n,' (',round(n/sum(n),2),')')) %>%
  select(-n)

ca <- rbind(ca, ca_everycase)

ca_fav <- d_contentAnalysis %>% 
  filter(citesReplication == T) %>% 
  filter(citationClassificationAgreed == 'favourable') %>%
  group_by(case, counterArguments, .drop = F) %>%
  summarise(n = n()) %>%
  mutate(fav = paste0(n,' (',round(n/sum(n),2),')')) %>%
  select(-n)

ca_fav_everycase <- d_contentAnalysis %>% 
  filter(citesReplication == T) %>% 
  filter(citationClassificationAgreed == 'favourable') %>%
  group_by(counterArguments, .drop = F) %>%
  summarise(n = n()) %>%
  mutate(case = factor('all cases'),
         fav = paste0(n,' (',round(n/sum(n),2),')')) %>%
  select(-n)

ca_fav <- rbind(ca_fav, ca_fav_everycase)

left_join(ca,ca_fav) %>%
  pivot_wider(id_cols = case, names_from = counterArguments, values_from = c(all, fav))
```

```{r}
d_contentAnalysis %>% 
  filter(citesReplication == T, counterArguments == T) %>% 
  summarise(evidence = sum(as.logical(evidenceCounter), na.rm=T), methods = sum(as.logical(methodsCounter), na.rm=T), expertise = sum(as.logical(expertiseCounter), na.rm=T))
```

```{r}
# meta-data for names does not appear to be standard. so extract surnames of original authors, automatically match, manually check

d_args <- d_contentAnalysis %>% 
  filter(counterArguments == T)
```
```{r}
# which journals are counter-argument articles published in?
ca_journals <- left_join(d_args, d_citations, by = 'UT') %>% 
  distinct(doi, .keep_all = T) %>% # unique articles only
  count(SO) %>%
  arrange(desc(n)) %>%
  rename('journal' = SO)
```
```{r}
# look at article types for counter-argument articles
ca_artTypes <- d_args %>% 
  distinct(doi, .keep_all = T) %>% # unique articles only
  count(articleType)

ca_noData <- ca_artTypes %>% filter(articleType %in% c('No empirical data')) %>% pull(n)
ca_novData <- ca_artTypes %>% filter(articleType %in% c('Empirical data - field study', 'Empirical data - laboratory study','Empirical data - multiple study types are reported','Empirical data - survey')) %>% pull(n) %>% sum()
ca_existData <- ca_artTypes %>% filter(articleType %in% c('Data synthesis - meta-analysis','Empirical data - commentary including analysis')) %>% pull(n) %>% sum()
```
```{r} 
# get original author surnames for all cases
org_auths_df <- data.frame(case = caseNames, authNames = c('Baumeister, Bratslavsky, Muraven, Tice', # Baumeister, RF; Bratslavsky, E; Muraven, M; Tice, DM,
           "Sripada, Kessler, Jonides", # Sripada, C; Kessler, D; Jonides, J,  
           "Strack, Martin, Stepper", # STRACK, F; MARTIN, LL; STEPPER, S,
           "Carter, Ferguson, Hassin", # Carter, TJ; Ferguson, MJ; Hassin, RR,
           "Caruso, Vohs, Baxter, Waytz")) # Caruso, EM; Vohs, KD; Baxter, B; Waytz, A)

org_auths_list <- str_c(org_auths_df$authNames, collapse = ', ') %>%
  str_replace_all(',','') %>%
  str_split(' ') %>%
  unlist()

# author surnames for counter-argument papers
ca_auth_list <- d_args %>% 
  distinct(doi, .keep_all = T) %>% 
  pull(authors) %>% 
  str_c(collapse = '; ')

# frequency of authors in counter-arg papers
ca_auth_freq <- ca_auth_list %>%
  str_split(';') %>%
  table(dnn = 'name') %>%
  as.data.frame() %>%
  arrange(desc(Freq)) %>%
  rowwise() %>% 
  mutate(surname = str_split(name,',')[[1]][1]) # extract surnames

# frequency of original authors also authoring counter-arg papers
ca_auth_surnames <- ca_auth_freq$surname %>% str_c(collapse = ' ')
org_auths_ca <- org_auths_list[str_detect(ca_auth_surnames, org_auths_list)]

d_coauthors <- d_coauthors %>%
  rowwise() %>%
  mutate(surnames = str_split(Authors,'; ')[[1]] %>% str_extract("[^,]+") %>% str_c(sep = '', collapse = ';'))

co_auths_list <- d_coauthors$surnames %>% str_to_lower() %>% str_c(';', collapse='') %>% str_split(';') %>% as.data.frame(col.names = 'coauthors') %>% filter(coauthors != 'baumeister') %>% distinct(coauthors) %>% count(coauthors) %>% filter(coauthors != '') %>% pull(coauthors)

d_args <- d_args %>%
  distinct(doi, .keep_all = T) %>% 
  rowwise() %>%
  mutate(surnames = str_split(authors,'; ')[[1]] %>% str_extract("[^,]+") %>% str_c(sep = '', collapse = ' '),
         org_auth_detected = ifelse(any(str_detect(str_to_lower(surnames),str_to_lower(paste0("\\b",org_auths_list, "\\b")))),T,F),
         coauth_detected = ifelse(any(str_detect(str_to_lower(surnames),str_to_lower(paste0("\\b",co_auths_list,"\\b")))),T,F))

# after automated matching, we did a manual check to verify
# because the automated matching was based on surnames, sometimes the wrong individual is flagged as a match. Manual checking highlights these cases. And below we override the match classification in these cases by replacing it with zero. Note that the automated matching also cannot distinguish between detecting 1 or more than one match in a given string. So below there are examples of a manual override e.g., changing to 2 when there are 2 relevant individuals in the author list.

d_args <- d_args %>%
  mutate(coauth_detected = case_when(
    authors == "Daley, Ken; Howell, Robert" ~ F,
    authors == "Englert, Chris; Zavery, Alafia; Bertrams, Alex" ~ T, # 2 detected
    authors == "Wang, Yan; Wang, Guosen; Chen, Qiuju; Li, Lin" ~ F,
    authors == "Kuehl, Tim; Bertrams, Alex" ~ T,
    authors == "Friese, Malte; Loschelder, David D.; Gieseler, Karolin; Frankenbach, Julius; Inzlicht, Michael" ~ T,
    authors == "Staller, Mario S.; Mueller, Marcel; Christiansen, Paul; Zaiser, Benjamin; Koerner, Swen; Cole, Jon C." ~ F,
    authors == "Achtziger, Anja; Alos-Ferrer, Carlos; Wagner, Alexander K." ~ T,
    authors == "Dang, Junhua; Liu, Ying; Liu, Xiaoping; Mao, Lihua" ~ F,
    authors == "Wolff, W; Baumann, L; Englert, C" ~ T,
    authors == "Alos-Ferrer, Carlos; Ritschel, Alexander; Garcia-Segarra, Jaume; Achtziger, Anja" ~ T,
    authors == "Stroebe, Wolfgang" ~ T,
    authors == "Schwarz, Norbert; Lee, Spike W. S." ~ T,
    authors == "Nelson, Leif D.; Simmons, Joseph; Simonsohn, Uri" ~ F,
    authors == "Strack, Fritz" ~ F, # this is an original author
    authors == "Cheng, Yongtian; Li, Johnson Ching-Hong; Liu, Xiyao" ~ F,
    authors == "Stroebe, Wolfgang" ~ T,
    authors == "Caruso, Eugene M.; Shapira, Oren; Landy, Justin F." ~ F, # this is an original author
    authors == "Vohs, Kathleen D." ~ F, # this is an original author
    T ~ F # everything else is false
  ))

d_args %>% filter(coauth_detected == T) %>% count(articleType)
d_args %>% filter(org_auth_detected == T) %>% count(articleType)

co_auths_overlap <- d_args %>% count(coauth_detected) %>% filter(coauth_detected == T) %>% pull(n)
org_auths_overlap <- d_args %>% count(org_auth_detected) %>% filter(org_auth_detected == T) %>% pull(n)
```

In additional exploratory analyses (not pre-registered) we examined other characteristics of the 45 unique articles that contained counter-arguments. The articles were published in `r nrow(ca_journals)` individual journals, with *Frontiers in Psychology* publishing `r ca_journals %>% filter(journal == 'FRONTIERS IN PSYCHOLOGY') %>% pull(n)` of the articles, *Social Psychology* publishing `r ca_journals %>% filter(journal == 'SOCIAL PSYCHOLOGY') %>% pull(n)` of the articles and all other journals publishing only 1 or 2 of the articles. `r ca_noData` of the articles did not involve empirical data, `r ca_existData` involved reanalysis or meta-analysis of existing data, and `r ca_novData` involved collection of novel data. The articles had `r nrow(ca_auth_freq)` individual authors. `r org_auths_overlap` articles were (co)authored of one of the original authors and `r co_auths_overlap` articles were (co)authored by a collaborator of one of the first authors of the original articles. `r ca_auth_freq %>% filter(Freq > 1) %>% nrow()` other individuals appeared as an author on 2 of the articles (no individual appeared as an author on more than 2 articles).

```{r}
# how many articles citing the original study also cited the replication?
d_contentAnalysis %>% filter(excluded == F) %>% count(citesReplication, citationClassificationAgreed, counterArguments)
```

# Discussion

# Open practices statement

The study protocol (hypotheses, methods, and analysis plan) was pre-registered on April 7th 2018 (https://osf.io/eh5qd/). An amended protocol was registered on May 1st 2019 (https://osf.io/pdvb5/). All deviations from these protocol or additional exploratory analyses are explicitly acknowledged. All data exclusions and measures conducted during this study are reported. All data, materials, and analysis scripts related to this study are publicly available on The Open Science Framework (https://osf.io/w8h2q/). To facilitate reproducibility this manuscript was written by interleaving regular prose and analysis code (TBA) using knitr [@Xie:2018aa] and papaja [@R-papaja], and is available in a Code Ocean container (TBA) which re-creates the software environment in which the original analyses were performed,

# Funding statement

# Conflict of interest statement

The authors declare no conflicts of interest.

# Author contributions:

\newpage

```{r render_appendix, include=FALSE}
render_appendix("appendix.Rmd")
```

# References

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

### 2021-02-04 .binder/install.R
binder setup

@@ -1,0 +1,10 @@
+install.packages(
+  "ggpubr",
+  "here",
+  "kableExtra",
+  "knitr",
+  "papaja",
+  "rmarkdown",
+  "tidylog",
+  "tidyverse"
+)

### 2021-02-04 .binder/install.R
removing binder files

@@ -1,10 +1,0 @@
-install.packages(
-  "ggpubr",
-  "here",
-  "kableExtra",
-  "knitr",
-  "papaja",
-  "rmarkdown",
-  "tidylog",
-  "tidyverse"
-)

### 2021-02-06 manuscript/manuscript (conflicted).Rmd
initial transfer of methods and results from google doc to r markdown complete

--- 2021-02-04 manuscript/manuscript.Rmd
@@
 
+# Methods
+
+The study protocol (rationale, methods, and analysis plan) was pre-registered on April 7th 2018 (https://osf.io/eh5qd/).
+An amended protocol was registered part way through data collection on May 1st 2019, primarily because we extended the sampling frame to cover additional months (https://osf.io/pdvb5/).
+All deviations from these protocols are explicitly acknowledged in Supplementary Information \@ref(appA).
+All data exclusions and measures conducted during this study are reported.
+
+## Design
+This was a retrospective observational study consisting of four case studies.
+Primary outcome variables were annual citation counts for original studies, citation valence (favourable, equivocal, unfavourable), co-citation of original and replication studies, frequency/type of counter-arguments.
+
+## Sample
+We examined four case studies in which a prominent pre-registered and multi-laboratory replication study strongly contradicted and outweighed the findings of an original study (Table 2).
+
+```{r studyDetails, results = "asis"}
+data_frame(
+  'Original Study' = c('Baumeister et al. (May, 1998)','Sripada et al. (April, 2014)','Strack et al. (May, 1988)','Caruso et al. (July, 2012)','Carter et al. (July, 2011)'),
+  'Replication study' = c('Hagger et al. (July, 2016)**','Hagger et al. (July, 2016)**','Wagenmakers et al. (October, 2016)','Klein et al. (January, 2014)','Klein et al. (January, 2014)'),
+  'Effect' = c('Ego-depletion', 'Ego-depletion','Facial feedback','Money priming','Flag priming'),
+  'Total citations to original*' = c('1974','36','708','57','54'),
+  'Original sample size' = c('k = 1 N = 67', 'k = 1 N = 26', 'k = 1 N = 92', 'k = 1 N = 30','k = 1 N = 70'),
+  'Replication sample size' = c('k = 23 N = 2,141', 'k = 23 N = 2,141', 'k = 17 N = 2,124', 'k = 36 N = 6,333','k = 36 N = 4,896'),
+  'Original effect size [95\\% CI]' = c('d = 2.05 [1.31, 2.79]', 'd = 0.68 [0.09, 1.27]','MD = 0.82 [-0.05, 1.69]','d = 0.8 [0.05. 1.54]', 'd = 0.50 [.01, .99]'),
+  'Replication effect size [95\\% CI]' =c('d = 0.04 [−0.07, 0.15]', 'd = 0.04 [−0.07, 0.15]', 'MD = 0.03 [-0.11, 0.16]', 'd = .01 [-.06, 0.09]', 'd = .01 [-.07, 0.08]')) %>% 
+  kable(booktabs=T, longtable=T) %>% 
+  kable_styling(latex_options="scale_down") %>%
+  column_spec(1, width = "1em") %>%
+  column_spec(2, width = "1em") %>%
+  column_spec(3, width = "1em") %>%
+  column_spec(4, width = "1em") %>%
+  column_spec(5, width = "1em") %>%
+  column_spec(6, width = "1em") %>%
+  column_spec(7, width = "1em") %>%
+  column_spec(8, width = "1em")
+```
+
+```{r}
+# identify how many times valence classifications were changed following discussion between primary and secondary coder
+classChange <- d_contentAnalysis %>% filter(excluded == F) %>% mutate(classificationChange = citationClassificationOriginal != citationClassificationAgreed) %>% count(classificationChange) %>% mutate(percent = round(n/sum(n)*100))
+```
+
+Valence classifications by the primary coder were modified after discussion with the secondary coder in `r classChange %>% filter(classificationChange == T) %>% pull(n)` (`r classChange %>% filter(classificationChange == T) %>% pull(percent)`%) cases.
+
 # Results
@@
 noCite_n <- excluded_reason %>% filter(exclusionReason == 'Does not cite target or replication') %>% pull(n)
+notexcluded_n <- d_contentAnalysis %>% filter(excluded == F) %>% nrow()
 ```
 
-We excluded `r excluded_n` citing articles from the qualitative analysis because we could not access the full text (n = `r noAccess_n`), they were non-English language (n = `r nonEnglish_n`), they included a citation to the original study in the reference section, but not in the main text (n = `r refCiteOnly_n`), or after manual inspection ßthey did not actually appear to cite the original study at all (n = `r noCite_n`).
+In total, `r d_citations %>% nrow()` articles cited one of the original studies of which `r d_contentAnalysis %>% nrow()` articles (after taking a 40% random sample in the Baumeister case) fell within the time period designated for qualitative assessment.
+Of these `r d_contentAnalysis %>% nrow()` articles, we excluded `r excluded_n` from the qualitative analysis because (1) we could not access the full text (n = `r noAccess_n`); (2) they were non-English language (n = `r nonEnglish_n`); (3) they included a citation to the original study in the reference section, but not in the main text (n = `r refCiteOnly_n`); or (4) manual inspection indicated that they did not actually appear to cite the original study at all (n = `r noCite_n`).
+Research design classifications for the remaining `r notexcluded_n` articles included in the qualitative analysis are shown in Table \@ref(tab:articleTypes).
 
@@
 
-```{r}
-artTyp <- d_contentAnalysis %>% 
+```{r articleTypes}
+d_contentAnalysis %>% 
   filter(excluded == F) %>% 
   count(articleType) %>%
-  mutate(percent = round(n/sum(n),2)*100)
-
-noData <- artTyp %>% filter(articleType == 'No empirical data')
-metaAnalysis <- artTyp %>% filter(articleType == 'Data synthesis - meta-analysis')
-caseStudy <- artTyp %>% filter(articleType == 'Empirical data - case study')
-comment <- artTyp %>% filter(articleType == 'Empirical data - commentary including analysis')
-fieldStudy <- artTyp %>% filter(articleType == 'Empirical data - field study')
-labStudy <- artTyp %>% filter(articleType == 'Empirical data - laboratory study')
-multipleStudy <- artTyp %>% filter(articleType == 'Empirical data - multiple study types are reported')
-surveyStudy <- artTyp %>% filter(articleType == 'Empirical data - survey')
+  mutate('Count (%)' = paste0(n,' (',round(n/sum(n),2)*100,')')) %>% 
+  select('Article type' = articleType, 'Count (%)', -n) %>%
+  apa_table(caption = 'Counts and percentages for research design classifications of articles included in qualitative analyses')
 ```
-
-The research design of citing articles was classified as 'laboratory study' (n = `r labStudy$n`, `r labStudy$percent`%), 'no empirical data' (n = `r noData$n`, `r noData$percent`%), 'survey' (n = `r surveyStudy$n`, `r surveyStudy$percent`%), 'field study' (n = `r fieldStudy$n`, `r fieldStudy$percent`%), 'multiple study types' (n = `r multipleStudy$n`, `r multipleStudy$percent`%), 'meta-analysis' (n = `r metaAnalysis$n`, `r metaAnalysis$percent`%), 'commentary including data analysis' (n = `r comment$n`, `r comment$percent`%), or 'case study' (n = `r caseStudy$n`, `r caseStudy$percent`%).
 
@@
 
-```{r}
-# identify how many times valence classifications were changed following discussion between primary and secondary coder
-classChange <- d_contentAnalysis %>% filter(excluded == F) %>% mutate(classificationChange = citationClassificationOriginal != citationClassificationAgreed) %>% count(classificationChange) %>% mutate(percent = round(n/sum(n)*100))
-```
-
-Figure \@ref(fig:citeCurves) shows standardized annual citation counts for each original study, the respective reference class (citations to all articles published in the same year and same journal as the original study), and classifications of citation valence (favourable, equivocal, unfavourable).
-The data can also be viewed in tabular format in (LINK TO APPENDIX).
-Valence classifications by the primary coder were modified after discussion with the secondary coder in `r classChange %>% filter(classificationChange == T) %>% pull(n)` (`r classChange %>% filter(classificationChange == T) %>% pull(percent)`%) cases.
+Figure \@ref(fig:citeCurves) shows standardized annual citation counts for each original study and the respective reference class (citations to all articles published in the same year and same journal as the original study), and classifications of citation valence (favourable, equivocal, unfavourable, unclassifiable or excluded).
+The data can also be viewed in tabular format in Supplementary Table \@ref(tab:tabularData).
+All counts (n) reported in the text and table are raw counts (i.e., not standardized).
 

### 2021-02-06 manuscript/manuscript.Rmd
initial transfer of methods and results from google doc to r markdown complete

@@
 
-lang              : "english"
+lang              : "en-US"
 class             : "man"
@@
 # loads raw data, performs preprocessing, saves preprocessed files
-source(here('analysis','preprocessing.R'))
+source(here("analysis", "preprocessing.R"))
 ```
@@
 # loads the processed data files output by the preprocessing performed above
-load(here('data','processed','d_citations.rds'))
-load(here('data','processed','d_reference.rds'))
-load(here('data','processed','d_contentAnalysis.rds'))
-load(here('data','processed','d_coauthors.rds'))
+load(here("data", "processed", "d_citations.rds"))
+load(here("data", "processed", "d_reference.rds"))
+load(here("data", "processed", "d_contentAnalysis.rds"))
+load(here("data", "processed", "d_coauthors.rds"))
 ```
@@
 # load custom functions
-source(here('analysis','functions.R'))
+source(here("analysis", "functions.R"))
 ```
@@
 # identify the five different case studies we are looking at
-caseNames <- c('baumeister', 'sripada', 'strack', 'carter', 'caruso')
+caseNames <- c("baumeister", "sripada", "strack", "carter", "caruso")
 replicationYears <- c(2016, 2016, 2016, 2014, 2014)
@@
 teal <- "#00798c"
-blue <- 'dodgerblue4'
-red <-"#CC6677"
+blue <- "dodgerblue4"
+red <- "#CC6677"
 yellow <- "#DDCC77"
@@
 
-p1 <- c(lightgrey,darkgrey,red,yellow,teal)
-p2 <- c(lightgrey,beige,blue)
+p1 <- c(lightgrey, darkgrey, red, yellow, teal)
+p2 <- c(lightgrey, beige, blue)
 
@@
 d_summary <- data.frame() # create empty list to hold data frames for each case
-for(i in seq(1,5)){ # loop through cases
+for (i in seq(1, 5)) { # loop through cases
   d <- standardizeCitations(
@@
     contentData = d_contentAnalysis %>% filter(case == caseNames[i]),
-    replicationYear = replicationYears[i])
+    replicationYear = replicationYears[i]
+  )
   d_summary <- bind_rows(d_summary, d) # append to dataframe
@@
 
+```{r patterns}
+tibble(
+  "Progressive responses" = c(1, 2),
+  "Regressive responses" = c(1, 2)
+) %>%
+  apa_table(caption = "Progressive or regressive responses to strongly contradictory replication results and their expected impact on citation patterns for original studies.")
+```
+
+# Methods
+
+The study protocol (rationale, methods, and analysis plan) was pre-registered on April 7th 2018 (https://osf.io/eh5qd/).
+An amended protocol was registered part way through data collection on May 1st 2019, primarily because we extended the sampling frame to cover additional months (https://osf.io/pdvb5/).
+All deviations from these protocols are explicitly acknowledged in Supplementary Information \@ref(appA).
+All data exclusions and measures conducted during this study are reported.
+
+## Design
+This was a retrospective observational study consisting of four case studies.
+Primary outcome variables were annual citation counts for original studies, citation valence (favourable, equivocal, unfavourable), co-citation of original and replication studies, frequency/type of counter-arguments.
+
+## Sample
+We examined four case studies in which a prominent pre-registered and multi-laboratory replication study strongly contradicted and outweighed the findings of an original study (Table \@ref(tab:studyDetails)).
+
+```{r studyDetails}
+tibble(
+  "Original Study" = c("Baumeister et al. (May, 1998)", "Sripada et al. (April, 2014)", "Strack et al. (May, 1988)", "Caruso et al. (July, 2012)", "Carter et al. (July, 2011)"),
+  "Replication study" = c("Hagger et al. (July, 2016)**", "Hagger et al. (July, 2016)**", "Wagenmakers et al. (October, 2016)", "Klein et al. (January, 2014)", "Klein et al. (January, 2014)"),
+  "Effect" = c("Ego-depletion", "Ego-depletion", "Facial feedback", "Money priming", "Flag priming"),
+  "Total citations to original*" = c("1974", "36", "708", "57", "54"),
+  "Original sample size" = c("k = 1 N = 67", "k = 1 N = 26", "k = 1 N = 92", "k = 1 N = 30", "k = 1 N = 70"),
+  "Replication sample size" = c("k = 23 N = 2,141", "k = 23 N = 2,141", "k = 17 N = 2,124", "k = 36 N = 6,333", "k = 36 N = 4,896"),
+  "Original effect size [95% CI]" = c("d = 2.05 [1.31, 2.79]", "d = 0.68 [0.09, 1.27]", "MD = 0.82 [-0.05, 1.69]", "d = 0.8 [0.05. 1.54]", "d = 0.50 [.01, .99]"),
+  "Replication effect size [95% CI]" = c("d = 0.04 [-0.07, 0.15]", "d = 0.04 [-0.07, 0.15]", "MD = 0.03 [-0.11, 0.16]", "d = .01 [-.06, 0.09]", "d = .01 [-.07, 0.08]")
+) %>%
+  kable(booktabs = T, longtable = T, caption = "Sample sizes and effect sizes for replication studies and original studies.
+d = Cohen’s d; MD = mean difference; k = number of data collection sites; N = total number of participants; CI = confidence interval.
+Publication dates are earliest available (i.e., ‘online first’ where relevant).") %>%
+  kable_styling(font_size = 10) %>%
+  column_spec(1, "1.8cm") %>%
+  column_spec(2, "1.8cm") %>%
+  column_spec(3, "1.5cm") %>%
+  column_spec(4, "1.5cm") %>%
+  column_spec(5, "1.5cm") %>%
+  column_spec(6, "1.5cm") %>%
+  column_spec(7, "1.8cm") %>%
+  column_spec(8, "1.8cm") %>%
+  footnote(
+    symbol = c("Total citations to the original study between the publication date and 31st December, 2019.", "For methodological reasons (see Hagger et al., 2016), the ego-depletion replication was aimed at a classic study in the field (Baumeister et al., 1998), but actually employed a modified computer-based version of the original paradigm (Sripada et al., 2014).
+We examined post-replication citation patterns for both studies."), symbol_manual = c("*", "**"),
+    footnote_as_chunk = T, threeparttable = T
+  )
+```
+
+## Procedure
+
+### Annual citation counts
+Citation histories (i.e., bibliographic records for all articles that cite the original study) from the publication date of each original study through to 31st December, 2019, were downloaded from Clarivate Analytics Web of Science Core Collection accessed via the Charité – Universitätsmedizin Berlin on 12th August, 2020.
+We also obtained citation histories for a reference class - all articles published in the same journal and the same year as each original study - from the same source.
+For example, for Baumeister et al.
+(1998) the reference class was all articles published in 1998 in the Journal of Personality and Social Psychology.
+Citation counts were standardized in each case study by setting the citation count in the replication year to the standardized value of ‘100’ and then adjusting the counts in other years according to the same transformation ratio.
+For example, if the raw citation count in the replication year was 1000, citation counts in each year were standardized by dividing by 10.
+This computation was performed separately for the reference class and citations to the original article.
+
+### Qualitative assessment
+Qualitative assessment of citation patterns was limited to a time period starting one year prior to the year of publication of the replication study up until 31st December, 2019, excluding the year in which the replication was published.
+We excluded the replication year because it may be unreasonable to expect citing articles already in the publication pipeline to cite the replication study.
+For the Baumeister case, the qualitative analysis was based on a random sample of 40% of citing articles from the pre-replication period and post-replication period due to the large number of citations to the original study (n = `r d_citations %>% filter(case=='baumeister') %>% nrow()`; see Supplementary Information \@ref(appB) for details).
+
+For each citing article undergoing qualitative assessment, we attempted to retrieve the full text via at least two of the institutional libraries we are affiliated with.
+Inaccessible articles were excluded.
+For articles for which we could obtain the full text, we classified the research design according to the categories in Table \@ref(tab:articleTypes) and recorded whether the replication study was cited after manual inspection of the reference section (see Table \@ref(tab:patterns): citation balance/bias).
+
+```{r}
+# identify how many times valence classifications were changed following discussion between primary and secondary coder
+classChange <- d_contentAnalysis %>%
+  filter(excluded == F) %>%
+  mutate(classificationChange = citationClassificationOriginal != citationClassificationAgreed) %>%
+  count(classificationChange) %>%
+  mutate(percent = round(n / sum(n) * 100))
+```
+
+To examine the belief correction/perpetuation pattern (Table \@ref(tab:patterns)), the primary coder manually extracted the ‘citation context’ of the original study and the replication study (i.e., all relevant verbatim text surrounding each in-text citation).
+The primary coder then classified the citation valence as ‘favourable’, ‘equivocal’, ‘unfavourable’, or ‘unclassifiable’.
+Favourable citations were those used to support a positive claim about the phenomenon of interest whereas unfavourable citations were used to support a negative claim about the phenomenon of interest.
+Citations were considered equivocal if the authors did not take a predominantly favourable or unfavourable position.
+Citations that did not endorse or oppose the phenomenon of interest (for example, simply referring to the procedures of the original study) were designated as ‘unclassifiable’.
+Because this process was inherently subjective, the citation contexts and classifications were also examined by a secondary coder.
+Disagreements were resolved through discussion and a third coder arbitrated when necessary.
+Valence classifications by the primary coder were modified after discussion with the secondary coder in `r classChange %>% filter(classificationChange == T) %>% pull(n)` (`r classChange %>% filter(classificationChange == T) %>% pull(percent)`%) cases.
+
+To examine the explicit/absent defence pattern (Table \@ref(tab:patterns)), the primary coder flagged articles that co-cited the original and replication studies that also contained any explicit defence of the original study.
+Subsequently, two team members (ORA and SC) re-examined all of the flagged cases, extracted verbatim counter-arguments, and developed a post-hoc categorisation scheme that summarised them as concisely and informatively as possible (Table \@ref(tab:caCategories)).
+Coding disagreements were resolved through discussion and a third coder (TEH) arbitrated when necessary.
+
+In additional exploratory (not pre-registered) analyses, we examined overlap of authorship for articles that provided counter-arguments with (1) any of the authors of the original studies; and (2) any prior collaborators of the first authors of the original studies.
+These analyses are complicated by the fact that author names in bibliographic records do not always adhere to the same grammatical standards - for example, whether forenames are initialised or middle names are included - so it is not straightforward to isolate individual authors within bibliographic databases.
+In order to identify prior collaborators of the first authors of the original studies, we downloaded bibliographic records (on 2nd February, 2021) for all papers published by each of the original study first authors, according to their author record in the Web of Science Core Collection.
+These author records are automatically generated by an algorithm that attempts to identify all documents likely published by an individual author using several variations of their name (for example, “Hardwicke, Tom E.”, “Hardwicke, Tom”,  “Hardwicke, T.
+E.”), but errors can still occur and incomplete database coverage means that this method likely misses some of the authors’ prior publications and thus some of their collaborators.
+Nevertheless, the method supports a reasonable lower-bound estimate of authorship overlap with articles providing counter-arguments.
+To identify authorship overlap, we used string manipulation tools in R to extract only author surnames from bibliographic records and then used string matching to automatically detect the presence of original author or collaborator surnames amongst the surnames of authors of articles that provided counter-arguments.
+When a match was detected, it was verified by manual examination of the authors’ full names.
+
 # Results
@@
 ```{r}
-excluded_n <- d_contentAnalysis %>% filter(excluded == T) %>% nrow() 
-excluded_reason <-  d_contentAnalysis %>% filter(excluded == T) %>% count(exclusionReason)
-noAccess_n <- excluded_reason %>% filter(exclusionReason == 'No access') %>% pull(n)
-nonEnglish_n <- excluded_reason %>% filter(exclusionReason == 'Non-English language') %>% pull(n)
-refCiteOnly_n <- excluded_reason %>% filter(exclusionReason == 'Cited in references but not in text') %>% pull(n)
-noCite_n <- excluded_reason %>% filter(exclusionReason == 'Does not cite target or replication') %>% pull(n)
+excluded_n <- d_contentAnalysis %>%
+  filter(excluded == T) %>%
+  nrow()
+excluded_reason <- d_contentAnalysis %>%
+  filter(excluded == T) %>%
+  count(exclusionReason)
+noAccess_n <- excluded_reason %>%
+  filter(exclusionReason == "No access") %>%
+  pull(n)
+nonEnglish_n <- excluded_reason %>%
+  filter(exclusionReason == "Non-English language") %>%
+  pull(n)
+refCiteOnly_n <- excluded_reason %>%
+  filter(exclusionReason == "Cited in references but not in text") %>%
+  pull(n)
+noCite_n <- excluded_reason %>%
+  filter(exclusionReason == "Does not cite target or replication") %>%
+  pull(n)
+notexcluded_n <- d_contentAnalysis %>%
+  filter(excluded == F) %>%
+  nrow()
 ```
 
-We excluded `r excluded_n` citing articles from the qualitative analysis because we could not access the full text (n = `r noAccess_n`), they were non-English language (n = `r nonEnglish_n`), they included a citation to the original study in the reference section, but not in the main text (n = `r refCiteOnly_n`), or after manual inspection ßthey did not actually appear to cite the original study at all (n = `r noCite_n`).
+In total, `r d_citations %>% nrow()` articles cited one of the original studies of which `r d_contentAnalysis %>% nrow()` articles (after taking a 40% random sample in the Baumeister case) fell within the time period designated for qualitative assessment.
+Of these `r d_contentAnalysis %>% nrow()` articles, we excluded `r excluded_n` from the qualitative analysis because (1) we could not access the full text (n = `r noAccess_n`); (2) they were non-English language (n = `r nonEnglish_n`); (3) they included a citation to the original study in the reference section, but not in the main text (n = `r refCiteOnly_n`); or (4) manual inspection indicated that they did not actually appear to cite the original study at all (n = `r noCite_n`).
+Research design classifications for the remaining `r notexcluded_n` articles included in the qualitative analysis are shown in Table \@ref(tab:articleTypes).
 
@@
 
-```{r}
-artTyp <- d_contentAnalysis %>% 
+```{r articleTypes}
+d_contentAnalysis %>%
   filter(excluded == F) %>%
   count(articleType) %>%
-  mutate(percent = round(n/sum(n),2)*100)
-
-noData <- artTyp %>% filter(articleType == 'No empirical data')
-metaAnalysis <- artTyp %>% filter(articleType == 'Data synthesis - meta-analysis')
-caseStudy <- artTyp %>% filter(articleType == 'Empirical data - case study')
-comment <- artTyp %>% filter(articleType == 'Empirical data - commentary including analysis')
-fieldStudy <- artTyp %>% filter(articleType == 'Empirical data - field study')
-labStudy <- artTyp %>% filter(articleType == 'Empirical data - laboratory study')
-multipleStudy <- artTyp %>% filter(articleType == 'Empirical data - multiple study types are reported')
-surveyStudy <- artTyp %>% filter(articleType == 'Empirical data - survey')
+  mutate("Count (%)" = paste0(n, " (", round(n / sum(n), 2) * 100, ")")) %>%
+  select("Article type" = articleType, "Count (%)", -n) %>%
+  apa_table(caption = "Counts and percentages for research design classifications of articles included in qualitative analyses")
 ```
-
-The research design of citing articles was classified as 'laboratory study' (n = `r labStudy$n`, `r labStudy$percent`%), 'no empirical data' (n = `r noData$n`, `r noData$percent`%), 'survey' (n = `r surveyStudy$n`, `r surveyStudy$percent`%), 'field study' (n = `r fieldStudy$n`, `r fieldStudy$percent`%), 'multiple study types' (n = `r multipleStudy$n`, `r multipleStudy$percent`%), 'meta-analysis' (n = `r metaAnalysis$n`, `r metaAnalysis$percent`%), 'commentary including data analysis' (n = `r comment$n`, `r comment$percent`%), or 'case study' (n = `r caseStudy$n`, `r caseStudy$percent`%).
 
@@
 Annual citation counts are standardized against the year in which the replication was published (citation counts in the replication year, indicated by a black arrow, are set at the standardized value of 100).
-Citation valence classifications for the Baumeister case are extrapolated to all articles in the assessment period based on a 40% random sample.', fig.width=12, fig.height=12.5, fig.path='figs/'}
+Citation valence classifications for the Baumeister case are extrapolated to all articles in the assessment period based on a 40% random sample.', fig.width=12, fig.height=12.5, fig.path='figs/', dev=c('png', 'pdf')}
 ggarrange(
-  ggarrange(citationCurve('strack', thisTitle = 'Strack et al. (1988)', standardized = T, plotReference = T, areaPlot = "classification"),
-          citationCurve('baumeister', thisTitle = 'Baumeister et al. (1998)', standardized = T, plotReference = T, areaPlot = "classification"),
-          nrow = 2, ncol = 1, common.legend = T), 
-  ggarrange(citationCurve('sripada', thisTitle = 'Sripada et al. (2014)', standardized = T, plotReference = T, areaPlot = "classification"),
-          citationCurve('carter', thisTitle = 'Carter et al. (2011)', standardized = T, plotReference = T, areaPlot = "classification"),
-          citationCurve('caruso', thisTitle = 'Caruso et al. (2013)', standardized = T, plotReference = T, areaPlot = "classification"),
-          nrow = 1, ncol = 3, common.legend = F),
-  nrow = 2, ncol = 1, heights = c(2,1)) %>% 
+  ggarrange(citationCurve("strack", thisTitle = "Strack et al. (1988)", standardized = T, plotReference = T, areaPlot = "classification"),
+    citationCurve("baumeister", thisTitle = "Baumeister et al. (1998)", standardized = T, plotReference = T, areaPlot = "classification"),
+    nrow = 2, ncol = 1, common.legend = T
+  ),
+  ggarrange(citationCurve("sripada", thisTitle = "Sripada et al. (2014)", standardized = T, plotReference = T, areaPlot = "classification"),
+    citationCurve("carter", thisTitle = "Carter et al. (2011)", standardized = T, plotReference = T, areaPlot = "classification"),
+    citationCurve("caruso", thisTitle = "Caruso et al. (2013)", standardized = T, plotReference = T, areaPlot = "classification"),
+    nrow = 1, ncol = 3, common.legend = F
+  ),
+  nrow = 2, ncol = 1, heights = c(2, 1)
+) %>%
   annotate_figure(
     left = text_grob("Standardized citation count", rot = 90, size = 20),
-    bottom = text_grob("Publication year", size = 20))
+    bottom = text_grob("Publication year", size = 20)
+  )
 ```
+
+```{r}
+# compute maximum difference for 'other' cases (Sripada, Carter, Caruso)
+maxDiff <- max(c(d_summary %>% filter(case == "carter" & pubYear == "2019") %>% pull(citesTarget) - d_summary %>% filter(case == "carter" & pubYear == "2012") %>% pull(citesTarget), d_summary %>% filter(case == "caruso" & pubYear == "2019") %>% pull(citesTarget) - d_summary %>% filter(case == "caruso" & pubYear == "2013") %>% pull(citesTarget), d_summary %>% filter(case == "sripada" & pubYear == "2019") %>% pull(citesTarget) - d_summary %>% filter(case == "sripada" & pubYear == "2015") %>% pull(citesTarget)))
+```
+
+After the replication was published, citations to the reference classes were continuing their trend to plateau (Baumeister case) or increase (other cases).
+By contrast, citations to the original study appeared to undergo a modest decline in the Strack case (decreasing from `r d_summary %>% filter(case == 'strack', pubYear == '2015') %>% pull(citesTarget)` to `r d_summary %>% filter(case == 'strack', pubYear == '2019') %>% pull(citesTarget)` between 2015 and 2019), and a small decline followed by a small increase in the Baumeister case (increasing from `r d_summary %>% filter(case == 'baumeister', pubYear == '2015') %>% pull(citesTarget)` to `r d_summary %>% filter(case == 'baumeister', pubYear == '2019') %>% pull(citesTarget)` between 2015 and 2019).
+In the other cases (Sripada, Carter, Caruso), the total citation counts were much lower and there was considerable variability in the post-replication citation patterns; nevertheless, there was no substantial change in annual citations from pre- to post- replication in these three cases (the maximum difference was +`r maxDiff` citations).
+
+```{r}
+preRepFavRange <- d_summary %>%
+  filter(case == "carter" & pubYear == "2013" | case == "caruso" & pubYear == "2013" | case %in% c("sripada", "baumeister", "strack") & pubYear == "2015") %>%
+  pull(favourable_prop) %>%
+  range()
+
+preRepUnfav_Strack <- d_summary %>%
+  filter(case == "strack", pubYear == "2015") %>%
+  pull(unfavourable_prop)
+postRepUnfav_Strack <- d_summary %>%
+  filter(case == "strack", pubYear == "post-replication") %>%
+  pull(unfavourable_prop)
+preRepFav_Strack <- d_summary %>%
+  filter(case == "strack", pubYear == "2015") %>%
+  pull(favourable_prop)
+postRepFav_Strack <- d_summary %>%
+  filter(case == "strack", pubYear == "post-replication") %>%
+  pull(favourable_prop)
+
+preRepFav_Bau <- d_summary %>%
+  filter(case == "baumeister", pubYear == "2015") %>%
+  pull(favourable_prop)
+postRepFav_Bau <- d_summary %>%
+  filter(case == "baumeister", pubYear == "post-replication") %>%
+  pull(favourable_prop)
+
+unfavN_Bau2017 <- d_summary %>%
+  filter(case == "baumeister", pubYear == "2017") %>%
+  pull(unfavourable)
+unfavProp_Bau2017 <- d_summary %>%
+  filter(case == "baumeister", pubYear == "2017") %>%
+  pull(unfavourable_prop)
+unfavN_Bau2018 <- d_summary %>%
+  filter(case == "baumeister", pubYear == "2018") %>%
+  pull(unfavourable)
+unfavProp_Bau2018 <- d_summary %>%
+  filter(case == "baumeister", pubYear == "2018") %>%
+  pull(unfavourable_prop)
+unfavN_Bau2019 <- d_summary %>%
+  filter(case == "baumeister", pubYear == "2019") %>%
+  pull(unfavourable)
+unfavProp_Bau2019 <- d_summary %>%
+  filter(case == "baumeister", pubYear == "2019") %>%
+  pull(unfavourable_prop)
+```
+Prior to the replication, the vast majority of citations were favourable for all five articles (range `r round(preRepFavRange[1]*100,0)`% to `r round(preRepFavRange[2]*100,0)`%).
+In most cases (Strack, Sripada, Carter, and Caruso) there was a small post-replication increase in unfavourable citations and a small decrease in favourable citations, indicating a modest active correction pattern.
+However, the overall number of unfavourable citations was very low and there was still a substantial majority of favourable citations.
+For example, in the Strack case, unfavourable citations increased from `r preRepUnfav_Strack`% in the pre-replication period (2015) to `r round(postRepUnfav_Strack,2)*100`% in the post-replication period, whilst favourable citations decreased from `r round(preRepFav_Strack,2)*100`% to `r round(postRepFav_Strack,2)*100`%.
+In the Baumeister case, the proportion of favourable citations remained stable from pre-replication (`r round(preRepFav_Bau,2)*100`%) to post-replication (`r round(postRepFav_Bau,2)*100`%), a pattern consistent with belief perpetuation.
+The very small number of unfavourable citations (2017: n = `r unfavN_Bau2017`, `r round(unfavProp_Bau2017,2)*100`%; 2018: n = `r unfavN_Bau2018`, `r round(unfavProp_Bau2018,2)*100`%; 2019: n = `r unfavN_Bau2019`, `r round(unfavProp_Bau2019,2)*100`%) suggests that this is largely an unchallenged belief perpetuation pattern (see Table \@ref(tab:patterns)).
 
@@
 
+```{r}
+citesRep_Bau2017 <- d_summary %>%
+  filter(case == "baumeister", pubYear == "2017") %>%
+  pull(citesRep_yes_prop)
+citesRep_Bau2019 <- d_summary %>%
+  filter(case == "baumeister", pubYear == "2019") %>%
+  pull(citesRep_yes_prop)
+
+citesRep_Strack2017 <- d_summary %>%
+  filter(case == "strack", pubYear == "2017") %>%
+  pull(citesRep_yes_prop)
+citesRep_Strack2019 <- d_summary %>%
+  filter(case == "strack", pubYear == "2019") %>%
+  pull(citesRep_yes_prop)
+
+citesRep_carter_caruso_max <- d_summary %>%
+  filter(case %in% c("carter", "caruso")) %>%
+  pull(citesRep_yes_prop) %>%
+  max(na.rm = T)
+
+citesRep_sripada <- d_summary %>%
+  filter(case == "sripada") %>%
+  pull(citesRep_yes_prop) %>%
+  min(na.rm = T)
+```
+
 Figure \@ref(fig:repPlot) shows the proportion of citing articles that also cited or did not cite the replication study after it was published (excluding the publication year itself).
-In the Strack and Baumeister cases, a considerable majority of articles citing the original study did not cite the replication study, indicating substantial *citation bias*.
-In the Baumeister case the proportion of articles citing the replication study actually fell from 20% to 16% between 2017-2019.
-In the Strack case, the proportion increased from 13% to 39%.
-In the Carter and Caruso cases, the proportion never exceeded 40%, also consistent with substantial *citation bias*.
-In the Sripada case, it was much more common for the replication study to be cited (>88%) reflecting a *balanced citation* effect.
+The data can also be viewed in tabular format in Supplementary Table \@ref(tab:tabularData).
+In the Strack and Baumeister cases, a considerable majority of articles citing the original study did not cite the replication study, indicating substantial citation bias.
+In the Baumeister case the proportion of articles citing the replication study remained stable (`r round(citesRep_Bau2017,2)*100`% in 2017, `r round(citesRep_Bau2019,2)*100`% in 2019).
+In the Strack case, the proportion increased from `r round(citesRep_Strack2017,2)*100`% to `r round(citesRep_Strack2019,2)*100`%.
+In the Carter and Caruso cases, the proportion never exceeded `r round(citesRep_carter_caruso_max,2)*100`%, also consistent with substantial citation bias.
+In the Sripada case, it was much more common for the replication study to be cited (>`r round(citesRep_sripada,2)*100`%) reflecting a balanced citation pattern.
 
@@
 Annual citation counts are standardized against the year in which the replication was published (citation counts in the replication year, indicated by a black arrow, are set at the standardized value of 100).
-Replication citation proportions for the Baumeister case are extrapolated to all articles in the assessment period based on a 40% random sample.', fig.width=12, fig.height=12.5, fig.path='figs/'}
+Replication citation proportions for the Baumeister case are extrapolated to all articles in the assessment period based on a 40% random sample.', fig.width=12, fig.height=12.5, fig.path='figs/', dev=c('png', 'pdf')}
 ggarrange(
-  ggarrange(citationCurve('strack', thisTitle = 'Strack et al. (1988)', areaPlot = 'citesReplication', zoom = T),
-          citationCurve('baumeister', thisTitle = 'Baumeister et al. (1998)', areaPlot = 'citesReplication', zoom = T),
-          nrow = 2, ncol = 1, common.legend = T), 
-  ggarrange(citationCurve('sripada', thisTitle = 'Sripada et al. (2014)', areaPlot = 'citesReplication', zoom = T),
-          citationCurve('carter', thisTitle = 'Carter et al. (2011)', areaPlot = 'citesReplication', zoom = T),
-          citationCurve('caruso', thisTitle = 'Caruso et al. (2013)', areaPlot = 'citesReplication', zoom = T),
-          nrow = 1, ncol = 3, common.legend = F),
-  nrow = 2, ncol = 1, heights = c(2,1)) %>% 
+  ggarrange(citationCurve("strack", thisTitle = "Strack et al. (1988)", areaPlot = "citesReplication", zoom = T),
+    citationCurve("baumeister", thisTitle = "Baumeister et al. (1998)", areaPlot = "citesReplication", zoom = T),
+    nrow = 2, ncol = 1, common.legend = T
+  ),
+  ggarrange(citationCurve("sripada", thisTitle = "Sripada et al. (2014)", areaPlot = "citesReplication", zoom = T),
+    citationCurve("carter", thisTitle = "Carter et al. (2011)", areaPlot = "citesReplication", zoom = T),
+    citationCurve("caruso", thisTitle = "Caruso et al. (2013)", areaPlot = "citesReplication", zoom = T),
+    nrow = 1, ncol = 3, common.legend = F
+  ),
+  nrow = 2, ncol = 1, heights = c(2, 1)
+) %>%
   annotate_figure(
     left = text_grob("Standardized citation count", rot = 90, size = 20),
-    bottom = text_grob("Publication year", size = 20))
+    bottom = text_grob("Publication year", size = 20)
+  )
 ```
@@
 ```{r}
+# prepare table of counter-argument frequency
+
+## get proportion of counter arguments for each individual case
 ca <- d_contentAnalysis %>%
@@
   summarise(n = n()) %>%
-  mutate(all = paste0(n,' (',round(n/sum(n),2),')')) %>%
+  mutate(all = paste0(n, " (", round(n / sum(n), 2), ")")) %>%
   select(-n)
 
+## get proportion of counter arguments across all cases
 ca_everycase <- d_contentAnalysis %>%
@@
   summarise(n = n()) %>%
-  mutate(case = factor('all cases'),
-         all = paste0(n,' (',round(n/sum(n),2),')')) %>%
+  mutate(
+    case = factor("all cases"),
+    all = paste0(n, " (", round(n / sum(n), 2), ")")
+  ) %>%
   select(-n)
@@
 
+## get proportion of counter arguments for each individual case where articles cited original favourably
 ca_fav <- d_contentAnalysis %>%
   filter(citesReplication == T) %>%
-  filter(citationClassificationAgreed == 'favourable') %>%
+  filter(citationClassificationAgreed == "favourable") %>%
   group_by(case, counterArguments, .drop = F) %>%
   summarise(n = n()) %>%
-  mutate(fav = paste0(n,' (',round(n/sum(n),2),')')) %>%
+  mutate(fav = paste0(n, " (", round(n / sum(n), 2), ")")) %>%
   select(-n)
 
+## get proportion of counter arguments across all cases where articles cited original favourably
 ca_fav_everycase <- d_contentAnalysis %>%
   filter(citesReplication == T) %>%
-  filter(citationClassificationAgreed == 'favourable') %>%
+  filter(citationClassificationAgreed == "favourable") %>%
   group_by(counterArguments, .drop = F) %>%
   summarise(n = n()) %>%
-  mutate(case = factor('all cases'),
-         fav = paste0(n,' (',round(n/sum(n),2),')')) %>%
+  mutate(
+    case = factor("all cases"),
+    fav = paste0(n, " (", round(n / sum(n), 2), ")")
+  ) %>%
   select(-n)
@@
 
-left_join(ca,ca_fav) %>%
-  pivot_wider(id_cols = case, names_from = counterArguments, values_from = c(all, fav))
+# for caFreq table (below)
+caFreq <- left_join(ca, ca_fav) %>%
+  pivot_wider(id_cols = case, names_from = counterArguments, values_from = c(all, fav)) %>%
+  rename("No" = all_FALSE, "No " = fav_FALSE, "Yes" = all_TRUE, "Yes " = fav_TRUE)
+```
+
+```{r }
+# for counter-argument categories table (below)
+counterArgCategories <- d_contentAnalysis %>%
+  filter(citesReplication == T, counterArguments == T) %>%
+  summarise(evidence = sum(as.logical(evidenceCounter), na.rm = T), methods = sum(as.logical(methodsCounter), na.rm = T), expertise = sum(as.logical(expertiseCounter), na.rm = T))
+```
+
+Table \@ref(tab:caFreq) shows whether articles that cited the original study and replication study (“co-citing articles”), and the subset of co-citing articles that cited the original study favourably, provided any explicit counter-arguments to defend the credibility of the original finding (an explicit defence) or not (an absent defence).
+Overall, fewer than half of the `r d_contentAnalysis %>% filter(citesReplication==T) %>% nrow()` co-citing articles provided any counter-arguments.
+Of the `r d_contentAnalysis %>% filter(citesReplication==T,citationClassificationAgreed=='favourable') %>% nrow()` co-citing articles that cited the original study favourably, around half provided counter-arguments.
+We identified `r counterArgCategories %>% mutate(totalCA = sum(evidence,methods,expertise)) %>% pull(totalCA)` discrete counter-arguments in `r d_contentAnalysis %>% filter(citesReplication==T,counterArguments==T) %>% nrow()` citing articles (`r d_contentAnalysis %>% filter(citesReplication==T,counterArguments==T) %>% distinct(doi) %>% nrow()` of which were unique articles, as `r d_contentAnalysis %>% filter(citesReplication==T,counterArguments==T) %>% count(doi) %>% filter(n>1) %>% nrow()` of them were cited in two of the case studies) and allocated them to one of three categories (Table \@ref(tab:caCategories)).
+
+```{r caFreq}
+caFreq %>%
+  kable("latex", booktabs = T, caption = "Counts and percentages (in brackets) for whether articles that cited both the original study and replication study provided any explicit argumentation to defend the original study.
+Data are displayed for co-citing articles with any citation valence classification and the subset of co-citing articles with favourable citation valence classifications.") %>%
+  add_header_above(c(" " = 1, "All citation valences" = 2, "Favourable citation valences" = 2))
+```
+
+```{r caCategories}
+counterArgCategories %>%
+  apa_table(caption = "Categorisation of counter-arguments provided to defend the original study in light of the contradictory replication result.
+57 discrete counter-arguments were identified in 50 articles (45 unique articles across cases).")
 ```
@@
 ```{r}
-d_contentAnalysis %>% 
-  filter(citesReplication == T, counterArguments == T) %>% 
-  summarise(evidence = sum(as.logical(evidenceCounter), na.rm=T), methods = sum(as.logical(methodsCounter), na.rm=T), expertise = sum(as.logical(expertiseCounter), na.rm=T))
+# exploratory analyses of articles containing counter-arguments
+
+# get all articles containing counter arguments
+d_args <- d_contentAnalysis %>%
+  filter(counterArguments == T)
+
+# which journals are counter-argument articles published in?
+ca_journals <- left_join(d_args, d_citations, by = "UT") %>%
+  distinct(doi, .keep_all = T) %>% # unique articles only
+  count(SO) %>%
+  arrange(desc(n)) %>%
+  rename("journal" = SO)
 ```
@@
 ```{r}
-# meta-data for names does not appear to be standard. so extract surnames of original authors, automatically match, manually check
-
-d_args <- d_contentAnalysis %>% 
-  filter(counterArguments == T)
-```
-```{r}
-# which journals are counter-argument articles published in?
-ca_journals <- left_join(d_args, d_citations, by = 'UT') %>% 
-  distinct(doi, .keep_all = T) %>% # unique articles only
-  count(SO) %>%
-  arrange(desc(n)) %>%
-  rename('journal' = SO)
-```
-```{r}
-# look at article types for counter-argument articles
+# get article types for counter-argument articles
 ca_artTypes <- d_args %>%
@@
 
-ca_noData <- ca_artTypes %>% filter(articleType %in% c('No empirical data')) %>% pull(n)
-ca_novData <- ca_artTypes %>% filter(articleType %in% c('Empirical data - field study', 'Empirical data - laboratory study','Empirical data - multiple study types are reported','Empirical data - survey')) %>% pull(n) %>% sum()
-ca_existData <- ca_artTypes %>% filter(articleType %in% c('Data synthesis - meta-analysis','Empirical data - commentary including analysis')) %>% pull(n) %>% sum()
+# extract information for in text reporting
+ca_noData <- ca_artTypes %>%
+  filter(articleType %in% c("No empirical data")) %>%
+  pull(n)
+ca_novData <- ca_artTypes %>%
+  filter(articleType %in% c("Empirical data - field study", "Empirical data - laboratory study", "Empirical data - multiple study types are reported", "Empirical data - survey")) %>%
+  pull(n) %>%
+  sum()
+ca_existData <- ca_artTypes %>%
+  filter(articleType %in% c("Data synthesis - meta-analysis", "Empirical data - commentary including analysis")) %>%
+  pull(n) %>%
+  sum()
 ```
+
 ```{r} 
+# now look at authorship overlap - a problem here is identifying unique individuals due to non-standard meta-data for author names.
+To address this we will convert full names to surnames then manually verify any automatically identified matches.
+
 # get original author surnames for all cases
-org_auths_df <- data.frame(case = caseNames, authNames = c('Baumeister, Bratslavsky, Muraven, Tice', # Baumeister, RF; Bratslavsky, E; Muraven, M; Tice, DM,
+org_auths_df <- data.frame(case = caseNames, authNames = c(
+  "Baumeister, Bratslavsky, Muraven, Tice", # Baumeister, RF; Bratslavsky, E; Muraven, M; Tice, DM,
   "Sripada, Kessler, Jonides", # Sripada, C; Kessler, D; Jonides, J,
@@
   "Carter, Ferguson, Hassin", # Carter, TJ; Ferguson, MJ; Hassin, RR,
-           "Caruso, Vohs, Baxter, Waytz")) # Caruso, EM; Vohs, KD; Baxter, B; Waytz, A)
+  "Caruso, Vohs, Baxter, Waytz"
+)) # Caruso, EM; Vohs, KD; Baxter, B; Waytz, A)
 
-org_auths_list <- str_c(org_auths_df$authNames, collapse = ', ') %>%
-  str_replace_all(',','') %>%
-  str_split(' ') %>%
+# convert to a list of original author surnames
+org_auths_list <- str_c(org_auths_df$authNames, collapse = ", ") %>%
+  str_replace_all(",", "") %>%
+  str_split(" ") %>%
   unlist()
@@
 
+# now extract surnames of co-authors of original first authors
 d_coauthors <- d_coauthors %>%
   rowwise() %>%
-  mutate(surnames = str_split(Authors,'; ')[[1]] %>% str_extract("[^,]+") %>% str_c(sep = '', collapse = ';'))
+  mutate(surnames = str_split(Authors, "; ")[[1]] %>% str_extract("[^,]+") %>% str_c(sep = "", collapse = ";"))
 
-co_auths_list <- d_coauthors$surnames %>% str_to_lower() %>% str_c(';', collapse='') %>% str_split(';') %>% as.data.frame(col.names = 'coauthors') %>% filter(coauthors != 'baumeister') %>% distinct(coauthors) %>% count(coauthors) %>% filter(coauthors != '') %>% pull(coauthors)
+# and convert to a list
+co_auths_list <- d_coauthors$surnames %>%
+  str_to_lower() %>%
+  str_c(";", collapse = "") %>%
+  str_split(";") %>%
+  as.data.frame(col.names = "coauthors") %>%
+  filter(coauthors != "baumeister") %>%
+  distinct(coauthors) %>%
+  count(coauthors) %>%
+  filter(coauthors != "") %>%
+  pull(coauthors)
+
+# now we will do some string matching to identify overlap of authorship between counter-argument article (CAA) authors and (1) original authors and (2) collaborators of original first authors.
 
@@
   rowwise() %>%
-  mutate(surnames = str_split(authors,'; ')[[1]] %>% str_extract("[^,]+") %>% str_c(sep = '', collapse = ' '),
-         org_auth_detected = ifelse(any(str_detect(str_to_lower(surnames),str_to_lower(paste0("\\b",org_auths_list, "\\b")))),T,F),
-         coauth_detected = ifelse(any(str_detect(str_to_lower(surnames),str_to_lower(paste0("\\b",co_auths_list,"\\b")))),T,F))
+  mutate(
+    surnames = str_split(authors, "; ")[[1]] %>% str_extract("[^,]+") %>% str_c(sep = "", collapse = " "), # extract CAA author surnames
+    org_auth_detected = ifelse(any(str_detect(str_to_lower(surnames), str_to_lower(paste0("\\b", org_auths_list, "\\b")))), T, F), # check if each CAA author surname is amongst original author surnames
+    coauth_detected = ifelse(any(str_detect(str_to_lower(surnames), str_to_lower(paste0("\\b", co_auths_list, "\\b")))), T, F)
+  ) # check if each CAA author surname is amongst original first author collaborator surnames
 
@@
 Manual checking highlights these cases.
-And below we override the match classification in these cases by replacing it with zero.
-Note that the automated matching also cannot distinguish between detecting 1 or more than one match in a given string.
-So below there are examples of a manual override e.g., changing to 2 when there are 2 relevant individuals in the author list.
+If necessary, we override the match classification of the automated matching.
 
@@
 
-d_args %>% filter(coauth_detected == T) %>% count(articleType)
-d_args %>% filter(org_auth_detected == T) %>% count(articleType)
+# extract information for in text reporting
+co_auths_overlap <- d_args %>%
+  count(coauth_detected) %>%
+  filter(coauth_detected == T) %>%
+  pull(n)
 
-co_auths_overlap <- d_args %>% count(coauth_detected) %>% filter(coauth_detected == T) %>% pull(n)
-org_auths_overlap <- d_args %>% count(org_auth_detected) %>% filter(org_auth_detected == T) %>% pull(n)
+org_auths_overlap <- d_args %>%
+  count(org_auth_detected) %>%
+  filter(org_auth_detected == T) %>%
+  pull(n)
+
+overlapArtTypes <- d_args %>%
+  filter(coauth_detected == T | org_auth_detected == T) %>%
+  count(articleType)
 ```
 
-In additional exploratory analyses (not pre-registered) we examined other characteristics of the 45 unique articles that contained counter-arguments.
-The articles were published in `r nrow(ca_journals)` individual journals, with *Frontiers in Psychology* publishing `r ca_journals %>% filter(journal == 'FRONTIERS IN PSYCHOLOGY') %>% pull(n)` of the articles, *Social Psychology* publishing `r ca_journals %>% filter(journal == 'SOCIAL PSYCHOLOGY') %>% pull(n)` of the articles and all other journals publishing only 1 or 2 of the articles.
+In additional exploratory analyses (not pre-registered) we examined other characteristics of the ` d_contentAnalysis %>% filter(citesReplication==T,counterArguments==T) %>% distinct(doi) %>% nrow()` unique articles that contained counter-arguments.
+The articles were published in `r nrow(ca_journals)` individual journals, with *Frontiers in Psychology* publishing `r ca_journals %>% filter(journal == 'FRONTIERS IN PSYCHOLOGY') %>% pull(n)` of the articles, *Social Psychology* publishing `r ca_journals %>% filter(journal == 'SOCIAL PSYCHOLOGY') %>% pull(n)` of the articles, and all other journals publishing only 1 or 2 of the articles.
 `r ca_noData` of the articles did not involve empirical data, `r ca_existData` involved reanalysis or meta-analysis of existing data, and `r ca_novData` involved collection of novel data.
-The articles had `r nrow(ca_auth_freq)` individual authors.
-`r org_auths_overlap` articles were (co)authored of one of the original authors and `r co_auths_overlap` articles were (co)authored by a collaborator of one of the first authors of the original articles.
-`r ca_auth_freq %>% filter(Freq > 1) %>% nrow()` other individuals appeared as an author on 2 of the articles (no individual appeared as an author on more than 2 articles).
-
-```{r}
-# how many articles citing the original study also cited the replication?
-d_contentAnalysis %>% filter(excluded == F) %>% count(citesReplication, citationClassificationAgreed, counterArguments)
-```
+The articles had `r nrow(ca_auth_freq)` individual authors of whom all contributed to a single article except for `r ca_auth_freq %>% filter(Freq > 1) %>% nrow()` individuals who had (co)authored 2 articles.
+`r org_auths_overlap` articles were (co)authored by one of the original authors and `r co_auths_overlap` articles were (co)authored by at least one prior collaborator of one of the first authors of the original articles.
+`r overlapArtTypes %>% filter(articleType == "No empirical data") %>% pull(n)` of these articles did not involve empirical data and `r overlapArtTypes %>% filter(articleType != "No empirical data") %>% pull(n) %>% sum()` of them involved novel data collection.
 
@@
 \setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
+\setlength{\leftskip}{0.5in

### 2021-02-06 manuscript/manuscript.tex
initial transfer of methods and results from google doc to r markdown complete

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  american,
  ,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Post-replication citation patterns in psychology: Four case studies},
  pdflang={en-US},
  pdfkeywords={Replication, self-correction, citations, citation bias, meta-research},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{Post-replication citation patterns}
\author{Tom E. Hardwicke\textsuperscript{1,2}, Dénes Szűcs\textsuperscript{3}, Robert T. Thibault\textsuperscript{4,5}, Sophia Crüwell\textsuperscript{2}, Olmo R. van den Akker\textsuperscript{6}, Michèle B. Nuijten\textsuperscript{6}, \& John P. A. Ioannidis\textsuperscript{2,7,8}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Psychology, University of Amsterdam\\\textsuperscript{2} Meta-Research Innovation Center Berlin (METRIC-B), QUEST Center for Transforming Biomedical Research, Charité – Universitätsmedizin Berlin\\\textsuperscript{3} Department of Psychology, University of Cambridge, UK\\\textsuperscript{4} School of Psychological Science, University of Bristol\\\textsuperscript{5} MRC Integrative Epidemiology Unit at the University of Bristol\\\textsuperscript{6} Department of Methodology and Statistics, Tilburg School of Social and Behavioral Sciences, Tilburg University\\\textsuperscript{7} Meta-Research Innovation Center at Stanford (METRICS), Stanford University\\\textsuperscript{8} Departments of Medicine, of Health Research and Policy, of Biomedical Data Science, and of Statistics, Stanford University}
\authornote{

Correspondence concerning this article should be addressed to Tom E. Hardwicke, Nieuwe Achtergracht 129B, Department of Psychology, University of Amsterdam, 1018 WS Amsterdam, The Netherlands. E-mail: tom.hardwicke@uva.nl}
\keywords{Replication, self-correction, citations, citation bias, meta-research}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[variant=american]{english}
\else
  \usepackage[shorthands=off,main=american]{babel}
\fi

\title{Post-replication citation patterns in psychology: Four case studies}

\date{}

\abstract{
Abstract here.
}

\begin{document}
\maketitle

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:patterns}Progressive or regressive responses to strongly contradictory replication results and their expected impact on citation patterns for original studies.}

\begin{tabular}{ll}
\toprule
Progressive responses & \multicolumn{1}{c}{Regressive responses}\\
\midrule
1.00 & 1.00\\
2.00 & 2.00\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

\hypertarget{methods}{%
\section{Methods}\label{methods}}

The study protocol (rationale, methods, and analysis plan) was pre-registered on April 7th 2018 (\url{https://osf.io/eh5qd/}). An amended protocol was registered part way through data collection on May 1st 2019, primarily because we extended the sampling frame to cover additional months (\url{https://osf.io/pdvb5/}). All deviations from these protocols are explicitly acknowledged in Supplementary Information \ref{appA}. All data exclusions and measures conducted during this study are reported.

\hypertarget{design}{%
\subsection{Design}\label{design}}

This was a retrospective observational study consisting of four case studies. Primary outcome variables were annual citation counts for original studies, citation valence (favourable, equivocal, unfavourable), co-citation of original and replication studies, frequency/type of counter-arguments.

\hypertarget{sample}{%
\subsection{Sample}\label{sample}}

We examined four case studies in which a prominent pre-registered and multi-laboratory replication study strongly contradicted and outweighed the findings of an original study (Table \ref{tab:studyDetails}).

\begingroup\fontsize{10}{12}\selectfont

\begin{ThreePartTable}
\begin{TableNotes}[para]
\item[*] Total citations to the original study between the publication date and 31st December, 2019.
\item[**] For methodological reasons (see Hagger et al., 2016), the ego-depletion replication was aimed at a classic study in the field (Baumeister et al., 1998), but actually employed a modified computer-based version of the original paradigm (Sripada et al., 2014). We examined post-replication citation patterns for both studies.
\end{TableNotes}
\begin{longtable}[t]{>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.5cm}>{\raggedright\arraybackslash}p{1.5cm}>{\raggedright\arraybackslash}p{1.5cm}>{\raggedright\arraybackslash}p{1.5cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}}
\caption{\label{tab:studyDetails}Sample sizes and effect sizes for replication studies and original studies. d = Cohen’s d; MD = mean difference; k = number of data collection sites; N = total number of participants; CI = confidence interval. Publication dates are earliest available (i.e., ‘online first’ where relevant).}\\
\toprule
Original Study & Replication study & Effect & Total citations to original* & Original sample size & Replication sample size & Original effect size [95\% CI] & Replication effect size [95\% CI]\\
\midrule
Baumeister et al. (May, 1998) & Hagger et al. (July, 2016)** & Ego-depletion & 1974 & k = 1 N = 67 & k = 23 N = 2,141 & d = 2.05 [1.31, 2.79] & d = 0.04 [-0.07, 0.15]\\
Sripada et al. (April, 2014) & Hagger et al. (July, 2016)** & Ego-depletion & 36 & k = 1 N = 26 & k = 23 N = 2,141 & d = 0.68 [0.09, 1.27] & d = 0.04 [-0.07, 0.15]\\
Strack et al. (May, 1988) & Wagenmakers et al. (October, 2016) & Facial feedback & 708 & k = 1 N = 92 & k = 17 N = 2,124 & MD = 0.82 [-0.05, 1.69] & MD = 0.03 [-0.11, 0.16]\\
Caruso et al. (July, 2012) & Klein et al. (January, 2014) & Money priming & 57 & k = 1 N = 30 & k = 36 N = 6,333 & d = 0.8 [0.05. 1.54] & d = .01 [-.06, 0.09]\\
Carter et al. (July, 2011) & Klein et al. (January, 2014) & Flag priming & 54 & k = 1 N = 70 & k = 36 N = 4,896 & d = 0.50 [.01, .99] & d = .01 [-.07, 0.08]\\
\bottomrule
\insertTableNotes
\end{longtable}
\end{ThreePartTable}
\endgroup{}

\hypertarget{procedure}{%
\subsection{Procedure}\label{procedure}}

\hypertarget{annual-citation-counts}{%
\subsubsection{Annual citation counts}\label{annual-citation-counts}}

Citation histories (i.e., bibliographic records for all articles that cite the original study) from the publication date of each original study through to 31st December, 2019, were downloaded from Clarivate Analytics Web of Science Core Collection accessed via the Charité -- Universitätsmedizin Berlin on 12th August, 2020. We also obtained citation histories for a reference class - all articles published in the same journal and the same year as each original study - from the same source. For example, for Baumeister et al.~(1998) the reference class was all articles published in 1998 in the Journal of Personality and Social Psychology. Citation counts were standardized in each case study by setting the citation count in the replication year to the standardized value of \enquote{100} and then adjusting the counts in other years according to the same transformation ratio. For example, if the raw citation count in the replication year was 1000, citation counts in each year were standardized by dividing by 10. This computation was performed separately for the reference class and citations to the original article.

\hypertarget{qualitative-assessment}{%
\subsubsection{Qualitative assessment}\label{qualitative-assessment}}

Qualitative assessment of citation patterns was limited to a time period starting one year prior to the year of publication of the replication study up until 31st December, 2019, excluding the year in which the replication was published. We excluded the replication year because it may be unreasonable to expect citing articles already in the publication pipeline to cite the replication study. For the Baumeister case, the qualitative analysis was based on a random sample of 40\% of citing articles from the pre-replication period and post-replication period due to the large number of citations to the original study (n = 1974; see Supplementary Information \ref{appB} for details).

For each citing article undergoing qualitative assessment, we attempted to retrieve the full text via at least two of the institutional libraries we are affiliated with. Inaccessible articles were excluded. For articles for which we could obtain the full text, we classified the research design according to the categories in Table \ref{tab:articleTypes} and recorded whether the replication study was cited after manual inspection of the reference section (see Table \ref{tab:patterns}: citation balance/bias).

To examine the belief correction/perpetuation pattern (Table \ref{tab:patterns}), the primary coder manually extracted the \enquote{citation context} of the original study and the replication study (i.e., all relevant verbatim text surrounding each in-text citation). The primary coder then classified the citation valence as \enquote{favourable}, \enquote{equivocal}, \enquote{unfavourable}, or \enquote{unclassifiable}. Favourable citations were those used to support a positive claim about the phenomenon of interest whereas unfavourable citations were used to support a negative claim about the phenomenon of interest. Citations were considered equivocal if the authors did not take a predominantly favourable or unfavourable position. Citations that did not endorse or oppose the phenomenon of interest (for example, simply referring to the procedures of the original study) were designated as \enquote{unclassifiable}. Because this process was inherently subjective, the citation contexts and classifications were also examined by a secondary coder. Disagreements were resolved through discussion and a third coder arbitrated when necessary. Valence classifications by the primary coder were modified after discussion with the secondary coder in 30 (5\%) cases.

To examine the explicit/absent defence pattern (Table \ref{tab:patterns}), the primary coder flagged articles that co-cited the original and replication studies that also contained any explicit defence of the original study. Subsequently, two team members (ORA and SC) re-examined all of the flagged cases, extracted verbatim counter-arguments, and developed a post-hoc categorisation scheme that summarised them as concisely and informatively as possible (Table \ref{tab:caCategories}). Coding disagreements were resolved through discussion and a third coder (TEH) arbitrated when necessary.

In additional exploratory (not pre-registered) analyses, we examined overlap of authorship for articles that provided counter-arguments with (1) any of the authors of the original studies; and (2) any prior collaborators of the first authors of the original studies. These analyses are complicated by the fact that author names in bibliographic records do not always adhere to the same grammatical standards - for example, whether forenames are initialised or middle names are included - so it is not straightforward to isolate individual authors within bibliographic databases. In order to identify prior collaborators of the first authors of the original studies, we downloaded bibliographic records (on 2nd February, 2021) for all papers published by each of the original study first authors, according to their author record in the Web of Science Core Collection. These author records are automatically generated by an algorithm that attempts to identify all documents likely published by an individual author using several variations of their name (for example, \enquote{Hardwicke, Tom E.}, \enquote{Hardwicke, Tom}, \enquote{Hardwicke, T. E.}), but errors can still occur and incomplete database coverage means that this method likely misses some of the authors' prior publications and thus some of their collaborators. Nevertheless, the method supports a reasonable lower-bound estimate of authorship overlap with articles providing counter-arguments. To identify authorship overlap, we used string manipulation tools in R to extract only author surnames from bibliographic records and then used string matching to automatically detect the presence of original author or collaborator surnames amongst the surnames of authors of articles that provided counter-arguments. When a match was detected, it was verified by manual examination of the authors' full names.

\hypertarget{results}{%
\section{Results}\label{results}}

In total, 2829 articles cited one of the original studies of which 632 articles (after taking a 40\% random sample in the Baumeister case) fell within the time period designated for qualitative assessment. Of these 632 articles, we excluded 69 from the qualitative analysis because (1) we could not access the full text (n = 58); (2) they were non-English language (n = 6); (3) they included a citation to the original study in the reference section, but not in the main text (n = 4); or (4) manual inspection indicated that they did not actually appear to cite the original study at all (n = 1). Research design classifications for the remaining 563 articles included in the qualitative analysis are shown in Table \ref{tab:articleTypes}.

\hypertarget{article-characteristics}{%
\subsection{Article characteristics}\label{article-characteristics}}

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:articleTypes}Counts and percentages for research design classifications of articles included in qualitative analyses}

\begin{tabular}{ll}
\toprule
Article type & \multicolumn{1}{c}{Count (\%)}\\
\midrule
Data synthesis - meta-analysis & 11 (2)\\
No empirical data & 163 (29)\\
Empirical data - case study & 1 (0)\\
Empirical data - commentary including analysis & 4 (1)\\
Empirical data - field study & 39 (7)\\
Empirical data - laboratory study & 245 (44)\\
Empirical data - multiple study types are reported & 23 (4)\\
Empirical data - survey & 77 (14)\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

\hypertarget{annual-citation-counts-and-citation-valence}{%
\subsection{Annual citation counts and citation valence}\label{annual-citation-counts-and-citation-valence}}

Figure \ref{fig:citeCurves} shows standardized annual citation counts for each original study and the respective reference class (citations to all articles published in the same year and same journal as the original study), and classifications of citation valence (favourable, equivocal, unfavourable, unclassifiable or excluded). The data can also be viewed in tabular format in Supplementary Table \ref{tab:tabularData}. All counts (n) reported in the text and table are raw counts (i.e., not standardized).

\begin{figure}
\centering
\includegraphics{figs/citeCurves-1.png}
\caption{\label{fig:citeCurves}Standardized annual citation counts (solid line) for the five original studies with citation valence (favourable, equivocal, unfavourable, unclassifiable) illustrated by coloured areas in pre-replication and post-replication assessment periods. Dashed line depicts citations to the reference class (all articles published in the same journal and same year as the target article). Annual citation counts are standardized against the year in which the replication was published (citation counts in the replication year, indicated by a black arrow, are set at the standardized value of 100). Citation valence classifications for the Baumeister case are extrapolated to all articles in the assessment period based on a 40\% random sample.}
\end{figure}

After the replication was published, citations to the reference classes were continuing their trend to plateau (Baumeister case) or increase (other cases). By contrast, citations to the original study appeared to undergo a modest decline in the Strack case (decreasing from 56 to 41 between 2015 and 2019), and a small decline followed by a small increase in the Baumeister case (increasing from 191 to 199 between 2015 and 2019). In the other cases (Sripada, Carter, Caruso), the total citation counts were much lower and there was considerable variability in the post-replication citation patterns; nevertheless, there was no substantial change in annual citations from pre- to post- replication in these three cases (the maximum difference was +8 citations).

Prior to the replication, the vast majority of citations were favourable for all five articles (range 67\% to 100\%). In most cases (Strack, Sripada, Carter, and Caruso) there was a small post-replication increase in unfavourable citations and a small decrease in favourable citations, indicating a modest active correction pattern. However, the overall number of unfavourable citations was very low and there was still a substantial majority of favourable citations. For example, in the Strack case, unfavourable citations increased from 0\% in the pre-replication period (2015) to 6\% in the post-replication period, whilst favourable citations decreased from 82\% to 71\%. In the Baumeister case, the proportion of favourable citations remained stable from pre-replication (71\%) to post-replication (73\%), a pattern consistent with belief perpetuation. The very small number of unfavourable citations (2017: n = 7, 7.00\%; 2018: n = 2, 2\%; 2019: n = 2, 4\%) suggests that this is largely an unchallenged belief perpetuation pattern (see Table \ref{tab:patterns}).

\hypertarget{citation-balance-and-citation-bias}{%
\subsection{Citation balance and citation bias}\label{citation-balance-and-citation-bias}}

Figure \ref{fig:repPlot} shows the proportion of citing articles that also cited or did not cite the replication study after it was published (excluding the publication year itself). The data can also be viewed in tabular format in Supplementary Table \ref{tab:tabularData}. In the Strack and Baumeister cases, a considerable majority of articles citing the original study did not cite the replication study, indicating substantial citation bias. In the Baumeister case the proportion of articles citing the replication study remained stable (20\% in 2017, 16\% in 2019). In the Strack case, the proportion increased from 13\% to 39\%. In the Carter and Caruso cases, the proportion never exceeded 40\%, also consistent with substantial citation bias. In the Sripada case, it was much more common for the replication study to be cited (\textgreater88\%) reflecting a balanced citation pattern.

\begin{figure}
\centering
\includegraphics{figs/repPlot-1.png}
\caption{\label{fig:repPlot}Standardized annual citation counts (solid line) for the five original studies with citation balance/bias (i.e., whether the replication is cited) illustrated by coloured areas in the post-replication assessment period. Dashed line depicts citations to the reference class (all articles published in the same journal and same year as the target article). Annual citation counts are standardized against the year in which the replication was published (citation counts in the replication year, indicated by a black arrow, are set at the standardized value of 100). Replication citation proportions for the Baumeister case are extrapolated to all articles in the assessment period based on a 40\% random sample.}
\end{figure}

\hypertarget{explicit-defence-and-absent-defence}{%
\subsection{Explicit defence and absent defence}\label{explicit-defence-and-absent-defence}}

Table \ref{tab:caFreq} shows whether articles that cited the original study and replication study (\enquote{co-citing articles}), and the subset of co-citing articles that cited the original study favourably, provided any explicit counter-arguments to defend the credibility of the original finding (an explicit defence) or not (an absent defence). Overall, fewer than half of the 121 co-citing articles provided any counter-arguments. Of the 59 co-citing articles that cited the original study favourably, around half provided counter-arguments. We identified 57 discrete counter-arguments in 50 citing articles (45 of which were unique articles, as 5 of them were cited in two of the case studies) and allocated them to one of three categories (Table \ref{tab:caCategories}).

\begin{table}

\caption{\label{tab:caFreq}Counts and percentages (in brackets) for whether articles that cited both the original study and replication study provided any explicit argumentation to defend the original study. Data are displayed for co-citing articles with any citation valence classification and the subset of co-citing articles with favourable citation valence classifications.}
\centering
\begin{tabular}[t]{lllll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{All citation valences} & \multicolumn{2}{c}{Favourable citation valences} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
case & No & Yes & No  & Yes \\
\midrule
baumeister & 23 (0.55) & 19 (0.45) & 11 (0.46) & 13 (0.54)\\
carter & 10 (0.77) & 3 (0.23) & 3 (0.75) & 1 (0.25)\\
caruso & 6 (0.6) & 4 (0.4) & 1 (0.5) & 1 (0.5)\\
sripada & 9 (0.5) & 9 (0.5) & 2 (0.4) & 3 (0.6)\\
strack & 23 (0.61) & 15 (0.39) & 12 (0.5) & 12 (0.5)\\
\addlinespace
all cases & 71 (0.59) & 50 (0.41) & 29 (0.49) & 30 (0.51)\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:caCategories}Categorisation of counter-arguments provided to defend the original study in light of the contradictory replication result. 57 discrete counter-arguments were identified in 50 articles (45 unique articles across cases).}

\begin{tabular}{lll}
\toprule
evidence & \multicolumn{1}{c}{methods} & \multicolumn{1}{c}{expertise}\\
\midrule
11 & 44 & 2\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

In additional exploratory analyses (not pre-registered) we examined other characteristics of the \texttt{d\_contentAnalysis\ \%\textgreater{}\%\ filter(citesReplication==T,counterArguments==T)\ \%\textgreater{}\%\ distinct(doi)\ \%\textgreater{}\%\ nrow()} unique articles that contained counter-arguments. The articles were published in 34 individual journals, with \emph{Frontiers in Psychology} publishing 7 of the articles, \emph{Social Psychology} publishing 4 of the articles, and all other journals publishing only 1 or 2 of the articles. 17 of the articles did not involve empirical data, 3 involved reanalysis or meta-analysis of existing data, and 25 involved collection of novel data. The articles had 112 individual authors of whom all contributed to a single article except for 9 individuals who had (co)authored 2 articles. 3 articles were (co)authored by one of the original authors and 9 articles were (co)authored by at least one prior collaborator of one of the first authors of the original articles. 7 of these articles did not involve empirical data and 5 of them involved novel data collection.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{open-practices-statement}{%
\section{Open practices statement}\label{open-practices-statement}}

The study protocol (hypotheses, methods, and analysis plan) was pre-registered on April 7th 2018 (\url{https://osf.io/eh5qd/}). An amended protocol was registered on May 1st 2019 (\url{https://osf.io/pdvb5/}). All deviations from these protocol or additional exploratory analyses are explicitly acknowledged. All data exclusions and measures conducted during this study are reported. All data, materials, and analysis scripts related to this study are publicly available on The Open Science Framework (\url{https://osf.io/w8h2q/}). To facilitate reproducibility this manuscript was written by interleaving regular prose and analysis code (TBA) using knitr ({\textbf{???}}) and papaja ({\textbf{???}}), and is available in a Code Ocean container (TBA) which re-creates the software environment in which the original analyses were performed,

\hypertarget{funding-statement}{%
\section{Funding statement}\label{funding-statement}}

\hypertarget{conflict-of-interest-statement}{%
\section{Conflict of interest statement}\label{conflict-of-interest-statement}}

The authors declare no conflicts of interest.

\hypertarget{author-contributions}{%
\section{Author contributions:}\label{author-contributions}}

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\setlength{\parindent}{-0.5in}

\textbackslash setlength\{\leftskip\}\{0.5in

\clearpage
\makeatletter
\efloat@restorefloats
\makeatother

\begin{appendix}
\hypertarget{appA}{%
\section{Appendix A: Protocol Changes}\label{appA}}

\hypertarget{appB}{%
\section{Appendix B title}\label{appB}}

Tabular data for citation curves and citation valence

\begin{longtable}[t]{lrllllll}
\caption{\label{tab:tabularData}Counts (percentages) for citations to the original study, exclusions, citation valence, and concomitant citations to replication study in the time periods subjected to qualitative analysis. Percentages for the Baumeister case are extrapolations based on 40\% random samples drawn from the pre-replication and post-replication time periods.}\\
\toprule
year & citing
articles & excluded & unclassifiable & favourable & equivocal & unfavourable & citesRep\\
\midrule
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{Baumeister}}\\
\hspace{1em}2015 & 191 & 8 (11\%) & 1 (1\%) & 53 (71\%) & 11 (15\%) & 2 (3\%) & -\\
\hspace{1em}2017 & 211 & 13 (13\%) & 4 (4\%) & 71 (70\%) & 7 (7\%) & 7 (7\%) & 20 (20\%)\\
\hspace{1em}2018 & 183 & 7 (8\%) & 7 (8\%) & 69 (77\%) & 5 (6\%) & 2 (2\%) & 15 (17\%)\\
\hspace{1em}2019 & 199 & 2 (4\%) & 3 (7\%) & 33 (73\%) & 5 (11\%) & 2 (4\%) & 7 (16\%)\\
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{Sripada}}\\
\hspace{1em}2015 & 6 & 1 (17\%) & 1 (17\%) & 4 (67\%) & 0 (0\%) & 0 (0\%) & -\\
\hspace{1em}2017 & 3 & 0 (0\%) & 1 (33\%) & 1 (33\%) & 1 (33\%) & 0 (0\%) & 3 (100\%)\\
\hspace{1em}2018 & 8 & 0 (0\%) & 0 (0\%) & 1 (12\%) & 3 (38\%) & 4 (50\%) & 7 (88\%)\\
\hspace{1em}2019 & 9 & 1 (11\%) & 0 (0\%) & 4 (44\%) & 2 (22\%) & 2 (22\%) & 8 (89\%)\\
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{Strack}}\\
\hspace{1em}2015 & 56 & 6 (11\%) & 4 (7\%) & 46 (82\%) & 0 (0\%) & 0 (0\%) & -\\
\hspace{1em}2017 & 55 & 7 (13\%) & 1 (2\%) & 45 (82\%) & 1 (2\%) & 1 (2\%) & 7 (13\%)\\
\hspace{1em}2018 & 47 & 7 (15\%) & 3 (6\%) & 31 (66\%) & 2 (4\%) & 4 (9\%) & 15 (32\%)\\
\hspace{1em}2019 & 41 & 7 (17\%) & 0 (0\%) & 26 (63\%) & 4 (10\%) & 4 (10\%) & 16 (39\%)\\
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{Carter}}\\
\hspace{1em}2013 & 6 & 0 (0\%) & 1 (17\%) & 5 (83\%) & 0 (0\%) & 0 (0\%) & -\\
\hspace{1em}2015 & 5 & 1 (20\%) & 1 (20\%) & 2 (40\%) & 1 (20\%) & 0 (0\%) & 2 (40\%)\\
\hspace{1em}2016 & 13 & 2 (15\%) & 2 (15\%) & 7 (54\%) & 0 (0\%) & 2 (15\%) & 5 (38\%)\\
\hspace{1em}2017 & 10 & 1 (10\%) & 0 (0\%) & 6 (60\%) & 1 (10\%) & 2 (20\%) & 4 (40\%)\\
\hspace{1em}2018 & 6 & 2 (33\%) & 2 (33\%) & 2 (33\%) & 0 (0\%) & 0 (0\%) & 2 (33\%)\\
\hspace{1em}2019 & 3 & 0 (0\%) & 0 (0\%) & 3 (100\%) & 0 (0\%) & 0 (0\%) & 0 (0\%)\\
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{Caruso}}\\
\hspace{1em}2013 & 1 & 0 (0\%) & 0 (0\%) & 1 (100\%) & 0 (0\%) & 0 (0\%) & -\\
\hspace{1em}2015 & 7 & 0 (0\%) & 0 (0\%) & 5 (71\%) & 1 (14\%) & 1 (14\%) & 2 (29\%)\\
\hspace{1em}2016 & 17 & 2 (12\%) & 3 (18\%) & 8 (47\%) & 2 (12\%) & 2 (12\%) & 2 (12\%)\\
\hspace{1em}2017 & 10 & 2 (20\%) & 0 (0\%) & 8 (80\%) & 0 (0\%) & 0 (0\%) & 2 (20\%)\\
\hspace{1em}2018 & 7 & 0 (0\%) & 1 (14\%) & 5 (71\%) & 1 (14\%) & 0 (0\%) & 1 (14\%)\\
\hspace{1em}2019 & 9 & 0 (0\%) & 0 (0\%) & 5 (56\%) & 1 (11\%) & 3 (33\%) & 3 (33\%)\\
\bottomrule
\end{longtable}

\begin{verbatim}
##         case             year citing\narticles  excluded unclassifiable
## 1 baumeister post-replication              593 22 (0.09)      14 (0.06)
## 2    sripada post-replication               20  1 (0.05)       1 (0.05)
## 3     strack post-replication              143 21 (0.15)       4 (0.03)
## 4     carter post-replication               37  6 (0.16)       5 (0.14)
## 5     caruso post-replication               50  4 (0.08)       4 (0.08)
##   favourable equivocal unfavourable  citesRep
## 1 173 (0.73) 17 (0.07)    11 (0.05) 42 (0.18)
## 2    6 (0.3)   6 (0.3)      6 (0.3)  18 (0.9)
## 3 102 (0.71)  7 (0.05)     9 (0.06) 38 (0.27)
## 4  20 (0.54)  2 (0.05)     4 (0.11) 13 (0.35)
## 5  31 (0.62)   5 (0.1)     6 (0.12)  10 (0.2)
\end{verbatim}
\end{appendix}

\end{document}

### 2021-02-06 manuscript/manuscript.Rmd
added discussion section

@@
 An amended protocol was registered part way through data collection on May 1st 2019, primarily because we extended the sampling frame to cover additional months (https://osf.io/pdvb5/).
-All deviations from these protocols are explicitly acknowledged in Supplementary Information \@ref(appA).
+All deviations from these protocols are explicitly acknowledged in Appendix \@ref(appA).
 All data exclusions and measures conducted during this study are reported.
@@
 We excluded the replication year because it may be unreasonable to expect citing articles already in the publication pipeline to cite the replication study.
-For the Baumeister case, the qualitative analysis was based on a random sample of 40% of citing articles from the pre-replication period and post-replication period due to the large number of citations to the original study (n = `r d_citations %>% filter(case=='baumeister') %>% nrow()`; see Supplementary Information \@ref(appB) for details).
+For the Baumeister case, the qualitative analysis was based on a random sample of 40% of citing articles from the pre-replication period and post-replication period due to the large number of citations to the original study (n = `r d_citations %>% filter(case=='baumeister') %>% nrow()`; see Appendix \@ref(appB) for details).
 
@@
 Figure \@ref(fig:citeCurves) shows standardized annual citation counts for each original study and the respective reference class (citations to all articles published in the same year and same journal as the original study), and classifications of citation valence (favourable, equivocal, unfavourable, unclassifiable or excluded).
-The data can also be viewed in tabular format in Supplementary Table \@ref(tab:tabularData).
+The data can also be viewed in tabular format in Appendix \@ref(tab:tabularData).
 All counts (n) reported in the text and table are raw counts (i.e., not standardized).
@@
 Figure \@ref(fig:repPlot) shows the proportion of citing articles that also cited or did not cite the replication study after it was published (excluding the publication year itself).
-The data can also be viewed in tabular format in Supplementary Table \@ref(tab:tabularData).
+The data can also be viewed in tabular format in Appendix \@ref(tab:tabularData).
 In the Strack and Baumeister cases, a considerable majority of articles citing the original study did not cite the replication study, indicating substantial citation bias.
@@
 
-In additional exploratory analyses (not pre-registered) we examined other characteristics of the ` d_contentAnalysis %>% filter(citesReplication==T,counterArguments==T) %>% distinct(doi) %>% nrow()` unique articles that contained counter-arguments.
+In additional exploratory analyses (not pre-registered) we examined other characteristics of the `r d_contentAnalysis %>% filter(citesReplication==T,counterArguments==T) %>% distinct(doi) %>% nrow()` unique articles that contained counter-arguments.
 The articles were published in `r nrow(ca_journals)` individual journals, with *Frontiers in Psychology* publishing `r ca_journals %>% filter(journal == 'FRONTIERS IN PSYCHOLOGY') %>% pull(n)` of the articles, *Social Psychology* publishing `r ca_journals %>% filter(journal == 'SOCIAL PSYCHOLOGY') %>% pull(n)` of the articles, and all other journals publishing only 1 or 2 of the articles.
@@
 
-# Discussion
+Discussion
+It has been proposed that replication studies can facilitate scientific self-correction by modifying scientists’ belief in the credibility of published findings (Ioannidis, 2012; Vazire & Holcombe, 2020; Zwaan et al., 2018); however, the extent to which this occurs in practice is unclear.
+In this study, we investigated how the research community responded to four strongly contradictory replications in the field of psychology by examining post-replication citation patterns for original studies.
+We observed some ‘progressive’ response in the form of modest active correction patterns (a small decline in favourable citations and a small increase in unfavourable citations), but also even more prominent ‘regressive’ responses in the form of unchallenged belief perpetuation (sustained levels of favourable citations and few unfavourable citations, particularly in the Baumeister case) and considerable citation bias (neglecting to cite the replication study; in all cases aside from Sripada).
+When authors cited the original study favourably despite the replication result, only half of the articles provided any explicit counter arguments in defence of the original study (an explicit defence).
+Overall, these findings are consistent with prior observations that favourable citation patterns can continue relatively unperturbed by subsequent publication of contradictory results in studies with lower risk of bias (Tatsioni et al., 2007) or even by full retraction (Budd et al., 1998), and that positive (supportive) evidence is preferentially cited relative to negative (non-supportive) evidence once a theory gets entrenched despite overwhelming evidence against it (Greenberg, 2009).
 
-# Open practices statement
+It is reasonable to question whether the replication results in these case studies should (rationally) have instigated belief change in the research community and thus triggered a more sizable decline in favourable citations than we observed.
+It is important to note that the case studies we selected were deliberately chosen because the replication results were superior to the original results in terms of both credibility and evidential value.
+The replications were pre-registered, muli-laboratory studies with large sample sizes.
+By contrast, the original studies had much smaller sample sizes, and arguably had a much higher risk of bias, as they were not pre-registered, were performed by single teams, and arose in domains affected by publication bias and other questionable research practices (E.
+C.
+Carter et al., 2015; Coles et al., 2019; Vadillo, 2019; Vadillo et al., 2016).
+Thus, it seems reasonable to suggest that the compelling replication results should have reduced belief in the credibility of the original results.
 
-The study protocol (hypotheses, methods, and analysis plan) was pre-registered on April 7th 2018 (https://osf.io/eh5qd/).
-An amended protocol was registered on May 1st 2019 (https://osf.io/pdvb5/).
-All deviations from these protocol or additional exploratory analyses are explicitly acknowledged.
-All data exclusions and measures conducted during this study are reported.
-All data, materials, and analysis scripts related to this study are publicly available on The Open Science Framework (https://osf.io/w8h2q/).
-To facilitate reproducibility this manuscript was written by interleaving regular prose and analysis code (TBA) using knitr [@Xie:2018aa] and papaja [@R-papaja], and is available in a Code Ocean container (TBA) which re-creates the software environment in which the original analyses were performed,
+The clear evidence of citation bias that our study documents may have two main contributory factors: (1) a lack of awareness about the replication results; and/or (2) a decision to ignore the replication results.
+The practical issue of awareness is not necessarily straightforward to address.
+Individual scientists can find it difficult to keep up-to-date with the voluminous literature that is relevant to their research.
+Recently, the reference manager Zotero has introduced a feature that alerts users when an article in their database has been retracted (Zotero, 2019).
+One could imagine a similar feature being introduced for replication studies, perhaps based on databases that explicitly identify replication studies such as Curate Science (https://curatescience.org/).
+However, it is much less straightforward to define a relevant replication study (Neuliep & Crandall, 1993) and users would need to be alerted that this requires some scientific judgement rather than simple article meta-data.
+Another solution could be to encourage researchers to focus less on individual studies and more on up-to-date evidence summaries (i.e., reviews and meta-analyses) in which relevant evidence is systematically identified and collated.
+This would require that high-quality and contemporary evidence summaries are available; however, in psychology systematic reviews and meta-analyses can be of low quality and their results may still be inflated and irreproducible (Kvarven et al., 2020; Maassen et al., 2020; Polanin et al., 2020).
+Moreover, most empirical studies are not included in any form of evidence synthesis (Hardwicke et al., 2020).
+
+The second issue of authors ignoring highly relevant replications is undesirable and implies a biased appraisal of the evidence.
+Contradictory replication results do not necessarily nullify the credibility of an original study (Collins, 1985; Earp & Trafimow, 2015; Maxwell et al., 2015), but we would still expect highly relevant replication results to be cited and explicitly debated.
+In fact, we found that when authors continued to favourably cite original studies despite the replication results, around half did not provide explicit counter-arguments in defence of the original study.
+Although our study did not examine researchers’ individual beliefs, one recent study found that when research psychologists were confronted with replication evidence, they often did update their (self-reported) beliefs, albeit modestly (McDiarmid et al., 2021).
+However, there are several reasons to be uncertain about whether the results from this artificial setting might generalize to real-world settings, including potential observer bias (i.e., participants behaving differently because they are under observation and/or responding to the perceived expectations of the research team) and the possibility that individuals may behave differently in settings where they have substantial personal investment and may be publicly scrutinized.
+Additionally, cognitive psychology studies have obtained some evidence of a ‘continued influence effect’ wherein an individual's beliefs and behaviour can continue to be influenced by false or misleading information despite subsequent efforts to reject it (Lewandowsky et al., 2012).
+It is plausible that various cognitive biases, such as confirmation bias (preferentially seeking out and processing evidence that supports pre-existing beliefs) or motivated reasoning (constructing and evaluating arguments based on what is desirable rather than what is rationally justifiable), may partly explain researchers' tendency to ignore or dismiss the replication evidence (Bishop, 2019; Kunda, 1990; Nickerson, 1998).
+
+We observed that when counter-arguments were raised, most of them tried to dismiss the contradictory replication by claiming that the original and the replication studies differed in important ways that moderated the absence/presence of the effect under investigation.
+In some cases, authors pointed to evidence from other studies as a rationale for their continued belief in the effect.
+In a minority of cases, the authors challenged the competence of replicators.
+We found that articles presenting counter-arguments were published in a variety of journals (rather than clustered in a few journals), and involved collection of new data in around half of the cases.
+They were also published by a sizeable group of investigators, only a minority of whom were one of the original authors or had previously collaborated with one of the original first authors.
+This suggests that the explicit defence of the original study came from a variety of sources rather than being confined to a small number of investigators.
+However, note that this analysis may underestimate authorship overlap due to difficulties isolating individual researcher identities (see methods section) and articles published in the same year as the replication were not included.
+
+The findings presented here are inherently limited by the observational nature of the study design, which complicates straightforward conclusions about the causal impact of the replications.
+Although the use of a reference class enables us to detect the influence of exogenous factors to some extent, we cannot rule out their contribution.
+For example, the modest decline in favourable citations observed in most cases could be attributable, at least in part, to a more general awareness in the research community about methodological issues (for example, that the sample sizes of the original studies may not have provided sufficient statistical power).
+We have also focused only on the replication study and the original study in each case study, without considering the impact of other potentially relevant events.
+Notably, meta-research studies have detected signatures of publication bias and other questionable research practices in the fields to which these case studies belong (E.
+C.
+Carter et al., 2015; Coles et al., 2019; Vadillo, 2019; Vadillo et al., 2016) and other relevant replication studies contesting prior findings have been published (e.g., Rohrer et al., 2015).
+
+We have only been able to gauge reactions to replications to the extent that they are reflected in citation patterns.
+Thus, for example, our method would not detect if there had been a self-correction effect amongst individuals who would not typically cite the original study, such as students, members of the public, or researchers working in other fields.
+A contested study may continue to be  cited favourably by its proponents who remain working in the field.
+This will suffice to create belief perpetuation in the published literature, even though other scientists may simply no longer be interested in getting involved with such a strongly contested research topic.
+Relatedly, we did not examine citation patterns beyond 3-5 years post-replication.
+Some perspectives envision scientific self-correction unfolding over a much longer time scale (Lauden, 1981; Peterson & Panofsky, 2020) and suggest the process is characterised less by the impact of individual study results and more by the gradual accumulation of converging evidence, gradual revision of theoretical understanding, and/or informal sociological processes (e.g., researchers choosing alternative topics to study).
+Thus, whilst the current findings may contradict the expectations of a more direct and expedited view of the corrective impact of replication studies (Ioannidis, 2012; Vazire & Holcombe, 2020), they are not necessarily inconsistent with a slower and more indirect process of self-correction.
+
+Generalization beyond these four case studies requires caution.
+Of note, the replication studies examined here were some of the first large-scale multi-laboratory replication attempts conducted in the field of psychology.
+This gave them particular prominence and initiated considerable debate, resulting in broader ramifications beyond the research community that typically studies the topics under scrutiny (Nelson et al., 2018).
+Also of note, we deliberately selected case studies where the replication studies were high-profile and had yielded high credibility evidence that strongly contradicted and outweighed the original findings.
+A self-correction effect may be less expected in cases where replication results are more ambiguous, less consequential, or less well-known.
+For example, in a situation where two high credibility studies with similar evidential value yield contradictory results, it would be premature to lose confidence in one of the studies before further investigation has probed the cause of the discrepancy.
+Pursuit of potential moderating factors may be entirely rational in such circumstances (Gershman, 2019).
+
+We should also highlight that particular aspects of our study were inherently subjective, specifically, the identification of citation context, the classification of citation valence, and the identification, extraction, and categorisation of counter-arguments.
+In order to minimize subjectivity, a team of six investigators performed all coding in duplicate, with a third investigator arbitrating if necessary.
+As disagreements between primary and secondary coders were infrequent, we are confident that the classifications are meaningful, but there may be some edge cases when an independent observer might reasonably disagree with our classifications.
+
+In conclusion, post-replication citation patterns in four case studies indicated that the anticipated corrective impact of strongly contradictory replication results did not materialise to any substantive degree.
+A lack of awareness of replications and/or a decision to discount them appears to have played a significant role.
+This highlights potential practical problems with the discoverability of replication studies and psychological or sociological issues related to belief change.
+The findings also indicate that scientific self-correction may not be as expedient or straightforward as one might hope (Ioannidis, 2012), adding further impetus towards efforts to improve the quality of the academic literature (Hardwicke et al., 2020; Nelson et al., 2018).
+
+# Materials, data, and analysis code availability
+All data, materials, and analysis scripts related to this study are publicly available on the Open Science Framework (https://osf.io/w8h2q/).
+To facilitate reproducibility this manuscript was written by interleaving regular prose and analysis code using knitr (Xie, 2017) and papaja (Aust & Barth, 2020), and is available in a Code Ocean container (TO BE ADDED) which re-creates the software environment in which the original analyses were performed.
 
 # Funding statement
+The authors received no specific funding for this work.
+The Meta-Research Innovation Center at Stanford (METRICS) is supported by a grant from the Laura and John Arnold Foundation.
+The Meta-Research Innovation Center Berlin (METRIC-B) is supported by a grant from the Einstein Foundation and Stiftung Charité.
+The work of John Ioannidis is supported by an unrestricted grant from Sue and Bob O’Donnell.
+Robert T.
+Thibault is supported by a postdoctoral fellowship from the Fonds de la recherche en santé du Québec.
+Tom E.
+Hardwicke receives funding from the European Union’s Horizon 2020 research and innovation programme under the Marie Skłodowska-Curie grant agreement No.
+841188.
+Denes Szucs receives funding from the James S.
+McDonnell Foundation 21st Century Science Initiative in Understanding Human Cognition (grant number 220020370).
+Olmo R.
+van den Akker is supported by a Consolidator Grant (IMPROVE) from the European Research Council (ERC; grant no.
+726361).
 
 # Conflict of interest statement
-
 The authors declare no conflicts of interest.
 
-# Author contributions:
+# Author contributions
+TEH, DS, and JPAI designed the study.
+TEH, DS, RTT, SC, ORA, and MN performed the data extraction and coding.
+TEH and SC performed the data analysis.
+TEH wrote the manuscript.
+All authors provided feedback and approved the manuscript.
+
+# References
+
+Aust, F., & Barth, M. (2020). Papaja: Create apa manuscripts with rmarkdown. https://github.com/crsh/papaja
+
+Bastiaansen, J.
+A., de Vries, Y.
+A., & Munafò, M.
+R.
+(2015).
+Citation distortions in the literature on the serotonin-transporter-linked polymorphic region and amygdala activation.
+Biological Psychiatry, 78(8), E35–E36.
+https://doi.org/10.1016/j.biopsych.2014.12.007
+
+Baumeister, R.
+E., Bratslavsky, E., Muraven, M., & Tice, D.
+M.
+(1998).
+Ego depletion: Is the active self a limited resource?
+Journal of Personality and Social Psychology, 74(5), 1252–1265.
+https://doi.org/10.1037/0022-3514.74.5.1252
+
+Bishop, D.
+V.
+(2019).
+The psychology of experimental psychologists: Overcoming cognitive constraints to improve research: The 47th Sir Frederic Bartlett Lecture: Quarterly Journal of Experimental Psychology, 73(1), 1–19.
+https://doi.org/10.1177/1747021819886519
+
+Budd, J. M., Sievert, M., & Schultz, T. R. (1998). Phenomena of retraction: Reasons for retraction and citations to the publications. JAMA, 280(3), 296–297. https://doi.org/10.1001/jama.280.3.296
+
+Carter, E.
+C., Kofler, L.
+M., Forster, D.
+E., & McCullough, M.
+E.
+(2015).
+A series of meta-analytic tests of the depletion effect: Self-control does not seem to rely on a limited resource.
+Journal of Experimental Psychology.
+General, 144(4), 796–815.
+https://doi.org/10.1037/xge0000083
+
+Carter, T.
+J., Ferguson, M.
+J., & Hassin, R.
+R.
+(2011).
+A single exposure to the american flag shifts support toward republicanism up to 8 months later.
+Psychological Science, 22(8), 1011–1018.
+https://doi.org/10.1177/0956797611414726
+
+Caruso, E.
+M., Vohs, K.
+D., Baxter, B., & Waytz, A.
+(2013).
+Mere exposure to money increases endorsement of free-market systems and social inequality.
+Journal of Experimental Psychology: General, 142(2), 301–306.
+https://doi.org/10.1037/a0029288
+
+Coles, N.
+A., Larsen, J.
+T., & Lench, H.
+C.
+(2019).
+A meta-analysis of the facial feedback literature: Effects of facial feedback on emotional experience are small and variable.
+Psychological Bulletin, 145(6), 610–651.
+https://doi.org/10.1037/bul0000194
+
+Collins, H. M. (1985). Changing order: Replication and induction in scientific practice. Sage Publications.
+
+Earp, B. D., & Trafimow, D. (2015). Replication, falsification, and the crisis of confidence in social psychology. Frontiers in Psychology, 6, 621. https://doi.org/10.3389/fpsyg.2015.00621
+
+Etz, A., & Vandekerckhove, J. (2016). A Bayesian perspective on the Reproducibility Project: Psychology. PLOS ONE, 11(2), e0149794. https://doi.org/10.1371/journal.pone.0149794
+
+Gershman, S. J. (2019). How to never be wrong. Psychonomic Bulletin & Review, 26(1), 13–28. https://doi.org/10.3758/s13423-018-1488-8
+
+Greenberg, S. A. (2009). How citation distortions create unfounded authority: Analysis of a citation network. BMJ, 339. https://doi.org/10.1136/bmj.b2680
+
+Hagger, M. S., Chatzisarantis, N. L. D., Alberts, H., Anggono, C. O., Batailler, C., Birt, A. R., Brand, R., Brandt, M. J., Brewer, G., Bruyneel, S., Calvillo, D. P., 
+Campbell, W.
+K., Cannon, P.
+R., Carlucci, M., Carruth, N.
+P., Cheung, T., Crowell, A., De Ridder, D.
+T.
+D., Dewitte, S., … Zwienenberg, M.
+(2016).
+A multilab preregistered replication of the ego-depletion effect.
+Perspectives on Psychological Science, 11(4), 546–573.
+https://doi.org/10.1177/1745691616652873
+
+Hardwicke, T.
+E., Serghiou, S., Janiaud, P., Danchev, V., Crüwell, S., Goodman, S.
+N., & Ioannidis, J.
+P.
+A.
+(2020).
+Calibrating the Scientific Ecosystem Through Meta-Research.
+Annual Review of Statistics and Its Application, 7(1), 11–37.
+https://doi.org/10.1146/annurev-statistics-031219-041104
+
+Hardwicke, T.
+E., Thibault, R.
+T., Kosie, J., Wallach, J.
+D., Kidwell, M.
+C., & Ioannidis, J.
+(2021).
+Estimating the prevalence of transparency and reproducibility-related research practices in psychology (2014-2017).
+Perspectives on Psychological Science.
+https://doi.org/10.31222/osf.io/9sz2y
+
+Ioannidis, J. P. A. (2012). Why science is not necessarily self-correcting. Perspectives on Psychological Science, 7(6), 645–654. https://doi.org/10.1177/1745691612464056
+
+Klein, R.
+A., Ratliff, K.
+A., Vianello, M., Adams, R.
+B., Bahník, Š., Bernstein, M.
+J., Bocian, K., Brandt, M.
+J., Brooks, B., Brumbaugh, C.
+C., Cemalcilar, Z., Chandler, J., Cheong, W., Davis, W.
+E., Devos, T., Eisner, M., Frankowska, N., Furrow, D., Galliani, E.
+M., … Nosek, B.
+A.
+(2014).
+Investigating variation in replicability.
+Social Psychology, 45(3), 142–152.
+https://doi.org/10.1027/1864-9335/a000178
+
+Kühl, T., & Bertrams, A. (2019). Is learning with elaborative interrogation less desirable when learners are depleted? Frontiers in Psychology, 10, 707. https://doi.org/10.3389/fpsyg.2019.00707
+
+Kunda, Z. (1990). The case for motivated reasoning. Psychological Bulletin, 108(3), 480–498. https://doi.org/10.1037/0033-2909.108.3.480
+
+Kvarven, A., Strømland, E., & Johannesson, M.
+(2020).
+Comparing meta-analyses and preregistered multiple-laboratory replication projects.
+Nature Human Behaviour, 4(4), 423–434.
+https://doi.org/10.1038/s41562-019-0787-z
+
+Lakatos, I.
+(1970).
+Falsification and the methodology of scientific research programmes.
+In A.
+Musgrave & I.
+Lakatos (Eds.), Criticism and the Growth of Knowledge (Vol.
+4, pp.
+91–196).
+Cambridge University Press.
+https://doi.org/10.1017/CBO9781139171434.009
+
+Lauden, L. (1981). Peirce and the trivialization of the self-corrective thesis. In Science and Hypothesis (pp. 226–251). Springer.
+
+Lewandowsky, S., Ecker, U.
+K.
+H., Seifert, C.
+M., Schwarz, N., & Cook, J.
+(2012).
+Misinformation and its correction: Continued influence and successful debiasing.
+Psychological Science in the Public Interest, 13(3), 106–131.
+https://doi.org/10.1177/1529100612451018
+
+Maassen, E., van Assen, M.
+A.
+L.
+M., Nuijten, M.
+B., Olsson-Collentine, A., & Wicherts, J.
+M.
+(2020).
+Reproducibility of individual effect sizes in meta-analyses in psychology.
+PLOS ONE, 15(5), e0233107.
+https://doi.org/10.1371/journal.pone.0233107
+
+Maxwell, S.
+E., Lau, M.
+Y., & Howard, G.
+S.
+(2015).
+Is psychology suffering from a replication crisis?
+What does “failure to replicate” really mean?
+American Psychologist, 70(6), 487–498.
+https://doi.org/10.1037/a0039400
+
+McDiarmid, A., Tullett, A., Whitt, C.
+M., Vazire, S., Smaldino, P.
+E., & Stephens, E.
+E.
+(2021).
+Self-correction in psychological science: How do psychologists update their beliefs in response to replications?
+PsyArXiv.
+https://doi.org/10.31234/osf.io/hjcm4
+
+Nelson, L. D., Simmons, J., & Simonsohn, U. (2018). Psychology’s Renaissance. Annual Review of Psychology, 69(1), 511–534. https://doi.org/10.1146/annurev-psych-122216-011836
+
+Neuliep, J. W., & Crandall, R. (1993). Everyone was wrong: There are lots of replications out there. Journal of Social Behavior and Personality, 8(6), 1–8.
+
+Nickerson, R. S. (1998). Confirmation Bias: A Ubiquitous Phenomenon in Many Guises. Review of General Psychology, 2(2), 175–220. https://doi.org/10.1037/1089-2680.2.2.175
+
+Open Science Collaboration. (2015). Estimating the reproducibility of psychological science. Science, 349(6251), aac4716–aac4716. https://doi.org/10.1126/science.aac4716
+
+Peterson, D., & Panofsky, A. (2020). Self-correction in science: The diagnostic and integrative motives for replication. SocArXiv. https://doi.org/10.31235/osf.io/96qxv
+
+Polanin, J.
+R., Hennessy, E.
+A., & Tsuji, S.
+(2020).
+Transparency and reproducibility of meta-analyses in psychology: A meta-review.
+Perspectives on Psychological Science, 15(4), 1026–1041.
+https://doi.org/10.1177/1745691620906416
+
+Rohrer, D., Pashler, H., & Harris, C.
+R.
+(2015).
+Do subtle reminders of money change people’s political views?
+Journal of Experimental Psychology: General, 144(4), e73–e85.
+https://doi.org/10.1037/xge0000058
+
+Simons, D.
+J., Holcombe, A.
+O., & Spellman, B.
+A.
+(2014).
+An introduction to Registered Replication Reports at Perspectives on Psychological Science.
+Perspectives on Psychological Science, 9(5), 552–555.
+https://doi.org/10.1177/1745691614543974
+
+Sripada, C., Kessler, D., & Jonides, J.
+(2014).
+Methylphenidate blocks effort-induced depletion of regulatory control in healthy volunteers.
+Psychological Science, 25(6), 1227–1234.
+https://doi.org/10.1177/0956797614526415
+
+Strack, F. (2017). From data to truth in psychological science. A personal perspective. Frontiers in Psychology, 8, 702. https://doi.org/10.3389/fpsyg.2017.00702
+
+Strack, F., Martin, L.
+L., & Stepper, S.
+(1988).
+Inhibiting and facilitating conditions of the human smile: A nonobtrusive test of the Facial Feedback Hypothesis.
+Journal of Personality and Social Psychology, 54(5), 768–777.
+https://doi.org/10.1037/0022-3514.54.5.768
+
+Tatsioni, A., Bonitsis, N. G., & Ioannidis, J. P. A. (2007). Persistence of contradicted claims in the literature. JAMA, 298(21), 2517–2526. https://doi.org/10.1001/jama.298.21.2517
+
+Vadillo, M. A. (2019). Ego depletion may disappear by 2020. Social Psychology, 50(5–6), 282–291. https://doi.org/10.1027/1864-9335/a000375
+
+Vadillo, M.
+A., Hardwicke, T.
+E., & Shanks, D.
+R.
+(2016).
+Selection bias, vote counting, and money-priming effects: A comment on Rohrer, Pashler, and Harris (2015) and Vohs (2015).
+Journal of Experimental Psychology: General, 145(5), 655–663.
+https://doi.org/10.1037/xge0000157
+
+Vazire, S., & Holcombe, A. O. (2020). Where are the self-correcting mechanisms in science? PsyArXiv. https://doi.org/10.31234/osf.io/kgqzt
+
+Wagenmakers, E.-J., Beek, T., Dijkhoff, L., Gronau, Q.
+F., Acosta, A., Adams, R.
+B., Albohn, D.
+N., Allard, E.
+S., Benning, S.
+D., Blouin-Hudon, E.-M., Bulnes, L.
+C., Caldwell, T.
+L., Calin-Jageman, R.
+J., Capaldi, C.
+A., Carfagno, N.
+S., Chasten, K.
+T., Cleeremans, A., Connell, L., DeCicco, J.
+M., … Zwaan, R.
+A.
+(2016).
+Registered Replication Report: Strack, Martin, & Stepper (1988).
+Perspectives on Psychological Science, 11(6), 917–928.
+https://doi.org/10.1177/1745691616674458
+
+Xie, Y. (2017). Dynamic documents with R and knitr. CRC Press.
+
+Zotero. (2019, June 14). Retracted item notifications with retraction watch integration. https://www.zotero.org/blog/retracted-item-notifications
+
+Zwaan, R. A., Etz, A., Lucas, R. E., & Donnellan, M. B. (2018). Making replication mainstream. Behavioral and Brain Sciences, 41, e120. https://doi.org/10.1017/S0140525X17001972
 
@@
 ```
-
-# References
-
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in

### 2021-02-06 manuscript/appendix.Rmd
initial transfer of methods and results from google doc to r markdown complete
transfer from google doc to appendix complete

@@
-# Appendix A: Protocol Changes {#appA}
+# Protocol amendments {#appA}
 
-# Appendix B title {#appB}
+The study protocol (rationale, methods, and analysis plan) was pre-registered on April 7th 2018 (https://osf.io/eh5qd/).
+In order to extend the sampling frame and make other minor methodological adjustments, an amended protocol was registered part way through data collection on May 1st 2019 (https://osf.io/pdvb5/).
+Appendix Table \@ref(tab:protocolAmendments) outlines amendments between the pre-registered protocol, amended protocol, and final report.
 
-Tabular data for citation curves and citation valence
+```{r protocolAmendments}
+tibble(
+  "Issue" = c("Sampling frame", "Definition of pre-replication assessment period", "Exclusion of replication year from qualitative assessment", "Automated extraction of citation contexts and replication co-citation", "Counter-arguments", "Dual-coding for citation valence classification", "Examination of characteristics of articles providing counter-arguments.","Renaming of two citation patterns"),
+  "Amendment" = c("The end date of the sampling frame was initially April 6th 2018 (see pre-registered protocol).
+Later, we decided to extend the sampling frame to April 26th 2019 to learn more about how the citation curves developed over time (see amended protocol).
+Finally, we decided to extend the sampling frame to December 31st 2019 to ensure we had all citations for the year 2019.", "Initially, the pre-replication assessment period was defined as 2015 (i.e., replication year - 1) for the RRR cases and all pre-replication years for the other cases.
+In practice, when the replication year - 1 definition is used for all cases, only one study (Caruso) has a citing article outside of this period, and it is only a single article.
+Therefore, we have decided that it is more straightforward to use the same definition of the pre-replication assessment period (replication year - 1) for all cases.", "The pre-registered protocol neglected to mention that we intended to exclude the replication year from qualitative assessment.
+This change was documented in the amended protocol.", "Initially, we intended to use Matlab code to partly automate the extraction of ‘citation contexts’ (text surrounding in-text citations) and identify replication co-citation.
+However, we found that the size of the relevant citation context varied substantially between articles, so we decided to extract all citation contexts manually.
+We also decided to identify replication co-citation manually too.
+These changes were documented in the amended protocol.","In the amended protocol, we clarified that we would attempt to identify counter-arguments in all articles that cited the original study and the replication study.", "Initially, we planned to dual code small batches of articles depending on the degree of inter-rater reliability.
+However, after encountering a number of somewhat ambiguous cases in our initial coding, we decided to dual-code all articles in order to reduce subjectivity and increase consistency between coders.
+This is documented in the amended protocol.", "After completing the study we thought it would be informative to examine the article types and journals of articles that provided counter-arguments, and any overlap in authorship with the original study authors or prior collaborators of the first authors of the original studies.
+These analyses were not pre-registered.", "After completing the study we decided to revise the following terminology from the protocol in order to improve clarity: “principled defence” was changed to “explicit defence” and “hollow defence” was changed to “absent defence”.")
+) %>%
+  kable(booktabs=T,longtable=T,caption = "Protocol amendments.") %>%
+  column_spec(1,width='4cm') %>%
+  column_spec(2,width='11cm')
+```
 
-```{r results = "asis"}
+# Random sampling and extrapolation procedures for Baumeister case study {#appB}
+
+Due to a large number of citations, qualitative assessment in the Baumeister case was based on a random sample of 40% of citing articles in each of the pre-replication and post-replication time periods.
+We used the ‘sample’ function in the programming language R to draw random samples.
+Because the sampling frame was extended on three occasions (see Supplementary Information A), it was necessary to obtain random samples three times in order to maintain coverage of 40% for each time period.
+As we aimed to achieve a 40% sample across each time period, an equal sample in each individual year was not guaranteed, and in fact the proportion of citations coded in the post-refutation period does differ between years (48% 49%, and 23% respectively).
+
+We extrapolated the valence classifications and co-citation of replications from the random sample to all citing articles in the pre-replication (2015) and post-replication (2017, 2018, 2019) assessment periods.
+To do this, we multiplied each category count in a given year by the reciprocal of the proportion coded in that year.
+For example, in 2017, we classified 71 citations as ‘favorable’ amongst the 48% of citations coded in that year, so to extrapolate we computed 71*1/0.48 = 148.
+
+# Tabular data for citation curves, citation valence, and co-citation of replications {#appC}
+
+```{r tabularData, results = "asis"}
 options(knitr.kable.NA = "-") # set symbol to use for NA cells
@@
   select(year = pubYear, `citing\narticles` = citesTarget, excluded, unclassifiable, favourable, equivocal, unfavourable, citesRep) %>%
-  kable(caption = 'Counts (percentages) for citations to the original study, exclusions, citation valence, and concomitant citations to replication study in the time periods subjected to qualitative analysis.
-Percentages for the Baumeister case are extrapolations based on 40\\% random samples drawn from the pre-replication and post-replication time periods.', booktabs=T, longtable=T) %>%
+  kable(caption = 'Raw counts and percentages for citations to the original study, exclusions, citation valence, and concomitant citations to the replication study in qualitative assessment period.
+In the Baumeister case, numbers in square brackets in the citations column are the number of citing articles included in the random samples drawn from the pre-replication and post-replication time periods.
+Percentages for the Baumeister case are extrapolations based on the random samples.', booktabs=T, longtable=T) %>%
   pack_rows("Baumeister", 1, 4) %>%
@@
 ```{r}
-# DATA SUMMARISED FOR POST-REPLICATION PERIOD
-d_summary %>%
+# data summarised for post-replication period
+d_summary_postrep <- d_summary %>%
   filter(year == 'post-replication') %>%

### 2021-02-06 manuscript/manuscript.Rmd
trasnferred abstract from google docs to markdown

@@
 abstract: >
-   Abstract here.
+   Replication studies that contradict published findings may facilitate scientific self-correction by triggering a reappraisal of the original findings; however, the research community's response to replication studies has not been studied systematically.
+One approach for gauging responses to replication results is to examine how they impact citations to original studies.
+In this study, we explored post-replication citation patterns in the context of four prominent multi-laboratory replication attempts published in the field of psychology that strongly contradicted and outweighed prior findings.
+Generally, we observed a small post-replication decline in the number of favourable citations and a small increase in unfavourable citations.
+This indicates only modest corrective effects and implies considerable perpetuation of belief in the original findings.
+Replication results that strongly contradict an original finding do not necessarily nullify its credibility; however, one might at least expect the replication results to be acknowledged and explicitly debated in subsequent literature.
+By contrast, we found substantial citation bias: the majority of articles citing the original studies neglected to cite the relevant replication.
+Of those articles that did cite the replication, but continued to cite the original study favourably, approximately half offered an explicit defence of the original study.
+Our findings suggest that even replication results that strongly contradict original findings do not necessarily prompt a corrective response from the research community.
 

### 2021-02-06 manuscript/manuscript.Rmd
completed transfer of introduction from google doc to R markdown

@@
 
-bibliography      : ["references.bib"]
-
 figsintext        : yes
@@
 
+# Introduction
+
+It is often assumed that science is a self-correcting enterprise: the veracity of scientific knowledge should progressively improve as inaccurate claims are abandoned and accurate claims are reinforced (Vazire & Holcombe, 2020).
+Replication studies are considered to be a key driver of this process because they may indicate that prior results are exaggerated or erroneous (Ioannidis, 2012; Zwaan et al., 2018).
+Although interpreting the outcome of replication studies is not necessarily straightforward (Collins, 1985; Earp & Trafimow, 2015; Maxwell et al., 2015), one might expect a replication result that strongly contradicts[^footnote1] and outweighs the results of a prior (‘original’) study to impact how that study is cited in subsequent academic literature.
+For example, if a replication undermines belief in the credibility of an original finding, one might expect to see a change in the frequency and valence (favourability) of citations to the original study; specifically, a decrease in favourable citations accompanied by an increase in unfavourable citations.
+However, as discussed below, a variety of interesting patterns could emerge depending on how the research community responds to a replication result.
+The goal of the present study was to empirically explore and describe post-replication citation patterns in the context of four prominent multi-laboratory replication attempts published in the field of psychology that strongly contradicted and outweighed the findings of prior studies.
+
+[^footnote1]: Determining whether a replication finding contradicts an original finding is a complex issue which we largely side-step here by only examining cases that are strongly contradictory - that is, according to several reasonable quantitative criteria (e.g., effect size magnitude, p-values, Bayes Factors, etc.) the results of the original study and replication study lead to opposing inferences (e.g., absence vs.
+presence of an effect).
+Also note that one can accept that results are strongly contradictory and remain agnostic about the explanation for the contradiction.
+
+Table \@ref(tab:patterns) outlines several citation patterns that might follow a contradictory replication result, each reflecting different types of response by the research community.
+We have tentatively categorised these response patterns as being ‘progressive’ or ‘regressive’ depending on their expected impact on the accumulation of scientific knowledge[^footnote2].
+The first set of patterns, belief correction/perpetuation, refer to what is often considered the primary functional role of (contradictory) replication studies - to change belief in the credibility of exaggerated or erroneous original findings (Ioannidis, 2012; Vazire & Holcombe, 2020; Zwaan et al., 2018).
+In the absence of a explicit defence of an original finding that convincingly explains a strongly contradictory result that outweighs an original finding (see ‘explicit defence’ below), a progressive response might involve a decrease in favourable citations and an increase in unfavourable citations, reflecting updated beliefs about the credibility of the original finding.
+Conversely, a regressive response might involve maintenance of (or even increase in) favourable citations, and relatively few unfavourable citations, suggesting a perpetuation of belief in the credibility of the original finding despite the contradictory replication result.
+Prior research has documented how favourable citations to observational epidemiology studies can persist despite the claims of those studies being strongly contradicted in subsequent randomized trials (Tatsioni et al., 2007).
+Similarly, it has been reported that even when articles are retracted, they can continue to receive favourable citations (Budd et al., 1998).
+Thus, there is evidence that belief in the credibility of original findings can perpetuate even when subsequent events cast doubt on their credibility; however, we are unaware of similar evidence in the context of replication studies that were explicitly designed to test the reliability of prior findings.
+
+[^footnote2]: This terminology was inspired by, but does not directly mirror, terminology proposed by Imre Lakatos in his work on the rationality of the research community’s response when a scientific theory is contracted by empirical evidence (Lakatos, 1970).
+
+The second set of patterns in Table \@ref(tab:patterns), citation balance/bias, generally refers to whether positive (supportive) evidence is preferentially cited relative to negative (non-supportive) evidence.
+This pattern has previously been observed in the context of research on inclusion body myositis; citation content analysis showed that the accumulating literature heavily cited the theory that beta amyloid is involved, ignoring multiple studies that contradicted this theory (Greenberg, 2009; also see Bastiaansen et al., 2015).
+In the present study, these patterns specifically refer to whether articles citing an original study also cite the subsequent contradictory replication study.
+A progressive response would be to cite both studies (citation balance), as this involves considering and reporting highly relevant evidence (even if the implications of the replication are disputed, see ‘explicit defence’ below).
+By contrast, citation bias could occur if articles citing an original study neglect to cite a relevant replication study.
+Regardless of whether this occurs through lack of awareness or deliberate omission, it can be considered a regressive response because highly relevant evidence is not being reported or considered.
+
+The third set of patterns in Table \@ref(tab:patterns), explicit/absent defence, refers to whether researchers who continue to favourably cite the original finding despite the strongly contradictory replication result offer a concrete defence of the original study.
+As implied above, even when the result of a replication study strongly contradicts the results of an original study, this does not necessarily nullify the credibility of the original findings; the same criticisms that one might apply to an original study in order to infer that its findings are erroneous or exaggerated may also be applied to replication studies (Collins, 1985; Earp & Trafimow, 2015; Maxwell et al., 2015).
+Thus, if proponents of the original claim mount an explicit defence that counters the implications of the replication results, this might still be considered a progressive response (though obviously one could disagree with the arguments that are presented).
+By contrast, if favourable citations to the original study are not accompanied by explicit argumentation about the replication result (an absent defence) this might be considered a regressive response because the replication result is apparently discounted without providing any rationale.
+
 ```{r patterns}
 tibble(
-  "Progressive responses" = c(1, 2),
-  "Regressive responses" = c(1, 2)
+  "Progressive responses" = c("*Belief correction*
+A decrease in the number of favourable citations may reflect a decline in belief in the credibility of the original finding (a belief correction pattern).
+This may be accompanied by a relative increase in the number of unfavourable citations (an active belief correction pattern) or relatively fewer/no unfavourable citations (a passive belief correction pattern).", "*Citation balance*
+Articles that cite the original study may also cite the contradictory replication, reflecting that relevant evidence has been considered (a balanced citation pattern).", "*Explicit defence*
+Articles that favourably cite the original study and unfavourably cite the contradictory replication may offer concrete counter-arguments that state why the credibility of the original finding has not been undermined (an explicit defence pattern)."),
+  "Regressive responses" = c("*Belief perpetuation*
+A maintenance of, or increase in, the number of favourable citations may reflect that a maintenance of or increase in belief in the credibility of the original finding (a belief perpetuation pattern).
+This may be accompanied by a relative increase in unfavourable citations (a challenged belief perpetuation pattern) or relatively fewer/no unfavourable citations (an unchallenged belief perpetuation pattern).", "*Citation bias*
+Articles that cite the original study may neglect to also cite the contradictory replication, reflecting that relevant evidence has been neglected, either through lack of awareness or deliberate omission (a citation bias pattern).", "*Absent defence*
+Articles that favourably cite the original study and unfavourably cite the contradictory replication may offer no concrete counter-arguments that state why the credibility of the original finding has not been undermined (an absent defence pattern).")
 ) %>%
-  apa_table(caption = "Progressive or regressive responses to strongly contradictory replication results and their expected impact on citation patterns for original studies.")
-```
+  kable(booktabs=T,longtable=T,caption = "Progressive or regressive responses to strongly contradictory replication results and their expected impact on citation patterns for original studies.") %>%
+  column_spec(1,'8.5cm') %>%
+  column_spec(2,'8.5cm')
+```
+
+In the present study, we explored the post-replication citation patterns described above in the context of four case studies in the field of psychology where the findings of a replication study strongly contradicted and outweighed the findings of an original study (Table \@ref(tab:studyDetails)).
+In two of the cases, the replication studies were part of a single ‘Many Labs’ project (Klein et al., 2014) and addressed the ‘flag priming effect’ (T.
+J.
+Carter et al., 2011) and ‘money priming effect’ (Caruso et al., 2013) respectively.
+The other two cases involved ‘Registered Replication Reports’ (RRRs; Simons et al., 2014), examining influential demonstrations of the ‘facial feedback effect’ (cf.
+Strack et al., 1988; Wagenmakers et al., 2016) and the ‘ego-depletion effect’ (cf.
+Baumeister et al., 1998; Hagger et al., 2016; Sripada et al., 2014) respectively.
+For methodological reasons (see Hagger et al., 2016), the ego-depletion replication was aimed at a classic study in the field (Baumeister et al., 1998), but actually employed a modified computer-based version of the original paradigm (Sripada et al., 2014).
+Thus, for this particular case study we examined citation patterns to both of these original studies.
+We chose these particular case studies because they involved prominent pre-registered multi-laboratory replication attempts with sample sizes between 23 - 211 times larger than the original studies, thus providing highly visible and highly credible evidence that strongly contradicted and outweighed earlier findings[^footnote3].
+This facilitates additional interpretative clarity about the citation patterns one might expect to observe.
+
+[^footnote3]: We deliberately focused on a select group of case studies rather than other potentially larger samples in order to aid interpretative clarity; for example, the extent to which the results of the large scale Reproducibility Project in Psychology (RPP; Open Science Collaboration, 2015) actually contradicted the original studies has been contested (Etz & Vandekerckhove, 2016).
+Additionally, original studies were not actually cited in the RPP research report, which may have diluted awareness about relevant replication studies.
+
+The present study was exploratory in nature and intended to provide descriptive observations rather than test hypotheses.
+The three sets of expected citation patterns outlined in Table \@ref(tab:patterns) were used to guide our study design and interpretation, but we do not claim that this is a comprehensive typology of the post-replication patterns that may occur.
+Such patterns may be more complex and idiosyncratic in other topic domains.
+In order to examine patterns of belief correction/perpetuation, we downloaded citation histories (a list of citing articles) for each original study and classified the valence of a set of pre-replication and post-replication citations as favourable, equivocal, or unfavourable.
+To examine patterns of citation balance/bias, we manually checked whether post-replication citations of the original study were accompanied by citations to the replication study.
+Finally, to examine patterns of explicit/absent defence, we extracted and categorised any counter-arguments offered in articles that cited the replication.
 

### 2021-02-06 analysis/preprocessing.R
now outputting csv processed files as well as rds

@@ -19,6 +19,7 @@ d_citations <- d_citations %>%
   mutate(pubYear = as.numeric(PY))
 
 ### Save file
+write_csv(d_citations, path = here('data','processed','/d_citations.csv'))
 save(d_citations, file = here('data','processed','/d_citations.rds'))
 
 ## Reference data
@@ -37,6 +38,7 @@ d_reference <- d_reference %>%
   mutate(pubYear = as.numeric(year))
 
 ### Save file
+write_csv(d_reference, path = here('data','processed','/d_reference.csv'))
 save(d_reference, file = here('data','processed','/d_reference.rds'))
 
 ## Content analysis data
@@ -85,6 +87,7 @@ d_contentAnalysis <- d_contentAnalysis %>%
                               'Data synthesis' = 'Data synthesis - meta-analysis'))
 
 ### Save file
+write_csv(d_contentAnalysis, path = here('data','processed','/d_contentAnalysis.csv'))
 save(d_contentAnalysis, file = here('data','processed','/d_contentAnalysis.rds'))
 
 ## Coauthor data
@@ -101,5 +104,6 @@ for(i in caseNames){ # loop through case names
 ### Apply munging
 
 ### Save file
+write_csv(d_coauthors, path = here('data','processed','/d_coauthors.csv'))
 save(d_coauthors, file = here('data','processed','/d_coauthors.rds'))
    

### 2021-02-06 manuscript/manuscript.Rmd
turned off tidylog

@@
 library(tidyverse) # for data munging
-library(tidylog) # inline tidyverse feedback
 library(here) # for finding files

### 2021-02-07 manuscript/manuscript.Rmd
applied styler to analysis files

@@
 abstract: >
-   Replication studies that contradict published findings may facilitate scientific self-correction by triggering a reappraisal of the original findings; however, the research community's response to replication studies has not been studied systematically.
+   Replication studies that contradict prior findings may facilitate scientific self-correction by triggering a reappraisal of the original studies; however, the research community's response to replication results has not been studied systematically.
 One approach for gauging responses to replication results is to examine how they impact citations to original studies.
@@
 Replication results that strongly contradict an original finding do not necessarily nullify its credibility; however, one might at least expect the replication results to be acknowledged and explicitly debated in subsequent literature.
-By contrast, we found substantial citation bias: the majority of articles citing the original studies neglected to cite the relevant replication.
+By contrast, we found substantial citation bias: the majority of articles citing the original studies neglected to cite relevant replication results.
 Of those articles that did cite the replication, but continued to cite the original study favourably, approximately half offered an explicit defence of the original study.
@@
 We have tentatively categorised these response patterns as being ‘progressive’ or ‘regressive’ depending on their expected impact on the accumulation of scientific knowledge[^footnote2].
-The first set of patterns, belief correction/perpetuation, refer to what is often considered the primary functional role of (contradictory) replication studies - to change belief in the credibility of exaggerated or erroneous original findings (Ioannidis, 2012; Vazire & Holcombe, 2020; Zwaan et al., 2018).
-In the absence of a explicit defence of an original finding that convincingly explains a strongly contradictory result that outweighs an original finding (see ‘explicit defence’ below), a progressive response might involve a decrease in favourable citations and an increase in unfavourable citations, reflecting updated beliefs about the credibility of the original finding.
+The first set of patterns, belief correction/perpetuation, refer to what is often considered a primary functional role of (contradictory) replication studies - to change belief in the credibility of exaggerated or erroneous original findings (Ioannidis, 2012; Vazire & Holcombe, 2020; Zwaan et al., 2018).
+In the absence of an explicit defence of an original study that convincingly explains a strongly contradictory replication result (see ‘explicit defence’ below), a progressive response might involve a decrease in favourable citations and an increase in unfavourable citations, reflecting updated beliefs about the credibility of the original finding.
 Conversely, a regressive response might involve maintenance of (or even increase in) favourable citations, and relatively few unfavourable citations, suggesting a perpetuation of belief in the credibility of the original finding despite the contradictory replication result.
@@
 ) %>%
-  kable(booktabs=T,longtable=T,caption = "Progressive or regressive responses to strongly contradictory replication results and their expected impact on citation patterns for original studies.") %>%
-  column_spec(1,'8.5cm') %>%
-  column_spec(2,'8.5cm')
+  kable(booktabs = T, longtable = T, caption = "Progressive or regressive responses to strongly contradictory replication results and their expected impact on citation patterns for original studies.") %>%
+  column_spec(1, "8.5cm") %>%
+  column_spec(2, "8.5cm")
 ```
@@
   pull(authors) %>%
-  str_c(collapse = '; ')
+  str_c(collapse = "; ")
 
@@
 ca_auth_freq <- ca_auth_list %>%
-  str_split(';') %>%
-  table(dnn = 'name') %>%
+  str_split(";") %>%
+  table(dnn = "name") %>%
   as.data.frame() %>%
@@
   rowwise() %>%
-  mutate(surname = str_split(name,',')[[1]][1]) # extract surnames
+  mutate(surname = str_split(name, ",")[[1]][1]) # extract surnames
 
 # frequency of original authors also authoring counter-arg papers
-ca_auth_surnames <- ca_auth_freq$surname %>% str_c(collapse = ' ')
+ca_auth_surnames <- ca_auth_freq$surname %>% str_c(collapse = " ")
 org_auths_ca <- org_auths_list[str_detect(ca_auth_surnames, org_auths_list)]
@@
 All data, materials, and analysis scripts related to this study are publicly available on the Open Science Framework (https://osf.io/w8h2q/).
-To facilitate reproducibility this manuscript was written by interleaving regular prose and analysis code using knitr (Xie, 2017) and papaja (Aust & Barth, 2020), and is available in a Code Ocean container (TO BE ADDED) which re-creates the software environment in which the original analyses were performed.
+To facilitate reproducibility this manuscript was written by interleaving regular prose and analysis code using knitr (Xie, 2017) and papaja (Aust & Barth, 2020), and is available in a Code Ocean container (https://doi.org/10.24433/CO.4225975.v1) which re-creates the software environment in which the original analyses were performed.
 

### 2021-02-07 analysis/functions.R

@@ -117,33 +117,35 @@ standardizeCitations <- function(citationData, referenceData, contentData, repli
       replicationYear = # replication year
       ifelse(pubYear == replicationYear, TRUE, FALSE)
     )
-  
-  # add a row summarise the entire post-replication period
-  
-  # define post-replication period
-  postRepPeriod <- as.character(seq(replicationYear+1,2019,1))
 
-  d <- rbind(d, d %>% 
+  # add a row summarise the entire post-replication period
+
+  # define post-replication period
+  postRepPeriod <- as.character(seq(replicationYear + 1, 2019, 1))
+
+  d <- rbind(d, d %>%
     filter(year %in% postRepPeriod) %>%
-    summarise(case = thisCase, 
-              pubYear = 'post-replication', 
-              citesTarget = sum(citesTarget), 
-              citesTarget_std =sum(citesTarget_std), 
-              year = 'post-replication', 
-              citesRef = sum(citesRef),
-              citesRef_std = sum(citesRef_std), 
-              equivocal = sum(equivocal), 
-              favourable = sum(favourable), 
-              unclassifiable = sum(unclassifiable),
-              unfavourable = sum(unfavourable),
-              citesRep_no = sum(citesRep_no), 
-              citesRep_yes = sum(citesRep_yes), 
-              excluded_no = sum(excluded_no),
-              excluded_yes = sum(excluded_yes),
-              counterargs_no = sum(counterargs_no),
-              counterargs_yes =sum(counterargs_yes), 
-              standardizationYear = FALSE, 
-              replicationYear = FALSE))
+    summarise(
+      case = thisCase,
+      pubYear = "post-replication",
+      citesTarget = sum(citesTarget),
+      citesTarget_std = sum(citesTarget_std),
+      year = "post-replication",
+      citesRef = sum(citesRef),
+      citesRef_std = sum(citesRef_std),
+      equivocal = sum(equivocal),
+      favourable = sum(favourable),
+      unclassifiable = sum(unclassifiable),
+      unfavourable = sum(unfavourable),
+      citesRep_no = sum(citesRep_no),
+      citesRep_yes = sum(citesRep_yes),
+      excluded_no = sum(excluded_no),
+      excluded_yes = sum(excluded_yes),
+      counterargs_no = sum(counterargs_no),
+      counterargs_yes = sum(counterargs_yes),
+      standardizationYear = FALSE,
+      replicationYear = FALSE
+    ))
 
[…]

### 2021-02-07 analysis/preprocessing.R

@@ -1,14 +1,14 @@
 # Preprocessing
 
 # identify the five different case studies we are looking at
-caseNames <- c('baumeister', 'sripada', 'strack', 'carter', 'caruso')
+caseNames <- c("baumeister", "sripada", "strack", "carter", "caruso")
 
 ## Citation data
 
 ### Load files
 d_citations <- data.frame() # create empty list to hold data frames for each case
-for(i in caseNames){ # loop through case names
-  filePath <- here('data','primary', paste0(i,'Citations.csv')) # create file path
+for (i in caseNames) { # loop through case names
+  filePath <- here("data", "primary", paste0(i, "Citations.csv")) # create file path
   d <- read_csv(filePath, col_types = cols(.default = "c")) # load the data
   d <- d %>% mutate(case = i)
   d_citations <- bind_rows(d_citations, d) # append to dataframe
@@ -19,15 +19,15 @@ d_citations <- d_citations %>%
   mutate(pubYear = as.numeric(PY))
 
 ### Save file
-write_csv(d_citations, path = here('data','processed','/d_citations.csv'))
-save(d_citations, file = here('data','processed','/d_citations.rds'))
+write_csv(d_citations, path = here("data", "processed", "/d_citations.csv"))
+save(d_citations, file = here("data", "processed", "/d_citations.rds"))
 
 ## Reference data
 
 ### Load files
 d_reference <- data.frame() # create empty list to hold data frames for each case
-for(i in caseNames){ # loop through case names
-  filePath <- here('data','primary', paste0(i,'ReferenceClass.csv')) # create file path
+for (i in caseNames) { # loop through case names
+  filePath <- here("data", "primary", paste0(i, "ReferenceClass.csv")) # create file path
   d <- read_csv(filePath) # load the data
   d <- d %>% mutate(case = i)
   d_reference <- bind_rows(d_reference, d) # append to dataframe
@@ -38,22 +38,22 @@ d_reference <- d_reference %>%
   mutate(pubYear = as.numeric(year))
 
 ### Save file
-write_csv(d_reference, path = here('data','processed','/d_reference.csv'))
-save(d_reference, file = here('data','processed','/d_reference.rds'))
+write_csv(d_reference, path = here("data", "processed", "/d_reference.csv"))
+save(d_reference, file = here("data", "processed", "/d_reference.rds"))
 
 ## Content analysis data
 
 ### Load files
[…]

### 2021-02-07 manuscript/appendix.Rmd
finalising alignment of google doc and R markdown

@@
 tibble(
-  "Issue" = c("Sampling frame", "Definition of pre-replication assessment period", "Exclusion of replication year from qualitative assessment", "Automated extraction of citation contexts and replication co-citation", "Counter-arguments", "Dual-coding for citation valence classification", "Examination of characteristics of articles providing counter-arguments.","Renaming of two citation patterns"),
+  "Issue" = c("Sampling frame", "Definition of pre-replication assessment period", "Exclusion of replication year from qualitative assessment", "Automated extraction of citation contexts and replication co-citation", "Counter-arguments", "Dual-coding for citation valence classification", "Examination of characteristics of articles providing counter-arguments.","Renaming of two citation patterns", "Data exclusions"),
   "Amendment" = c("The end date of the sampling frame was initially April 6th 2018 (see pre-registered protocol).
 Later, we decided to extend the sampling frame to April 26th 2019 to learn more about how the citation curves developed over time (see amended protocol).
-Finally, we decided to extend the sampling frame to December 31st 2019 to ensure we had all citations for the year 2019.", "Initially, the pre-replication assessment period was defined as 2015 (i.e., replication year - 1) for the RRR cases and all pre-replication years for the other cases.
+Finally, we decided to extend the sampling frame to December 31st 2019 to ensure we had all citations for the year 2019.", "Initially, the pre-replication assessment period was defined as 2015 (i.e., replication year - 1) for the Registered Replication Report cases and all pre-replication years for the other cases.
 In practice, when the replication year - 1 definition is used for all cases, only one study (Caruso) has a citing article outside of this period, and it is only a single article.
@@
 This is documented in the amended protocol.", "After completing the study we thought it would be informative to examine the article types and journals of articles that provided counter-arguments, and any overlap in authorship with the original study authors or prior collaborators of the first authors of the original studies.
-These analyses were not pre-registered.", "After completing the study we decided to revise the following terminology from the protocol in order to improve clarity: “principled defence” was changed to “explicit defence” and “hollow defence” was changed to “absent defence”.")
+These analyses were not pre-registered.", "After completing the study we decided to revise the following terminology from the protocol in order to improve clarity: “principled defence” was changed to “explicit defence” and “hollow defence” was changed to “absent defence”.", "The criteria to exclude data from qualitative assessment (no access, non-English language, and no original study citation in the main text, were not pre-registered.
+These exclusions were necessary because they precluded the qualitative assessment.")
 ) %>%

### 2021-02-07 manuscript/manuscript.Rmd
finalising alignment of google doc and R markdown

@@
 
-The second set of patterns in Table \@ref(tab:patterns), citation balance/bias, generally refers to whether positive (supportive) evidence is preferentially cited relative to negative (non-supportive) evidence.
-This pattern has previously been observed in the context of research on inclusion body myositis; citation content analysis showed that the accumulating literature heavily cited the theory that beta amyloid is involved, ignoring multiple studies that contradicted this theory (Greenberg, 2009; also see Bastiaansen et al., 2015).
+The second set of patterns in Table \@ref(tab:patterns), citation balance/bias, generally refers to whether positive (supportive) evidence is preferentially cited relative to negative (non-supportive) evidence (Bastiaansen et al., 2015; Greenberg, 2009).
+This pattern has previously been observed in the context of research on inclusion body myositis; citation content analysis showed that the accumulating literature heavily cited the theory that beta amyloid is involved, ignoring multiple studies that contradicted this theory (Greenberg, 2009).
 In the present study, these patterns specifically refer to whether articles citing an original study also cite the subsequent contradictory replication study.
@@
 By contrast, citation bias could occur if articles citing an original study neglect to cite a relevant replication study.
-Regardless of whether this occurs through lack of awareness or deliberate omission, it can be considered a regressive response because highly relevant evidence is not being reported or considered.
+Regardless of whether this occurs through lack of awareness or deliberate omission, it can be considered a regressive response pattern because highly relevant evidence is not being reported or considered.
 
@@
 Carter et al., 2011) and ‘money priming effect’ (Caruso et al., 2013) respectively.
-The other two cases involved ‘Registered Replication Reports’ (RRRs; Simons et al., 2014), examining influential demonstrations of the ‘facial feedback effect’ (cf.
+The other two cases involved ‘Registered Replication Reports’ (Simons et al., 2014), examining influential demonstrations of the ‘facial feedback effect’ (cf.
 Strack et al., 1988; Wagenmakers et al., 2016) and the ‘ego-depletion effect’ (cf.
@@
 For methodological reasons (see Hagger et al., 2016), the ego-depletion replication was aimed at a classic study in the field (Baumeister et al., 1998), but actually employed a modified computer-based version of the original paradigm (Sripada et al., 2014).
-Thus, for this particular case study we examined citation patterns to both of these original studies.
+For this particular case study we examined citation patterns to both of these original studies.
 We chose these particular case studies because they involved prominent pre-registered multi-laboratory replication attempts with sample sizes between 23 - 211 times larger than the original studies, thus providing highly visible and highly credible evidence that strongly contradicted and outweighed earlier findings[^footnote3].
@@
 Such patterns may be more complex and idiosyncratic in other topic domains.
-In order to examine patterns of belief correction/perpetuation, we downloaded citation histories (a list of citing articles) for each original study and classified the valence of a set of pre-replication and post-replication citations as favourable, equivocal, or unfavourable.
-To examine patterns of citation balance/bias, we manually checked whether post-replication citations of the original study were accompanied by citations to the replication study.
-Finally, to examine patterns of explicit/absent defence, we extracted and categorised any counter-arguments offered in articles that cited the replication.
+In order to examine belief correction/perpetuation patterns, we downloaded citation histories (a list of citing articles) for each original study and classified the valence (favourable, equivocal, or unfavourable) of a set of pre-replication and post-replication citations.
+To examine citation balance/bias patterns, we manually checked whether post-replication citations of the original study were accompanied by citations to the replication study.
+Finally, to examine explicit/absent defence patterns, we extracted and categorised any counter-arguments offered in articles that cited the replication.
 
@@
 Citation counts were standardized in each case study by setting the citation count in the replication year to the standardized value of ‘100’ and then adjusting the counts in other years according to the same transformation ratio.
-For example, if the raw citation count in the replication year was 1000, citation counts in each year were standardized by dividing by 10.
-This computation was performed separately for the reference class and citations to the original article.
+For example, if the raw citation count in the replication year was 1000, citation counts in each year would be standardized by dividing by 10.
+This computation was performed separately for citations to the reference class and citations to the original article.
 
@@
 
-To examine the belief correction/perpetuation pattern (Table \@ref(tab:patterns)), the primary coder manually extracted the ‘citation context’ of the original study and the replication study (i.e., all relevant verbatim text surrounding each in-text citation).
+To examine the belief correction/perpetuation pattern (Table \@ref(tab:patterns)), one of six primary coders (TEH, DS, RTT, SC, ORA, and MN) manually extracted the ‘citation context’ of the original study and the replication study (i.e., all relevant verbatim text surrounding each in-text citation).
 The primary coder then classified the citation valence as ‘favourable’, ‘equivocal’, ‘unfavourable’, or ‘unclassifiable’.
@@
 Citations that did not endorse or oppose the phenomenon of interest (for example, simply referring to the procedures of the original study) were designated as ‘unclassifiable’.
-Because this process was inherently subjective, the citation contexts and classifications were also examined by a secondary coder.
+Because this process was inherently subjective, the citation contexts and classifications were also examined by one of six secondary coders (TEH, DS, RTT, SC, ORA, and MN).
 Disagreements were resolved through discussion and a third coder arbitrated when necessary.
@@
 
-To examine the explicit/absent defence pattern (Table \@ref(tab:patterns)), the primary coder flagged articles that co-cited the original and replication studies that also contained any explicit defence of the original study.
+To examine the explicit/absent defence pattern (Table \@ref(tab:patterns)), the primary coder flagged articles that co-cited the original and replication studies and also contained any explicit defence of the original study.
 Subsequently, two team members (ORA and SC) re-examined all of the flagged cases, extracted verbatim counter-arguments, and developed a post-hoc categorisation scheme that summarised them as concisely and informatively as possible (Table \@ref(tab:caCategories)).
@@
 These author records are automatically generated by an algorithm that attempts to identify all documents likely published by an individual author using several variations of their name (for example, “Hardwicke, Tom E.”, “Hardwicke, Tom”,  “Hardwicke, T.
-E.”), but errors can still occur and incomplete database coverage means that this method likely misses some of the authors’ prior publications and thus some of their collaborators.
+E.”), but errors can still occur and incomplete database coverage means that this method likely misses some of the authors’ prior publications and consequently some of their collaborators.
 Nevertheless, the method supports a reasonable lower-bound estimate of authorship overlap with articles providing counter-arguments.
@@
 When authors cited the original study favourably despite the replication result, only half of the articles provided any explicit counter arguments in defence of the original study (an explicit defence).
-Overall, these findings are consistent with prior observations that favourable citation patterns can continue relatively unperturbed by subsequent publication of contradictory results in studies with lower risk of bias (Tatsioni et al., 2007) or even by full retraction (Budd et al., 1998), and that positive (supportive) evidence is preferentially cited relative to negative (non-supportive) evidence once a theory gets entrenched despite overwhelming evidence against it (Greenberg, 2009).
+Overall, these findings are consistent with prior observations that favourable citation patterns appear relatively unperturbed by subsequent publication of contradictory results in studies with lower risk of bias (Tatsioni et al., 2007) or even by full retraction (Budd et al., 1998), and that positive (supportive) evidence is preferentially cited relative to negative (non-supportive) evidence once a theory gets entrenched despite overwhelming evidence against it (Bastiaansen et al., 2015; Greenberg, 2009).
 
@@
 One could imagine a similar feature being introduced for replication studies, perhaps based on databases that explicitly identify replication studies such as Curate Science (https://curatescience.org/).
-However, it is much less straightforward to define a relevant replication study (Neuliep & Crandall, 1993) and users would need to be alerted that this requires some scientific judgement rather than simple article meta-data.
+However, it is much less straightforward to define and identify a relevant replication study (Neuliep & Crandall, 1993) and users would need to be alerted that this requires some scientific judgement rather than simple article meta-data.
 Another solution could be to encourage researchers to focus less on individual studies and more on up-to-date evidence summaries (i.e., reviews and meta-analyses) in which relevant evidence is systematically identified and collated.
-This would require that high-quality and contemporary evidence summaries are available; however, in psychology systematic reviews and meta-analyses can be of low quality and their results may still be inflated and irreproducible (Kvarven et al., 2020; Maassen et al., 2020; Polanin et al., 2020).
-Moreover, most empirical studies are not included in any form of evidence synthesis (Hardwicke et al., 2020).
+This would require that high-quality and contemporary evidence summaries are available; however, in psychology systematic reviews and meta-analyses can be of low quality and their results may still be inflated and non-reproducible (Kvarven et al., 2020; Maassen et al., 2020; Polanin et al., 2020).
+Moreover, empirical studies are often not included in any form of evidence synthesis (Hardwicke et al., 2020).
 
@@
 In fact, we found that when authors continued to favourably cite original studies despite the replication results, around half did not provide explicit counter-arguments in defence of the original study.
-Although our study did not examine researchers’ individual beliefs, one recent study found that when research psychologists were confronted with replication evidence, they often did update their (self-reported) beliefs, albeit modestly (McDiarmid et al., 2021).
+Although our study did not examine researchers’ individual beliefs, one recent study reported that when research psychologists were confronted with replication evidence, they often did update their (self-reported) beliefs, albeit modestly (McDiarmid et al., 2021).
 However, there are several reasons to be uncertain about whether the results from this artificial setting might generalize to real-world settings, including potential observer bias (i.e., participants behaving differently because they are under observation and/or responding to the perceived expectations of the research team) and the possibility that individuals may behave differently in settings where they have substantial personal investment and may be publicly scrutinized.
@@
 
-The findings presented here are inherently limited by the observational nature of the study design, which complicates straightforward conclusions about the causal impact of the replications.
-Although the use of a reference class enables us to detect the influence of exogenous factors to some extent, we cannot rule out their contribution.
-For example, the modest decline in favourable citations observed in most cases could be attributable, at least in part, to a more general awareness in the research community about methodological issues (for example, that the sample sizes of the original studies may not have provided sufficient statistical power).
-We have also focused only on the replication study and the original study in each case study, without considering the impact of other potentially relevant events.
-Notably, meta-research studies have detected signatures of publication bias and other questionable research practices in the fields to which these case studies belong (E.
-C.
-Carter et al., 2015; Coles et al., 2019; Vadillo, 2019; Vadillo et al., 2016) and other relevant replication studies contesting prior findings have been published (e.g., Rohrer et al., 2015).
-
 We have only been able to gauge reactions to replications to the extent that they are reflected in citation patterns.
-Thus, for example, our method would not detect if there had been a self-correction effect amongst individuals who would not typically cite the original study, such as students, members of the public, or researchers working in other fields.
+Thus, for example, our method would not detect if there had been a correction effect amongst individuals who would not typically cite the original study, such as students, members of the public, or researchers working in other fields.
 A contested study may continue to be cited favourably by its proponents who remain working in the field.
@@
 Also of note, we deliberately selected case studies where the replication studies were high-profile and had yielded high credibility evidence that strongly contradicted and outweighed the original findings.
-A self-correction effect may be less expected in cases where replication results are more ambiguous, less consequential, or less well-known.
+A correction effect may be less expected in cases where replication results are more ambiguous, less consequential, or less well-known.
 For example, in a situation where two high credibility studies with similar evidential value yield contradictory results, it would be premature to lose confidence in one of the studies before further investigation has probed the cause of the discrepancy.
@@
 We should also highlight that particular aspects of our study were inherently subjective, specifically, the identification of citation context, the classification of citation valence, and the identification, extraction, and categorisation of counter-arguments.
-In order to minimize subjectivity, a team of six investigators performed all coding in duplicate, with a third investigator arbitrating if necessary.
+In order to minimize subjectivity, a team of six investigators performed coding in duplicate, with a third investigator arbitrating if necessary.
 As disagreements between primary and secondary coders were infrequent, we are confident that the classifications are meaningful, but there may be some edge cases when an independent observer might reasonably disagree with our classifications.
@@
 In conclusion, post-replication citation patterns in four case studies indicated that the anticipated corrective impact of strongly contradictory replication results did not materialise to any substantive degree.
-A lack of awareness of replications and/or a decision to discount them appears to have played a significant role.
+A lack of awareness of replications and/or a decision to discount or omit them appears to have played a significant role.
 This highlights potential practical problems with the discoverability of replication studies and psychological or sociological issues related to belief change.

### 2021-02-07 manuscript/manuscript.Rmd
finalizing alignment of google doc and r markdown

@@
   - id            : "8"
-    institution   : "Departments of Medicine, of Health Research and Policy, of Biomedical Data Science, and of Statistics, Stanford University"
+    institution   : "Departments of Medicine, of Epidemiology and Population Health, of Biomedical Data Science, and of Statistics, Stanford University"
 
@@
 Of these `r d_contentAnalysis %>% nrow()` articles, we excluded `r excluded_n` from the qualitative analysis because (1) we could not access the full text (n = `r noAccess_n`); (2) they were non-English language (n = `r nonEnglish_n`); (3) they included a citation to the original study in the reference section, but not in the main text (n = `r refCiteOnly_n`); or (4) manual inspection indicated that they did not actually appear to cite the original study at all (n = `r noCite_n`).
-Research design classifications for the remaining `r notexcluded_n` articles included in the qualitative analysis are shown in Table \@ref(tab:articleTypes).
+Article type classifications for the remaining `r notexcluded_n` articles included in the qualitative analysis are shown in Table \@ref(tab:articleTypes).
 
@@
   select("Article type" = articleType, "Count (%)", -n) %>%
-  apa_table(caption = "Counts and percentages for research design classifications of articles included in qualitative analyses")
+  apa_table(caption = "Counts and percentages for article type classifications of articles included in qualitative analyses")
 ```
@@
 
+The findings presented here are inherently limited by the observational nature of the study design, which complicates straightforward conclusions about the causal impact of the replications.
+Although the use of a reference class enables us to detect the influence of exogenous factors to some extent, we cannot rule out their contribution.
+For example, the modest decline in favourable citations observed in most cases could be attributable, at least in part, to a more general awareness in the research community about methodological issues (for example, that the sample sizes of the original studies may not have provided adequate statistical power).
+We have also focused only on the replication study and the original study in each case study, without considering the impact of other potentially relevant events.
+Notably, meta-research studies have detected signatures of publication bias and other questionable research practices in the fields to which these case studies belong (E.
+C.
+Carter et al., 2015; Coles et al., 2019; Vadillo, 2019; Vadillo et al., 2016) and other relevant replication studies contesting prior findings have been published (e.g., Rohrer et al., 2015).
+
 We have only been able to gauge reactions to replications to the extent that they are reflected in citation patterns.

### 2021-04-27 manuscript/appendix.Rmd
added raw citation curves to appendix

@@
 
-# Tabular data for citation curves, citation valence, and co-citation of replications {#appC}
+# Non-standardized data for citation curves, citation valence, and co-citation of replications {#appC}
 
@@
 ```
+
+```{r citeCurvesRaw, fig.cap='Annual citation counts (non-standardized; solid line) for the five original studies.
+The black arrow indicates the year in which the replication was published.', fig.width=12, fig.height=12.5, fig.path='figs/', dev=c('png', 'pdf')}
+ggarrange(
+  ggarrange(citationCurve("strack", thisTitle = "Strack et al. (1988)", standardized = F, plotReference = F, areaPlot = F),
+    citationCurve("baumeister", thisTitle = "Baumeister et al. (1998)", standardized = F, plotReference = F, areaPlot = F),
+    nrow = 2, ncol = 1, common.legend = T
+  ),
+  ggarrange(citationCurve("sripada", thisTitle = "Sripada et al. (2014)", standardized = F, plotReference = F, areaPlot = F),
+    citationCurve("carter", thisTitle = "Carter et al. (2011)", standardized = F, plotReference = F, areaPlot = F),
+    citationCurve("caruso", thisTitle = "Caruso et al. (2013)", standardized = F, plotReference = F, areaPlot = F),
+    nrow = 1, ncol = 3, common.legend = F
+  ),
+  nrow = 2, ncol = 1, heights = c(2, 1)
+) %>%
+  annotate_figure(
+    left = text_grob("Citation count (non-standardized)", rot = 90, size = 20),
+    bottom = text_grob("Publication year", size = 20)
+  )
+```

### 2021-04-27 analysis/functions.R
adjust curves func to allow raw curve plotting

@@ -286,7 +286,7 @@ citationCurve <- function(
     pull(pubYear)
 
   # set the top of the y-axis and increments depending on case (some have many citations, some have few)
-  if (standardized == T) { # if standardizing then all cases can use same y axis (except for Caruso when we standardize to the replication year...)
+  if (standardized == T) { # if standardizing then all cases can use same y axis (except for Carter and Caruso)
     if (thisCase == "caruso") {
       yMax <- 300
       yInc <- 50
@@ -297,12 +297,12 @@ citationCurve <- function(
       yMax <- 125
       yInc <- 25
     }
-  } else if (thisCase == "baumeister") {
+  } else if (thisCase == "baumeister") { # if not standardizing then need to set axes for each case individually
     yMax <- 250
     yInc <- 50
   } else if (thisCase == "strack") {
-    yMax <- 70
-    yInc <- 10
+    yMax <- 75
+    yInc <- 25
   } else {
     yMax <- 20
     yInc <- 5
@@ -339,6 +339,7 @@ citationCurve <- function(
       excluded_yes_std,
       citesTarget,
       citesTarget_std,
+      citesRef,
       citesRef_std,
       citesRep_yes_std,
       citesRep_no_std,
@@ -352,6 +353,7 @@ citationCurve <- function(
         excluded_yes_std,
         citesTarget,
         citesTarget_std,
+        citesRef,
         citesRef_std,
         citesRep_yes_std,
         citesRep_no_std,
@@ -367,6 +369,7 @@ citationCurve <- function(
         "Excluded" = "excluded_yes_std",
         "Target study (raw)" = "citesTarget",
         "Target study" = "citesTarget_std",
+        "Reference studies (raw)" = "citesRef",
         "Reference studies" = "citesRef_std",
         "Cites replication" = "citesRep_yes_std",
         "Does not cite replication" = "citesRep_no_std",
@@ -380,6 +383,7 @@ citationCurve <- function(
         "Excluded",
         "Target study (raw)",
         "Target study",
+        "Reference studies (raw)",
         "Reference studies",
         "Cites replication",
         "Does not cite replication",
@@ -428,17 +432,31 @@ citationCurve <- function(
   ## end accounting for absence of qual data in rep year
 
   if (plotReference == T) { # if user wants to plot the reference class
-    basePlot <- basePlot +
-      geom_line(
-        data = d %>% # plot reference class citations
-          filter(
-            var == "Reference studies",
-            dummyYears == F
[…]

### 2021-04-27 analysis/preprocessing.R
Update preprocessing.R

@@ -125,6 +125,3 @@ d_repCitations <- d_repCitations %>%
 ### Save file
 write_csv(d_repCitations, file = here("data", "processed", "/d_repCitations.csv"))
 save(d_repCitations, file = here("data", "processed", "/d_repCitations.rds"))
-
-
-

### 2021-04-27 manuscript/appendix.Rmd
round 1 revisions added

@@
 tibble(
-  "Issue" = c("Sampling frame", "Definition of pre-replication assessment period", "Exclusion of replication year from qualitative assessment", "Automated extraction of citation contexts and replication co-citation", "Counter-arguments", "Dual-coding for citation valence classification", "Examination of characteristics of articles providing counter-arguments.","Renaming of two citation patterns", "Data exclusions"),
+  "Issue" = c("Sampling frame", "Definition of pre-replication assessment period", "Exclusion of replication year from qualitative assessment", "Automated extraction of citation contexts and replication co-citation", "Counter-arguments", "Dual-coding for citation valence classification", "Examination of characteristics of articles providing counter-arguments.","Renaming of two citation patterns", "Data exclusions", "Citations to replication studies and co-citation of original studies"),
   "Amendment" = c("The end date of the sampling frame was initially April 6th 2018 (see pre-registered protocol).
@@
 These changes were documented in the amended protocol.","In the amended protocol, we clarified that we would attempt to identify counter-arguments in all articles that cited the original study and the replication study.", "Initially, we planned to dual code small batches of articles depending on the degree of inter-rater reliability.
-However, after encountering a number of somewhat ambiguous cases in our initial coding, we decided to dual-code all articles in order to reduce subjectivity and increase consistency between coders.
+However, we encountered a number of somewhat ambiguous cases in our initial coding which suggested there may be more subjectivity involved in the classification of citation valence than we had originally anticipated.
+For example, an article may contain several favourable statements and several unfavourable statements, in which case one needs to determine how to appropriately weight those statements in order to arrive at a classification that accurately represents the author’s viewpoint.
+To address this, we decided to minimize subjectivity by dual-coding all articles, with disagreements resolved through discussion (with a third coder if necessary).
 This is documented in the amended protocol.", "After completing the study we thought it would be informative to examine the article types and journals of articles that provided counter-arguments, and any overlap in authorship with the original study authors or prior collaborators of the first authors of the original studies.
 These analyses were not pre-registered.", "After completing the study we decided to revise the following terminology from the protocol in order to improve clarity: “principled defence” was changed to “explicit defence” and “hollow defence” was changed to “absent defence”.", "The criteria to exclude data from qualitative assessment (no access, non-English language, and no original study citation in the main text, were not pre-registered.
-These exclusions were necessary because they precluded the qualitative assessment.")
+These exclusions were necessary because they precluded the qualitative assessment.","Upon reviewer request, we obtain citation counts for replication studies and checked if they co-cited the relevant original study.
+These results are presented in Supplementary Information D.")
 ) %>%
@@
 
-```{r}
-# data summarised for post-replication period
-d_summary_postrep <- d_summary %>%
-  filter(year == 'post-replication') %>%
-  mutate(favourable = paste0(favourable,' (',round(favourable_prop,2),')'),
-         unfavourable = paste0(unfavourable,' (',round(unfavourable_prop,2),')'),
-         equivocal = paste0(equivocal,' (',round(equivocal_prop,2),')'),
-         unclassifiable = paste0(unclassifiable,' (',round(unclassifiable_prop,2),')'),
-         excluded = paste0(excluded_yes,' (',round(excluded_yes_prop,2),')'),
-         citesRep = paste0(citesRep_yes,' (',round(citesRep_yes_prop,2),')')) %>%
-  select(case, year = pubYear, `citing\narticles` = citesTarget, excluded, unclassifiable, favourable, equivocal, unfavourable, citesRep)
-```
-
 ```{r citeCurvesRaw, fig.cap='Annual citation counts (non-standardized; solid line) for the five original studies.
@@
 ```
+
+```{r}
+# data summarised for post-replication period
+d_summary_postrep <- d_summary %>%
+  filter(year == 'post-replication') %>%
+  mutate(favourable = paste0(favourable,' (',round(favourable_prop,2),')'),
+         unfavourable = paste0(unfavourable,' (',round(unfavourable_prop,2),')'),
+         equivocal = paste0(equivocal,' (',round(equivocal_prop,2),')'),
+         unclassifiable = paste0(unclassifiable,' (',round(unclassifiable_prop,2),')'),
+         excluded = paste0(excluded_yes,' (',round(excluded_yes_prop,2),')'),
+         citesRep = paste0(citesRep_yes,' (',round(citesRep_yes_prop,2),')')) %>%
+  select(case, year = pubYear, `citing\narticles` = citesTarget, excluded, unclassifiable, favourable, equivocal, unfavourable, citesRep)
+```
+
+# Citations to replication studies and co-citation of original studies {#appD}
+
+Upon reviewer request, we examined citation counts for replication studies and checked whether citing articles also co-cited the relevant original study.
+To obtain the data, we downloaded bibliographic records from the Clarivate Analytics Web of Science Core Collection accessed via the University of Amsterdam on 16th April, 2021 for articles that cited each replication study (up to the study endpoint - December, 2019) and cross-checked them with our sample of articles that cited the original study.
+As shown, the replication studies also have a life of their own and they are often cited independently of the specific original study.
+Often this is easy to explain, e.g., Klein et al.
+replicated 13 original studies, not just the two that were of interest in our analysis.
+These studies are also likely to have accrued citations by virtue of being amongst the first highly prominent examples of preregistered multi-laboratory replications in psychology.
+
+```{r}
+# identify how many studies citing the replication also cite the original study
+
+## extract citing articles for replication studies and original studies
+d_kleinCites <- d_repCitations %>% filter(case == 'klein')
+d_wagenmakersCites <- d_repCitations %>% filter(case == 'wagenmakers')
+d_haggerCites <- d_repCitations %>% filter(case == 'hagger')
+d_baumeisterCites <- d_citations %>% filter(case == 'baumeister')
+d_sripadaCites <- d_citations %>% filter(case == 'sripada')
+d_carusoCites <- d_citations %>% filter(case == 'caruso')
+d_carterCites <- d_citations %>% filter(case == 'carter')
+d_strackCites <- d_citations %>% filter(case == 'strack')
+
+## compute citation counts for replication studies
+hagCites <- nrow(d_haggerCites)
+wagenCites <- nrow(d_wagenmakersCites)
+kleinCites <- nrow(d_kleinCites)
+
+## compute how many articles citing the replication study also cite the relevant original study
+hagBauCocites <- sum(d_haggerCites$UT %in% d_baumeisterCites$UT)
+hagSriCocites <- sum(d_haggerCites$UT %in% d_sripadaCites$UT)
+wagenStrackCocites <- sum(d_wagenmakersCites$UT %in% d_strackCites$UT)
+kleinCartCocites <- sum(d_kleinCites$UT %in% d_carterCites$UT)
+kleinCaruCocites <- sum(d_kleinCites$UT %in% d_carusoCites$UT)
+
+## build table
+repCociteTable <- data.frame(
+  "Replication" = c("Hagger et al. (2016)", "Klein et al. (2014)", "Wagenmakers et al. (2016)"),
+  "Citation" = c(hagCites, kleinCites, wagenCites),
+  "Cocitations" = c(paste0(hagBauCocites," (Baumeister et al.
+1998) ",hagSriCocites," (Sripada et al.
+2014)"), paste0(kleinCartCocites," (Carter et al.
+2011) ",kleinCaruCocites," (Caruso et al.
+2012)"), paste0(wagenStrackCocites," (Strack et al.
+1988)"))
+  )
+```
+
+```{r}
+repCociteTable %>% kable(col.names = c("Replication study", "Citation count","Cocitations of original study"), caption = 'Citations counts for replication studies and co-citation counts to relevant original studies.', booktabs=T)
+```

### 2021-04-27 manuscript/manuscript.Rmd
round 1 revisions added

@@
 ---
-title             : "Post-replication citation patterns in psychology: Four case studies"
+title             : "Citation patterns following a strongly contradictory replication result: Four case studies from psychology"
 shorttitle        : "Post-replication citation patterns"
@@
 Prior research has documented how favourable citations to observational epidemiology studies can persist despite the claims of those studies being strongly contradicted in subsequent randomized trials (Tatsioni et al., 2007).
-Similarly, it has been reported that even when articles are retracted, they can continue to receive favourable citations (Budd et al., 1998).
+Similarly, it has been reported that even when articles are retracted, they can continue to receive favourable citations (Budd et al., 1998; Fernandez & Vadillo, 2020).
 Thus, there is evidence that belief in the credibility of original findings can perpetuate even when subsequent events cast doubt on their credibility; however, we are unaware of similar evidence in the context of replication studies that were explicitly designed to test the reliability of prior findings.
@@
 For methodological reasons (see Hagger et al., 2016), the ego-depletion replication was aimed at a classic study in the field (Baumeister et al., 1998), but actually employed a modified computer-based version of the original paradigm (Sripada et al., 2014).
-For this particular case study we examined citation patterns to both of these original studies.
+For this particular case study, we examined citation patterns to both of these original studies.
+
+We adopted a case study approach in order to develop a ‘narrow and deep’ understanding of the topic, as opposed to ‘broad and shallow’ approach, which would have required a deliberate representative sampling strategy.
 We chose these particular case studies because they involved prominent pre-registered multi-laboratory replication attempts with sample sizes between 23 - 211 times larger than the original studies, thus providing highly visible and highly credible evidence that strongly contradicted and outweighed earlier findings[^footnote3].
@@
 
-[^footnote3]: We deliberately focused on a select group of case studies rather than other potentially larger samples in order to aid interpretative clarity; for example, the extent to which the results of the large scale Reproducibility Project in Psychology (RPP; Open Science Collaboration, 2015) actually contradicted the original studies has been contested (Etz & Vandekerckhove, 2016).
+[^footnote3]: We deliberately focused on a select group of case studies rather than other potentially larger samples in order to aid interpretative clarity; for example, the extent to which the results of the large-scale Reproducibility Project in Psychology (RPP; Open Science Collaboration, 2015) actually contradicted the original studies has been contested (Etz & Vandekerckhove, 2016).
 Additionally, original studies were not actually cited in the RPP research report, which may have diluted awareness about relevant replication studies.
@@
 We examined four case studies in which a prominent pre-registered and multi-laboratory replication study strongly contradicted and outweighed the findings of an original study (Table \@ref(tab:studyDetails)).
+As shown, all the original studies found modestly large to very large effects and all of them were relatively small studies (thus they would have been underpowered to detect small effects).
+Conversely, all the replication efforts comprised of very large sample sizes, and they would be very well powered to detect even small effects; however, they all obtained null results.
 
@@
 In this study, we investigated how the research community responded to four strongly contradictory replications in the field of psychology by examining post-replication citation patterns for original studies.
-We observed some ‘progressive’ response in the form of modest active correction patterns (a small decline in favourable citations and a small increase in unfavourable citations), but also even more prominent ‘regressive’ responses in the form of unchallenged belief perpetuation (sustained levels of favourable citations and few unfavourable citations, particularly in the Baumeister case) and considerable citation bias (neglecting to cite the replication study; in all cases aside from Sripada).
+We observed some ‘progressive’ response patterns in the form of modest active correction patterns (a small decline in favourable citations and a small increase in unfavourable citations), but also even more prominent ‘regressive’ response patternsß in the form of unchallenged belief perpetuation (sustained levels of favourable citations and few unfavourable citations, particularly in the Baumeister case) and considerable citation bias (neglecting to cite the replication study; in all cases aside from Sripada).
 When authors cited the original study favourably despite the replication result, only half of the articles provided any explicit counter arguments in defence of the original study (an explicit defence).
-Overall, these findings are consistent with prior observations that favourable citation patterns appear relatively unperturbed by subsequent publication of contradictory results in studies with lower risk of bias (Tatsioni et al., 2007) or even by full retraction (Budd et al., 1998), and that positive (supportive) evidence is preferentially cited relative to negative (non-supportive) evidence once a theory gets entrenched despite overwhelming evidence against it (Bastiaansen et al., 2015; Greenberg, 2009).
+Overall, these findings are consistent with prior observations that favourable citation patterns appear relatively unperturbed by subsequent publication of contradictory results in studies with lower risk of bias (Tatsioni et al., 2007) or even by full retraction (Budd et al., 1998; Fernandez & Vadillo, 2020), and that positive (supportive) evidence is preferentially cited relative to negative (non-supportive) evidence once a theory gets entrenched, despite overwhelming evidence against it (Bastiaansen et al., 2015; Greenberg, 2009).
 
@@
 Thus, it seems reasonable to suggest that the compelling replication results should have reduced belief in the credibility of the original results.
+However, one could also contend that if there were some flaw in the replication study that undermined its validity (Brandt et al., 2014; Vazire et al., 2021; Fabrigar et al., 2020), then this may provide justification to continue favourably citing the original study.
+Indeed, in these particular case studies, the validity of the replication results has been challenged by proponents of the original findings (Table 5; Baumeister & Vohs, 2016; Strack, 2016).
+As this debate remains far from settled (Coles et al., 2019; Vadillo, 2019), ideally any favourable citation of the original studies should at a minimum be accompanied by co-citation of the replication results and some discussion of the discrepant findings.
 
@@
 Recently, the reference manager Zotero has introduced a feature that alerts users when an article in their database has been retracted (Zotero, 2019).
-One could imagine a similar feature being introduced for replication studies, perhaps based on databases that explicitly identify replication studies such as Curate Science (https://curatescience.org/).
+One could imagine a similar feature being introduced for replication studies, perhaps based on databases that explicitly identify replication studies.
 However, it is much less straightforward to define and identify a relevant replication study (Neuliep & Crandall, 1993) and users would need to be alerted that this requires some scientific judgement rather than simple article meta-data.
@@
 
-The second issue of authors ignoring highly relevant replications is undesirable and implies a biased appraisal of the evidence.
+The second issue of authors ignoring highly relevant replications seems undesirable and implies a biased appraisal or presentation of the evidence.
 Contradictory replication results do not necessarily nullify the credibility of an original study (Collins, 1985; Earp & Trafimow, 2015; Maxwell et al., 2015), but we would still expect highly relevant replication results to be cited and explicitly debated.
@@
 Although our study did not examine researchers’ individual beliefs, one recent study reported that when research psychologists were confronted with replication evidence, they often did update their (self-reported) beliefs, albeit modestly (McDiarmid et al., 2021).
-However, there are several reasons to be uncertain about whether the results from this artificial setting might generalize to real-world settings, including potential observer bias (i.e., participants behaving differently because they are under observation and/or responding to the perceived expectations of the research team) and the possibility that individuals may behave differently in settings where they have substantial personal investment and may be publicly scrutinized.
+However, there are several reasons to be uncertain about whether the results from this artificial setting might generalize to real-world settings, including potential participant reactivity (i.e., participants behaving differently because they are under observation and/or responding to the perceived expectations of the research team) and the possibility that individuals may behave differently in settings where they have substantial personal investment and may be publicly scrutinized.
 Additionally, cognitive psychology studies have obtained some evidence of a ‘continued influence effect’ wherein an individual's beliefs and behaviour can continue to be influenced by false or misleading information despite subsequent efforts to reject it (Lewandowsky et al., 2012).
@@
 
-We have only been able to gauge reactions to replications to the extent that they are reflected in citation patterns.
-Thus, for example, our method would not detect if there had been a correction effect amongst individuals who would not typically cite the original study, such as students, members of the public, or researchers working in other fields.
+We have only been able to gauge reactions to replications to the extent that they are reflected in relatively short-term citation patterns.
+A potential explanation of the apparently cursory treatment of the replication studies could be that researchers only became aware of them after their own research projects had begun and or even had been completed.
+If one of the original studies had been a key motivator for one’s own study, then it may be difficult to accommodate the strongly contradictory replication results.
+It may even be tempting to ignore them, or give them only superficial treatment.
+Moreover, researchers who are convinced that the replication study has squarely refuted the original, may no longer be interested in doing research on a topic where they see no future potential.
+Additionally, examination of citation patterns would not detect if there had been a correction effect amongst individuals who would not typically cite the original study, such as students, members of the public, or researchers working in other fields.
 A contested study may continue to be cited favourably by its proponents who remain working in the field.
@@
 Thus, whilst the current findings may contradict the expectations of a more direct and expedited view of the corrective impact of replication studies (Ioannidis, 2012; Vazire & Holcombe, 2020), they are not necessarily inconsistent with a slower and more indirect process of self-correction.
+Future research could employ longitudinal designs or older historical case studies in order to evaluate citation patterns unfolding over a longer time scale.
 
@@
 Baumeister, R.
-E., Bratslavsky, E., Muraven, M., & Tice, D.
+F., Bratslavsky, E., Muraven, M., & Tice, D.
 M.
@@
 
+Baumeister, R. F., & Vohs, K. D. (2016). Misguided effort with elusive implications. Perspectives on Psychological Science, 11(4), 574-575. https://doi.org/10.1177%2F1745691616652878
+
 Bishop, D.
@@
 https://doi.org/10.1177/1747021819886519
+
+Brandt, M.
+J., IJzerman, H., Dijksterhuis, A., Farach, F.
+J., Geller, J., Giner-Sorolla, R., Grange, J.
+A., Perugini, M., Spies, J.
+R., van ’t Veer, A.
+(2014).
+The Replication Recipe: What makes for a convincing replication?
+Journal of Experimental Social Psychology, 50, 217–24.
+https://doi.org/10.1016/j.jesp.2013.10.005.
 
@@
 
+Fabrigar, L.
+R., Wegener, D.
+T., Petty, R.
+E.
+(2020).
+A validity-based framework for understanding replication in psychology.
+Personality and Social Psychology Review, 24(4), 316-344.
+https://doi.org/10.1177/1088868320931366.
+
+Fernandez, L. M., & Vadillo, M. A. (2020). Retracted papers die hard: Diederik Stapel and the enduring influence of flawed science. PsyArXiv. https://doi.org/10.31234/osf.io/cszpy
+
 Gershman, S. J. (2019). How to never be wrong. Psychonomic Bulletin & Review, 26(1), 13–28. https://doi.org/10.3758/s13423-018-1488-8
@@
 
-Hagger, M. S., Chatzisarantis, N. L. D., Alberts, H., Anggono, C. O., Batailler, C., Birt, A. R., Brand, R., Brandt, M. J., Brewer, G., Bruyneel, S., Calvillo, D. P., 
-Campbell, W.
+Hagger, M.
+S., Chatzisarantis, N.
+L.
+D., Alberts, H., Anggono, C.
+O., Batailler, C., Birt, A.
+R., Brand, R., Brandt, M.
+J., Brewer, G., Bruyneel, S., Calvillo, D.
+P., Campbell, W.
 K., Cannon, P.
@@
 Perspectives on Psychological Science.
-https://doi.org/10.31222/osf.io/9sz2y
+https://doi.org/10.1177%2F1745691620979806
 
@@
 https://doi.org/10.1177/0956797614526415
+
+Strack, F. (2016). Reflection on the smiling registered replication report. Perspectives on Psychological Science, 11(6), 929–930. https://doi.org/10.1177/1745691616674460
 

### 2021-04-27 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
 
-We adopted a case study approach in order to develop a ‘narrow and deep’ understanding of the topic, as opposed to ‘broad and shallow’ approach, which would have required a deliberate representative sampling strategy.
+We adopted a case study approach in order to develop a ‘narrow and deep’ understanding of the topic, as opposed to a ‘broad and shallow’ approach, which would have required a deliberate representative sampling strategy.
 We chose these particular case studies because they involved prominent pre-registered multi-laboratory replication attempts with sample sizes between 23 - 211 times larger than the original studies, thus providing highly visible and highly credible evidence that strongly contradicted and outweighed earlier findings[^footnote3].

### 2021-04-27 manuscript/manuscript.Rmd
updated code ocean link

@@
 All data, materials, and analysis scripts related to this study are publicly available on the Open Science Framework (https://osf.io/w8h2q/).
-To facilitate reproducibility this manuscript was written by interleaving regular prose and analysis code using knitr (Xie, 2017) and papaja (Aust & Barth, 2020), and is available in a Code Ocean container (https://doi.org/10.24433/CO.4225975.v1) which re-creates the software environment in which the original analyses were performed.
+To facilitate reproducibility this manuscript was written by interleaving regular prose and analysis code using knitr (Xie, 2017) and papaja (Aust & Barth, 2020), and is available in a Code Ocean container (https://doi.org/10.24433/CO.4225975.v2) which re-creates the software environment in which the original analyses were performed.
 

### 2021-07-14 analysis/preprocessing.R
Update preprocessing.R

@@ -19,7 +19,7 @@ d_citations <- d_citations %>%
   mutate(pubYear = as.numeric(PY))
 
 ### Save file
-write_csv(d_citations, path = here("data", "processed", "/d_citations.csv"))
+write_csv(d_citations, file = here("data", "processed", "/d_citations.csv"))
 save(d_citations, file = here("data", "processed", "/d_citations.rds"))
 
 ## Reference data
@@ -38,7 +38,7 @@ d_reference <- d_reference %>%
   mutate(pubYear = as.numeric(year))
 
 ### Save file
-write_csv(d_reference, path = here("data", "processed", "/d_reference.csv"))
+write_csv(d_reference, file = here("data", "processed", "/d_reference.csv"))
 save(d_reference, file = here("data", "processed", "/d_reference.rds"))
 
 ## Content analysis data
@@ -51,6 +51,9 @@ for (i in caseNames) { # loop through case names
   d <- d %>% mutate(case = i)
   d_contentAnalysis <- bind_rows(d_contentAnalysis, d) # append to dataframe
 }
+
+#### also load the articles that became accessible or were translated during round 2 revisions
+d_round2access <- read_csv(here("data", "primary", "contentAnalysis_round2access.csv"), col_types = cols(.default = "c")) # load the data
 
 ### Apply munging
 d_contentAnalysis <- d_contentAnalysis %>%
@@ -88,7 +91,7 @@ d_contentAnalysis <- d_contentAnalysis %>%
   ))
 
 ### Save file
-write_csv(d_contentAnalysis, path = here("data", "processed", "/d_contentAnalysis.csv"))
+write_csv(d_contentAnalysis, file = here("data", "processed", "/d_contentAnalysis.csv"))
 save(d_contentAnalysis, file = here("data", "processed", "/d_contentAnalysis.rds"))
 
 ## Coauthor data
@@ -103,7 +106,7 @@ for (i in caseNames) { # loop through case names
 }
 
 ### Save file
-write_csv(d_coauthors, path = here("data", "processed", "/d_coauthors.csv"))
+write_csv(d_coauthors, file = here("data", "processed", "/d_coauthors.csv"))
 save(d_coauthors, file = here("data", "processed", "/d_coauthors.rds"))
 
 ## Replication study citation data [analyses requested by a reviewer]

### 2021-07-14 analysis/preprocessing.R
adding accessible and translated articles
during round 2 of revisions at AMPPSS, reviewer requested additional effort to retrieve inaccessible articles and translate non-english article. Changes introduced here to incorporate new data after gaining access to and translating additional articles

@@ -1,5 +1,8 @@
 # Preprocessing
 
+# specify custom notin function
+`%notin%` = function(x,y) !(x %in% y)
+
 # identify the five different case studies we are looking at
 caseNames <- c("baumeister", "sripada", "strack", "carter", "caruso")
 
@@ -55,6 +58,21 @@ for (i in caseNames) { # loop through case names
 #### also load the articles that became accessible or were translated during round 2 revisions
 d_round2access <- read_csv(here("data", "primary", "contentAnalysis_round2access.csv"), col_types = cols(.default = "c")) # load the data
 
+#### save a version of the round 2 access data for reporting how many we could or could not access
+d_round2access_meta <- d_round2access %>% select(wos_id, access, source, emailContacted, emailBounced, emailResponseReceived, translationNeeded, translationNote)
+write_csv(d_round2access_meta, file = here("data", "processed", "/d_round2access.csv"))
+save(d_round2access_meta, file = here("data", "processed", "/d_round2access.rds"))
+
+#### integrate the newly accessible/translated article content analysis data with the previous data
+#### firstly align columns in the newly accessible articles df with the content analysis df
+d_round2access <- d_round2access %>% select(firstCoder, secondCoder, doi, authors,	pubYear,	excluded,	exclusionReason,	articleType,	citesReplication,	citationClassificationOriginal = citationClassificationFirstCoder, citationClassificationAgreed, counterArguments,	evidenceCounter,	evidenceVerbatim,	methodsCounter,	methodsVerbatim,	expertiseCounter,	expertiseVerbatim, UT =	wos_id,	case)
+
+#### remove the cases excluded for access or translation reasons from the content analysis dataframe and replace with those from the newly accessible articles dataframe
+
+d_contentAnalysis <- d_contentAnalysis %>% mutate(replace = ifelse(exclusionReason %in% c("No access", "Non-English language"), T, F)) # tag articles to replace
+d_contentAnalysis <- d_contentAnalysis %>% filter(replace == F) %>% select(-replace) # remove articles to be replaced
+d_contentAnalysis <- rbind(d_contentAnalysis,d_round2access) # combine newly accessible/translated articles with content analysis data
+
 ### Apply munging
 d_contentAnalysis <- d_contentAnalysis %>%
   mutate_at(vars( # change the columns below
@@ -82,12 +100,12 @@ d_contentAnalysis <- d_contentAnalysis %>%
 # homogenize article type labels
[…]

### 2021-07-14 manuscript/appendix.Rmd
updated after round 2 revisions

--- 2021-04-27 manuscript/appendix.Rmd
@@
 tibble(
-  "Issue" = c("Sampling frame", "Definition of pre-replication assessment period", "Exclusion of replication year from qualitative assessment", "Automated extraction of citation contexts and replication co-citation", "Counter-arguments", "Dual-coding for citation valence classification", "Examination of characteristics of articles providing counter-arguments.","Renaming of two citation patterns", "Data exclusions"),
+  "Issue" = c("Sampling frame", "Definition of pre-replication assessment period", "Exclusion of replication year from qualitative assessment", "Automated extraction of citation contexts and replication co-citation", "Counter-arguments", "Dual-coding for citation valence classification", "Examination of characteristics of articles providing counter-arguments.","Renaming of two citation patterns", "Data exclusions", "Citations to replication studies and co-citation of original studies", "Additional access attempts and translation"),
   "Amendment" = c("The end date of the sampling frame was initially April 6th 2018 (see pre-registered protocol).
@@
 These changes were documented in the amended protocol.","In the amended protocol, we clarified that we would attempt to identify counter-arguments in all articles that cited the original study and the replication study.", "Initially, we planned to dual code small batches of articles depending on the degree of inter-rater reliability.
-However, after encountering a number of somewhat ambiguous cases in our initial coding, we decided to dual-code all articles in order to reduce subjectivity and increase consistency between coders.
+However, we encountered a number of somewhat ambiguous cases in our initial coding which suggested there may be more subjectivity involved in the classification of citation valence than we had originally anticipated.
+For example, an article may contain several favourable statements and several unfavourable statements, in which case one needs to determine how to appropriately weight those statements in order to arrive at a classification that accurately represents the author’s viewpoint.
+To address this, we decided to minimize subjectivity by dual-coding all articles, with disagreements resolved through discussion (with a third coder if necessary).
 This is documented in the amended protocol.", "After completing the study we thought it would be informative to examine the article types and journals of articles that provided counter-arguments, and any overlap in authorship with the original study authors or prior collaborators of the first authors of the original studies.
-These analyses were not pre-registered.", "After completing the study we decided to revise the following terminology from the protocol in order to improve clarity: “principled defence” was changed to “explicit defence” and “hollow defence” was changed to “absent defence”.", "The criteria to exclude data from qualitative assessment (no access, non-English language, and no original study citation in the main text, were not pre-registered.
-These exclusions were necessary because they precluded the qualitative assessment.")
+These analyses were not pre-registered.", "After completing the study we decided to revise the following terminology from the protocol in order to improve clarity: “principled defence” was changed to “explicit defence” and “hollow defence” was changed to “absent defence”.", "The criteria to exclude data from qualitative assessment (no access and no original study citation in the main text, were not pre-registered.
+These exclusions were necessary because they precluded the qualitative assessment.","Upon reviewer request, we obtained citation counts for replication studies and checked if they co-cited the relevant original study.
+These results are presented in the main results section.", "Upon reviewer request, we took additional steps to obtain articles that we were previously unable to access, and to translate articles that were written in a non-English language.
+The newly accessible and translated articles then underwent qualitative analyses.
+These new data had minimal impact on the results, we have therefore incorporated them into the main paper.
+Supplementary Information D describes the process in detail and also contains copies of the original graphs before the data from the newly accessible and translated articles were incorporated.")
 ) %>%
@@
 ```
+
+```{r}
+# data summarised for post-replication period
+d_summary_postrep <- d_summary %>%
+  filter(year == 'post-replication') %>%
+  mutate(favourable = paste0(favourable,' (',round(favourable_prop,2),')'),
+         unfavourable = paste0(unfavourable,' (',round(unfavourable_prop,2),')'),
+         equivocal = paste0(equivocal,' (',round(equivocal_prop,2),')'),
+         unclassifiable = paste0(unclassifiable,' (',round(unclassifiable_prop,2),')'),
+         excluded = paste0(excluded_yes,' (',round(excluded_yes_prop,2),')'),
+         citesRep = paste0(citesRep_yes,' (',round(citesRep_yes_prop,2),')')) %>%
+  select(case, year = pubYear, `citing\narticles` = citesTarget, excluded, unclassifiable, favourable, equivocal, unfavourable, citesRep)
+```
+
+# Additional access attempts and translation {#appD}
+
+```{r}
+d_round2access <- read_csv(here('data','processed','d_round2access.csv'))
+```
+
+Initially, we excluded 64 of 632 articles from the qualitative analysis because (1) we could not access the full text via at least two of our institutional library affiliations (n = 58); or (2) they were non-English language (n = 6).
+Following requests from a reviewer, we took additional steps to access and translate articles (as described in the methods section).
+
+Our additional access attempts were successful for `r d_round2access %>% filter(access == T) %>% nrow()` articles.
+`r d_round2access %>% filter(access == T) %>% count(source) %>% filter(source == "UVA search") %>% pull(n)` were obtained via Institutional Library search, `r d_round2access %>% filter(access == T) %>% count(source) %>% filter(source == "general search") %>% pull(n)` were obtained via general Internet search, `r d_round2access %>% filter(access == T) %>% count(source) %>% filter(source == "email") %>% pull(n)` were obtained via author e-mail requests, and `r d_round2access %>% filter(access == T) %>% count(source) %>% filter(source == "email; inter-library loan") %>% pull(n)` was obtained via Interlibrary Loan from the University of Illinois.
+The `R d_round2access %>% filter(access == F) %>% nrow()` remaining articles could not be accessed despite trying all of these methods and contacting a librarian for assistance.
+
+`r d_round2access %>% filter(translationNeeded == T) %>% nrow()` articles were written in non-English language.
+We translated the relevant citation contexts to English.
+For `r d_round2access %>% filter(translationNote == "translated by SC from German (SC is native speaker)") %>% nrow()` articles written in German, translation was performed by one of the authors (SC) who is a native German speaker.
+For the other `r d_round2access %>% filter(translationNote != "translated by SC from German (SC is native speaker)") %>% nrow()` articles, written in Czech, Spanish, Korean, and Russian, we used Google Translate.
+
+The newly accessible and translated articles underwent qualitative analysis according to same procedures as other articles.
+The new data was incorporated into the main analysis, the results of which are presented in the main body of the paper.
+The new data had minimal impact on the results.

### 2021-07-14 manuscript/manuscript.Rmd
updated after round 2 revisions

@@
 lineno            : yes
+always_allow_html: true
 
@@
 output            : 
-  papaja::apa6_pdf:
-    includes:
-      after_body: "appendix.tex"
+  papaja::apa6_word
 ---
@@
 Similarly, it has been reported that even when articles are retracted, they can continue to receive favourable citations (Budd et al., 1998; Fernandez & Vadillo, 2020).
-Thus, there is evidence that belief in the credibility of original findings can perpetuate even when subsequent events cast doubt on their credibility; however, we are unaware of similar evidence in the context of replication studies that were explicitly designed to test the reliability of prior findings.
+Thus, there is evidence that belief in the credibility of original findings can perpetuate even when subsequent events cast doubt on their credibility; however, we are unaware of similar evidence in the context of studies that were explicitly designed to test the replicabiity of prior findings.
 
@@
 All deviations from these protocols are explicitly acknowledged in Appendix \@ref(appA).
-All data exclusions and measures conducted during this study are reported.
+We report how we determined our sample size, all data exclusions, all manipulations, and all measures in the study.
 
@@
 
-For each citing article undergoing qualitative assessment, we attempted to retrieve the full text via at least two of the institutional libraries we are affiliated with.
-Inaccessible articles were excluded.
+For each citing article undergoing qualitative assessment, we attempted to retrieve the full text via several methods in the following order: (1) search of at least two of the institutional libraries we are affiliated with; (2) general Internet search for the article title, including the Google and Google Scholar search engines and Research Gate; (3) e-mail requests to the corresponding author; (4) inter-library loan request.
+Articles that remained inaccessible after all of these methods were exhausted were excluded.
+Articles written in non-English language were translated by one of the authors or using Google Translate (see Supplementary Information D).
 For articles for which we could obtain the full text, we classified the research design according to the categories in Table \@ref(tab:articleTypes) and recorded whether the replication study was cited after manual inspection of the reference section (see Table \@ref(tab:patterns): citation balance/bias).
@@
 
+### Citations to replication studies and co-citation of original studies
+
+Upon reviewer request, we examined citation counts for replication studies and checked whether citing articles also co-cited the relevant original study.
+To obtain the data, we downloaded bibliographic records from the Clarivate Analytics Web of Science Core Collection accessed via the University of Amsterdam on 16th April, 2021 for articles that cited each replication study (up to the study endpoint - December, 2019) and cross-checked them with our sample of articles that cited the original study.
+As shown, the replication studies also have a life of their own and they are often cited independently of the specific original study.
+Often this is easy to explain, e.g., Klein et al.
+replicated 13 original studies, not just the two that were of interest in our analysis.
+These studies are also likely to have accrued citations by virtue of being amongst the first highly prominent examples of preregistered multi-laboratory replications in psychology.
+
+```{r}
+# identify how many studies citing the replication also cite the original study
+
+## extract citing articles for replication studies and original studies
+d_kleinCites <- d_repCitations %>% filter(case == 'klein')
+d_wagenmakersCites <- d_repCitations %>% filter(case == 'wagenmakers')
+d_haggerCites <- d_repCitations %>% filter(case == 'hagger')
+d_baumeisterCites <- d_citations %>% filter(case == 'baumeister')
+d_sripadaCites <- d_citations %>% filter(case == 'sripada')
+d_carusoCites <- d_citations %>% filter(case == 'caruso')
+d_carterCites <- d_citations %>% filter(case == 'carter')
+d_strackCites <- d_citations %>% filter(case == 'strack')
+
+## compute citation counts for replication studies
+hagCites <- nrow(d_haggerCites)
+wagenCites <- nrow(d_wagenmakersCites)
+kleinCites <- nrow(d_kleinCites)
+
+## compute how many articles citing the replication study also cite the relevant original study
+hagBauCocites <- sum(d_haggerCites$UT %in% d_baumeisterCites$UT)
+hagSriCocites <- sum(d_haggerCites$UT %in% d_sripadaCites$UT)
+wagenStrackCocites <- sum(d_wagenmakersCites$UT %in% d_strackCites$UT)
+kleinCartCocites <- sum(d_kleinCites$UT %in% d_carterCites$UT)
+kleinCaruCocites <- sum(d_kleinCites$UT %in% d_carusoCites$UT)
+
+## build table
+repCociteTable <- data.frame(
+  "Replication" = c("Hagger et al. (2016)", "Klein et al. (2014)", "Wagenmakers et al. (2016)"),
+  "Citation" = c(hagCites, kleinCites, wagenCites),
+  "Cocitations" = c(paste0(hagBauCocites," (Baumeister et al.
+1998) ",hagSriCocites," (Sripada et al.
+2014)"), paste0(kleinCartCocites," (Carter et al.
+2011) ",kleinCaruCocites," (Caruso et al.
+2012)"), paste0(wagenStrackCocites," (Strack et al.
+1988)"))
+  )
+```
+
+```{r}
+repCociteTable %>% kable(col.names = c("Replication study", "Citation count","Cocitations of original study"), caption = 'Citations counts for replication studies and co-citation counts to relevant original studies.', booktabs=T)
+```
+
 # Results
@@
 noAccess_n <- excluded_reason %>%
-  filter(exclusionReason == "No access") %>%
-  pull(n)
-nonEnglish_n <- excluded_reason %>%
-  filter(exclusionReason == "Non-English language") %>%
+  filter(exclusionReason == "noAccess") %>%
   pull(n)
@@
 In total, `r d_citations %>% nrow()` articles cited one of the original studies of which `r d_contentAnalysis %>% nrow()` articles (after taking a 40% random sample in the Baumeister case) fell within the time period designated for qualitative assessment.
-Of these `r d_contentAnalysis %>% nrow()` articles, we excluded `r excluded_n` from the qualitative analysis because (1) we could not access the full text (n = `r noAccess_n`); (2) they were non-English language (n = `r nonEnglish_n`); (3) they included a citation to the original study in the reference section, but not in the main text (n = `r refCiteOnly_n`); or (4) manual inspection indicated that they did not actually appear to cite the original study at all (n = `r noCite_n`).
+Of these `r d_contentAnalysis %>% nrow()` articles, we excluded `r excluded_n` from the qualitative analysis because (1) we could not access the full text (n = `r noAccess_n`); (2) they included a citation to the original study in the reference section, but not in the main text (n = `r refCiteOnly_n`); or (3) manual inspection indicated that they did not actually appear to cite the original study at all (n = `r noCite_n`).
 Article type classifications for the remaining `r notexcluded_n` articles included in the qualitative analysis are shown in Table \@ref(tab:articleTypes).
@@
 In this study, we investigated how the research community responded to four strongly contradictory replications in the field of psychology by examining post-replication citation patterns for original studies.
-We observed some ‘progressive’ response patterns in the form of modest active correction patterns (a small decline in favourable citations and a small increase in unfavourable citations), but also even more prominent ‘regressive’ response patternsß in the form of unchallenged belief perpetuation (sustained levels of favourable citations and few unfavourable citations, particularly in the Baumeister case) and considerable citation bias (neglecting to cite the replication study; in all cases aside from Sripada).
+We observed some ‘progressive’ response patterns in the form of modest active correction (a small decline in favourable citations and a small increase in unfavourable citations), but also even more prominent ‘regressive’ response patterns in the form of unchallenged belief perpetuation (sustained levels of favourable citations and few unfavourable citations, particularly in the Baumeister case) and considerable citation bias (neglecting to cite the replication study; in all cases aside from Sripada).
 When authors cited the original study favourably despite the replication result, only half of the articles provided any explicit counter arguments in defence of the original study (an explicit defence).
@@
 It is important to note that the case studies we selected were deliberately chosen because the replication results were superior to the original results in terms of both credibility and evidential value.
-The replications were pre-registered, muli-laboratory studies with large sample sizes.
+The replications were pre-registered, multi-laboratory studies with large sample sizes.
 By contrast, the original studies had much smaller sample sizes, and arguably had a much higher risk of bias, as they were not pre-registered, were performed by single teams, and arose in domains affected by publication bias and other questionable research practices (E.
@@
 Future research could employ longitudinal designs or older historical case studies in order to evaluate citation patterns unfolding over a longer time scale.
+Finally, as we only examined citation patterns, this study could not capture other potential responses to replication results, such as changes in research practices.
 
@@
 
-```{r render_appendix, include=FALSE}
-render_appendix("appendix.Rmd")
-```
+```{r child = "appendix.Rmd"}
+```

### 2021-07-19 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
 All data, materials, and analysis scripts related to this study are publicly available on the Open Science Framework (https://osf.io/w8h2q/).
-To facilitate reproducibility this manuscript was written by interleaving regular prose and analysis code using knitr (Xie, 2017) and papaja (Aust & Barth, 2020), and is available in a Code Ocean container (https://doi.org/10.24433/CO.4225975.v2) which re-creates the software environment in which the original analyses were performed.
+To facilitate reproducibility this manuscript was written by interleaving regular prose and analysis code using knitr (Xie, 2017) and papaja (Aust & Barth, 2020), and is available in a Code Ocean container (https://doi.org/10.24433/CO.4225975.v3) which re-creates the software environment in which the original analyses were performed.
 

### 2021-07-19 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
 output            : 
-  papaja::apa6_word
+  papaja::apa6_pdf:
+    includes:
+        after_body: "appendix.tex"
 ---
@@
 
-### Citations to replication studies and co-citation of original studies
-
-Upon reviewer request, we examined citation counts for replication studies and checked whether citing articles also co-cited the relevant original study.
-To obtain the data, we downloaded bibliographic records from the Clarivate Analytics Web of Science Core Collection accessed via the University of Amsterdam on 16th April, 2021 for articles that cited each replication study (up to the study endpoint - December, 2019) and cross-checked them with our sample of articles that cited the original study.
-As shown, the replication studies also have a life of their own and they are often cited independently of the specific original study.
-Often this is easy to explain, e.g., Klein et al.
-replicated 13 original studies, not just the two that were of interest in our analysis.
-These studies are also likely to have accrued citations by virtue of being amongst the first highly prominent examples of preregistered multi-laboratory replications in psychology.
-
-```{r}
-# identify how many studies citing the replication also cite the original study
-
-## extract citing articles for replication studies and original studies
-d_kleinCites <- d_repCitations %>% filter(case == 'klein')
-d_wagenmakersCites <- d_repCitations %>% filter(case == 'wagenmakers')
-d_haggerCites <- d_repCitations %>% filter(case == 'hagger')
-d_baumeisterCites <- d_citations %>% filter(case == 'baumeister')
-d_sripadaCites <- d_citations %>% filter(case == 'sripada')
-d_carusoCites <- d_citations %>% filter(case == 'caruso')
-d_carterCites <- d_citations %>% filter(case == 'carter')
-d_strackCites <- d_citations %>% filter(case == 'strack')
-
-## compute citation counts for replication studies
-hagCites <- nrow(d_haggerCites)
-wagenCites <- nrow(d_wagenmakersCites)
-kleinCites <- nrow(d_kleinCites)
-
-## compute how many articles citing the replication study also cite the relevant original study
-hagBauCocites <- sum(d_haggerCites$UT %in% d_baumeisterCites$UT)
-hagSriCocites <- sum(d_haggerCites$UT %in% d_sripadaCites$UT)
-wagenStrackCocites <- sum(d_wagenmakersCites$UT %in% d_strackCites$UT)
-kleinCartCocites <- sum(d_kleinCites$UT %in% d_carterCites$UT)
-kleinCaruCocites <- sum(d_kleinCites$UT %in% d_carusoCites$UT)
-
-## build table
-repCociteTable <- data.frame(
-  "Replication" = c("Hagger et al. (2016)", "Klein et al. (2014)", "Wagenmakers et al. (2016)"),
-  "Citation" = c(hagCites, kleinCites, wagenCites),
-  "Cocitations" = c(paste0(hagBauCocites," (Baumeister et al.
-1998) ",hagSriCocites," (Sripada et al.
-2014)"), paste0(kleinCartCocites," (Carter et al.
-2011) ",kleinCaruCocites," (Caruso et al.
-2012)"), paste0(wagenStrackCocites," (Strack et al.
-1988)"))
-  )
-```
-
-```{r}
-repCociteTable %>% kable(col.names = c("Replication study", "Citation count","Cocitations of original study"), caption = 'Citations counts for replication studies and co-citation counts to relevant original studies.', booktabs=T)
-```
-
 # Results
@@
 Article type classifications for the remaining `r notexcluded_n` articles included in the qualitative analysis are shown in Table \@ref(tab:articleTypes).
-
-## Article characteristics
 
@@
 
-Discussion
+## Citations to replication studies and co-citation of original studies
+
+Upon reviewer request, we examined citation counts for replication studies and checked whether citing articles also co-cited the relevant original study.
+To obtain the data, we downloaded bibliographic records from the Clarivate Analytics Web of Science Core Collection accessed via the University of Amsterdam on 16th April, 2021 for articles that cited each replication study (up to the study endpoint - December, 2019) and cross-checked them with our sample of articles that cited the original study.
+As shown, the replication studies also have a life of their own and they are often cited independently of the specific original study.
+Often this is easy to explain, e.g., Klein et al.
+replicated 13 original studies, not just the two that were of interest in our analysis.
+These studies are also likely to have accrued citations by virtue of being amongst the first highly prominent examples of preregistered multi-laboratory replications in psychology.
+
+```{r}
+# identify how many studies citing the replication also cite the original study
+
+## extract citing articles for replication studies and original studies
+d_kleinCites <- d_repCitations %>% filter(case == 'klein')
+d_wagenmakersCites <- d_repCitations %>% filter(case == 'wagenmakers')
+d_haggerCites <- d_repCitations %>% filter(case == 'hagger')
+d_baumeisterCites <- d_citations %>% filter(case == 'baumeister')
+d_sripadaCites <- d_citations %>% filter(case == 'sripada')
+d_carusoCites <- d_citations %>% filter(case == 'caruso')
+d_carterCites <- d_citations %>% filter(case == 'carter')
+d_strackCites <- d_citations %>% filter(case == 'strack')
+
+## compute citation counts for replication studies
+hagCites <- nrow(d_haggerCites)
+wagenCites <- nrow(d_wagenmakersCites)
+kleinCites <- nrow(d_kleinCites)
+
+## compute how many articles citing the replication study also cite the relevant original study
+hagBauCocites <- sum(d_haggerCites$UT %in% d_baumeisterCites$UT)
+hagSriCocites <- sum(d_haggerCites$UT %in% d_sripadaCites$UT)
+wagenStrackCocites <- sum(d_wagenmakersCites$UT %in% d_strackCites$UT)
+kleinCartCocites <- sum(d_kleinCites$UT %in% d_carterCites$UT)
+kleinCaruCocites <- sum(d_kleinCites$UT %in% d_carusoCites$UT)
+
+## build table
+repCociteTable <- data.frame(
+  "Replication" = c("Hagger et al. (2016)", "Klein et al. (2014)", "Wagenmakers et al. (2016)"),
+  "Citation" = c(hagCites, kleinCites, wagenCites),
+  "Cocitations" = c(paste0(hagBauCocites," (Baumeister et al.
+1998) ",hagSriCocites," (Sripada et al.
+2014)"), paste0(kleinCartCocites," (Carter et al.
+2011) ",kleinCaruCocites," (Caruso et al.
+2012)"), paste0(wagenStrackCocites," (Strack et al.
+1988)"))
+  )
+```
+
+```{r}
+repCociteTable %>% kable(col.names = c("Replication study", "Citation count","Cocitations of original study"), caption = 'Citations counts for replication studies and co-citation counts to relevant original studies.', booktabs=T)
+```
+
+# Discussion
 It has been proposed that replication studies can facilitate scientific self-correction by modifying scientists’ belief in the credibility of published findings (Ioannidis, 2012; Vazire & Holcombe, 2020; Zwaan et al., 2018); however, the extent to which this occurs in practice is unclear.
@@
 
-```{r child = "appendix.Rmd"}
-```
+```{r render_appendix, include=FALSE}
+render_appendix("appendix.Rmd")
+```
